### 2017-10-01 paper.tex
First sketch of the paper

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}

%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09

\usepackage{cite}

% Math operators
\newcommand{\scal}[2]{\left\langle #1 , #2 \right\rangle}
\DeclareMathOperator{\IR}{\mathbb{R}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\One}{\mathbbm{1}}
\DeclareMathOperator{\Ccal}{\mathcal{C}}
\DeclareMathOperator{\logsumexp}{logsumexp}
\DeclareMathOperator{\diag}{diag}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\renewcommand{\epsilon}{\varepsilon}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

\title{Over-relaxed Sinkhorn's Algorithm for Regularized Optimal Transport}

\author{
Alexis THIBAULT\\
\'Ecole Normale Sup\'erieure\\
Paris, France\\
\texttt{alexis.thibault@ens.fr}
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
This article describes a method for quickly calculating the solution to the regularized optimal transport problem. It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn's algorithm). The idea is to over-relax the Bregman projection operators, allowing for faster convergence. In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
\end{abstract}

\section{Iterative Bregman projections}

\subsection{Regularized optimal transport}

\[
K^\epsilon(\gamma) = \scal{c}{\gamma}
+ \epsilon \sum_{i,j} \gamma_{i,j} (\log(\gamma_{i,j}) - 1)
\]
\begin{equation}\label{eq:reg_ot_pb}
\gamma^* = \argmin_{\Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma^*)
\end{equation}
Constraint sets:
\begin{align*}
A_1 &: \begin{cases}
\IR^{mn} \rightarrow \IR^m \\
(A_1 x)_i = \sum_j x_{i,j}
\end{cases} &
A_2 &: \begin{cases}
\IR^{mn} \rightarrow \IR^n\\
(A_2 x)_j = \sum_i x_{i,j}
\end{cases}
\end{align*}
\begin{align*}
\Ccal_1 &= \left\{ \gamma\in\IR^{mn} \mid A_1 \gamma = \mu \right\} &
\Ccal_2 &= \left\{ \gamma\in\IR^{mn} \mid A_2 \gamma = \nu \right\}
\end{align*}
Kullback-Leibler divergence:
\[ 
KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}
\]
Bregman projection onto $\Ccal$:
\[
P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} KL(\gamma,\xi)
\]
\[\argmin_{\gamma \in \Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma) = P_{\Ccal_1 \cap \Ccal_2} (e^{-c/\epsilon})\]
Optimal transport plan: Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$.

\subsection{Sinkhorn's algorithm}

\begin{align*}
\gamma^0 &= e^{-c/\epsilon} &
\gamma^{k+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^k))
\end{align*}
Bregman's result: iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.

Hence this algorithm is correct:
\[\lim_{k\rightarrow +\infty} \gamma^k = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*\]

\subsection{Numerical version}

Projections actually correspond to diagonal scalings.
\begin{align*}
P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &
a &=  \frac{\mu}{A_1 \gamma} \\
P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &
b &= \frac{\nu}{A_2 \gamma}
\end{align*}
(pointwise division).

In practice, simply store $(a^k, b^k)$:
\[
\gamma^k = \diag(a^k) \gamma^0 \diag{b^k}
\]
\begin{align*}
a^{k+1} &= \frac{\mu}{\gamma^0 b^k} &
b^{k+1} &= \frac{\nu}{^t \gamma^0 a^{k+1}} 
\end{align*}

Efficient implementation: matrix operations on GPU.

\section{Over-relaxed Sinkhorn's algorithm}

\subsection{Over-relaxed projections}

Let us define the $\omega$-over-relaxed projection operator by:
\begin{equation}\label{eq:def_or_proj}
\log P^\omega_{\Ccal_k}(\gamma) = (1-\omega) \log \gamma + \omega \log P_{\Ccal_k}(\gamma),
\end{equation}
where the logarithm is taken coordinate-wise.
Note that $P^0 = Id$, $P^1 = P$, and $P^2$ is an involution (this is non-trivial, and uses the fact that $\Ccal_k$ is an affine subspace).

\subsection{Lyapunov function}
Let $\gamma^*$ denote the solution of the regularized OT problem.
The function $F$ is defined as:
\begin{equation}\label{eq:lyapunov_function}
F(\gamma) = KL(\gamma^*, \gamma)
\end{equation}
We shall use it as a Lyapunov function, so that $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.

Note that it is not necessary to know $\gamma^*$ in order to calculate the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$, as shown by the following lemma.
\begin{lemma}\label{lemma:lyapunov_decrease}
	Take $\gamma$ in $\IR^{mn}_{+*}$. The decrease in value of the Lyapunov function can be calculated with the following formula:
	\begin{equation} \label{eq:kl_diff_scal}
	F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma)) = 
	\scal{\mu}{\varphi_\omega \left(\frac{A_1 \gamma}{\mu}\right)},
	\end{equation}
	where
	\begin{equation}
	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega e^x
	\end{equation}
	is a real function, applied coordinate-wise.
\end{lemma}
Thus, calculating the decrease in value of the function $F$ is relatively cheap and straightforward.

\subsection{New algorithm}

\begin{theorem}
	Let $\theta$ and $\omega$ be such that $1\le \theta < \omega < 2$. Let $U$ be the open set defined by:
	\begin{equation}\label{eq:open_set_U}
	U = \left\{
	\gamma \in \IR_{+*}^{nm} \mid
	F(P^\omega_{\Ccal_1}(\gamma)) < F(\gamma)
	\text{ and }
	F(P^\omega_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma))) < F(P^\theta_{\Ccal_1}(\gamma))
	\right\}.
	\end{equation}
	Set the initial value $\gamma^0 = e^{-c/\epsilon}$, and iterate:
	\begin{align*}
	\gamma^{k+1} =
	\begin{cases}
	P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^k)) & \text{if } \gamma \in U \\
	P_{\Ccal_2}(P_{\Ccal_1})(\gamma^k) & \text{otherwise.}
	\end{cases}
	\end{align*}
	Then the iterates $(\gamma^k)$ converge to $\gamma^*$.
\end{theorem}
{\color{red} $\longrightarrow$ Has to calculate 3 projections sometimes. This result is weaker than the one from my report, but easier to prove. Maybe we could avoid having $U$, and rather perform the test independently on each variable ? (Or we could find a cheaper test).}
\begin{remark}
	Given lemma \ref{lemma:lyapunov_decrease}, we can see that $\gamma^* \in U$. Therefore, iterations eventually all use the over-relaxed operators.
\end{remark}

\begin{proof}
	(New proof, using the Lyapunov function. Much shorter and cleaner than the one from my report.)
\end{proof}

\section{Local rate of convergence ?}

\section{Experimental results}
{\color{red} TODO}

\section*{Acknowledgments}

%\subsubsection*{References}

\bibliographystyle{apalike}
\bibliography{references}

\end{document}

### 2017-10-06 paper.tex
Add proof of global convergence

@@
 \]
+Two discrete probability measures $\mu^k \in \IR_{+*}^{n_k}$.
 \begin{equation}\label{eq:reg_ot_pb}
@@
 A_1 &: \begin{cases}
-\IR^{mn} \rightarrow \IR^m \\
+\IR^{n_1 n_2} \rightarrow \IR^{n_1} \\
 (A_1 x)_i = \sum_j x_{i,j}
@@
 A_2 &: \begin{cases}
-\IR^{mn} \rightarrow \IR^n\\
+\IR^{n_1 n_2} \rightarrow \IR^{n_2}\\
 (A_2 x)_j = \sum_i x_{i,j}
@@
 \begin{align*}
-\Ccal_1 &= \left\{ \gamma\in\IR^{mn} \mid A_1 \gamma = \mu \right\} &
-\Ccal_2 &= \left\{ \gamma\in\IR^{mn} \mid A_2 \gamma = \nu \right\}
+\Ccal_k &= \left\{ \gamma\in\IR^{n_1 n_2} \mid A_k \gamma = \mu^k \right\}
 \end{align*}
@@
 \gamma^0 &= e^{-c/\epsilon} &
-\gamma^{k+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^k))
+\gamma^{l+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))
 \end{align*}
@@
 Hence this algorithm is correct:
-\[\lim_{k\rightarrow +\infty} \gamma^k = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*\]
+\[\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*\]
 
@@
 P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &
-a &=  \frac{\mu}{A_1 \gamma} \\
+a &=  \frac{\mu^1}{A_1 \gamma} \\
 P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &
-b &= \frac{\nu}{A_2 \gamma}
+b &= \frac{\mu^2}{A_2 \gamma}
 \end{align*}
@@
 
-In practice, simply store $(a^k, b^k)$:
+In practice, simply store $(a^l, b^l)$:
 \[
-\gamma^k = \diag(a^k) \gamma^0 \diag{b^k}
+\gamma^l = \diag(a^l) \gamma^0 \diag{b^l}
 \]
 \begin{align*}
-a^{k+1} &= \frac{\mu}{\gamma^0 b^k} &
-b^{k+1} &= \frac{\nu}{^t \gamma^0 a^{k+1}} 
+a^{l+1} &= \frac{\mu^1}{\gamma^0 b^l} &
+b^{l+1} &= \frac{\mu^2}{^t \gamma^0 a^{l+1}} 
 \end{align*}
@@
 \end{equation}
-We shall use it as a Lyapunov function, so that $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+It shall be used as a Lyapunov function, so that $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+
+\begin{lemma} \label{lemma:KL_compact}
+	For any $M \in \IR_+^*$, the sublevel set $\left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
+\end{lemma}
+\begin{proof}
+	{\color{red} TODO}
+\end{proof}
 
@@
 	\begin{equation}
-	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega e^x
+	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega \log x
 	\end{equation}
@@
 	\begin{align*}
-	\gamma^{k+1} =
+	\gamma^{l+1} =
 	\begin{cases}
-	P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^k)) & \text{if } \gamma \in U \\
+	P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l)) & \text{if } \gamma \in U \\
 	P_{\Ccal_2}(P_{\Ccal_1})(\gamma^k) & \text{otherwise.}
@@
 	\end{align*}
-	Then the iterates $(\gamma^k)$ converge to $\gamma^*$.
+	Then the iterates $(\gamma^l)$ converge to $\gamma^*$.
 \end{theorem}
@@
 \begin{remark}
-	Given lemma \ref{lemma:lyapunov_decrease}, we can see that $\gamma^* \in U$. Therefore, iterations eventually all use the over-relaxed operators.
+	Given lemma \ref{lemma:lyapunov_decrease}, it can be seen that $\gamma^* \in U$. Therefore, iterations eventually all use the over-relaxed operators.
 \end{remark}
 
+\begin{lemma}
+	\label{lemma:trivial_intersection}
+	Let us take $\gamma^0$ in $\IR_{+*}^{n_1 n_2}$,
+	and denote
+	\[
+	S = \left\{
+	\diag(a) \gamma^0 \diag(b),\quad
+	(a,b) \in \IR_{+*}^{n_1 + n_2}
+	\right\}
+	\]
+	the set of matrices that are diagonally similar to $\gamma^0$.
+	Then the set $S \cap \Ccal_1 \cap \Ccal_2$ contains exactly one element $\gamma^* = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0)$.
+\end{lemma}
+
+\begin{lemma}\label{lemma:F_P_theta}
+	Let $1\le \theta < \omega$. Then, for any $\gamma \in \IR_{+*}^{nm}$, one has
+	\begin{equation}\label{eq:F_P_theta}
+	F(P^\theta_{\Ccal_k}(\gamma)) \le F(P^\omega_{\Ccal_k}(\gamma)).
+	\end{equation}
+	Moreover, equality occurs if and only if $\gamma \in \Ccal_k$.
+\end{lemma}
 \begin{proof}
+	Thanks to lemma \ref{lemma:lyapunov_decrease}, one knows that
+	\[
+	F(P^\theta_{\Ccal_k}(\gamma)) - F(P^\omega_{\Ccal_k}(\gamma))
+	= \scal{\mu}{(\varphi_\omega - \varphi_\theta) \left( \frac{A_k \gamma}{\mu} \right) } .
+	\]
+	Moreover, $t \in [1,\infty) \mapsto \varphi_t(x)$ is nonincreasing:
+	\[
+	\frac{d}{dt} \varphi_t(x) = \log x (x^{1-t} - 1).
+	\]
+	For $x\neq 1$, it is even strictly decreasing.
+	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_1 \gamma = \mu$.
+\end{proof}
+
+\begin{proof}[Proof of the theorem]
 	(New proof, using the Lyapunov function. Much shorter and cleaner than the one from my report.)
+	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^l)$ are thus diagonally similar to $\gamma^0$:
+	\[
+	\forall l\ge0,\quad \gamma^l \in S
+	\]
+
+	The set $U$ is constructed in such a way that, by applying lemma \ref{lemma:F_P_theta} several times, one ensures that the sequence $(F(\gamma^l))$ is nonincreasing.
+Moreover, it is strictly decreasing as long as $\gamma^l \not \in \Ccal_1 \cap \Ccal_2$.
+	Hence, 
+	\[
+	F(\gamma^l) \le F(\gamma^0).
+	\]
+	By lemma \ref{lemma:KL_compact}, the sequence $(\gamma^l)$ is thus precompact.
+Let us show that any accumulation point $\zeta$ belongs to $\Ccal_1 \cap \Ccal_2$.
+Since $S$ is closed in $\IR_{+*}^{n_1 n_2}$, by lemma \ref{lemma:trivial_intersection} this would mean that $\zeta = \gamma^*$, and would be sufficient to prove the convergence.
+
+	Let $(\gamma^{k_l})$ be a subsequence of $(\gamma^l)$ that converges to $\zeta$.
+	Let us denote, for convenience, $T_1 = P_{\Ccal_2}^\theta \circ P_{\Ccal_1}^\theta$ and $T_2 = P_{\Ccal_2} \circ P_{\Ccal_1}$ the two operators used in the iteration, and $T$ the compound operator, so that $T_{|U} \equiv T_{1|U}$ and $T_{|U^c} \equiv T_{2|U^c}$.
+	\begin{itemize}
+		\item If $\zeta \in U$, then since $U$ is open, all $(\gamma^{k_l})$ are eventually in $U$ as well.
+		Thus, for any large enough $l$, $ T(\gamma^{k_l}) = T_1(\gamma^{k_l}) $. By continuity of $T_1$ on $U$, since $(\gamma^{k_l})$ converges to $\zeta$, 
+		\[T(\zeta) = \lim_{l\rightarrow \infty} T(\gamma^{k_l}).\] Thus
+		 \[F(T(\zeta)) = \lim_{l \rightarrow \infty} F(T(\gamma^{k_l})) \ge \lim_{l\rightarrow \infty} F(\gamma^{k_{l+1}}) .\]
+		 \[
+		 F(T(\zeta)) \ge F(\zeta)
+		 \] Since $\zeta$ is inside the set $U$, which ensures strict decrease of the Lyapunov function, this is absurd.
+		 \item If $\zeta \in U^c$ and an infinite number of $(\gamma^{k_l})$ are in $U^c$, then the same argument of continuity applied to $T_2$ yields
+		 \[
+		 F(T(\zeta)) = F(\zeta).
+		 \]
+		 The function $\varphi_1$ from lemma \ref{lemma:lyapunov_decrease} is strictly positive except at $x=1$.
+Since the case of equality occurs for both projections, this ensures that $\zeta \in \Ccal_1 \cap \Ccal_2$.
+
+		 \item If $\zeta \in U^c$ and all but finitely many $\gamma^{k_l}$ are in $U$, then $\zeta$ is on the border $\partial U$. In particular it belongs to $\bar{U}$, and therefore
+		 \[
+		 F(P^\omega_{\Ccal_1}(\zeta)) \le F(\zeta) \quad
+		 \text{ and } \quad
+		 F(P^\omega_{\Ccal_2}(P^\theta_{\Ccal_1}(\zeta))) \le F(P^\theta_{\Ccal_1}(\zeta)).
+		 \]
+		 By lemma \ref{lemma:F_P_theta}, the inequality \begin{equation} \label{eq:ineq1}
+		 F(P_{\Ccal_1}^\theta(\zeta)) \le F(P_{\Ccal_1}^\omega(\zeta)) \end{equation}
+		 is valid; it is strict unless $\zeta \in \Ccal_1$.
+		 The same lemma yields
+		 \begin{equation} \label{eq:ineq2}
+		 F(P_{\Ccal_2}^\theta(P_{\Ccal_1}^\theta(\zeta)) \le F(P_{\Ccal_2}^\omega(P_{\Ccal_1}^\theta(\zeta)), \end{equation}
+		 with a strict inequality unless $P_{\Ccal_1}^\theta(\zeta) \in \Ccal_2$.
+		 To sum up, we have the series of inequalities:
+		 \[
+		 F(\zeta) 
+		 \overset{\zeta \in \bar{U}}{\ge}
+		 F(P^\omega_{\Ccal_1}(\zeta))
+		 \overset{\text{(\ref{eq:ineq1})}}{\ge}
+		 F(P^\theta_{\Ccal_1}(\zeta))
+		 \overset{\zeta \in U}{\ge}
+		 F(P_{\Ccal_2}^\omega(P_{\Ccal_1}^\theta(\zeta)))
+		 \overset{\text{(\ref{eq:ineq2})}}{\ge}
+		 F(P_{\Ccal_2}^\theta(P_{\Ccal_1}^\theta(\zeta)))
+		 \]
+
+		 Yet, by continuity of $T_1 = P_{\Ccal_2}^\theta \circ P_{\Ccal_1}^\theta $ on $\bar{U}$, we have
+		 \[
+		 F(T_1(\zeta)) = F(\zeta).
+		 \]
+		 Thus both (\ref{eq:ineq1}) and (\ref{eq:ineq2}) fall in the case of equality, which means $\zeta \in \Ccal_1$ and $P^\theta_{\Ccal_1}(\zeta) = \zeta \in \Ccal_2$.
+	\end{itemize}
 \end{proof}

### 2017-10-09 paper.tex
Redaction

@@
 
-\section{Iterative Bregman projections}
+\section{Introduction}
+BLABLA to write
 
-\subsection{Regularized optimal transport}
-
-\[
-K^\epsilon(\gamma) = \scal{c}{\gamma}
-+ \epsilon \sum_{i,j} \gamma_{i,j} (\log(\gamma_{i,j}) - 1)
-\]
-Two discrete probability measures $\mu^k \in \IR_{+*}^{n_k}$.
-\begin{equation}\label{eq:reg_ot_pb}
-\gamma^* = \argmin_{\Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma^*)
-\end{equation}
-Constraint sets:
+We consider two discrete probability measures $\mu^k \in \IR_{+*}^{n_k}$.
+Let us define the two following linear operators
 \begin{align*}
@@
 \IR^{n_1 n_2} \rightarrow \IR^{n_2}\\
-(A_2 x)_j = \sum_i x_{i,j}
+(A_2 x)_j = \sum_i x_{i,j},
 \end{cases}
 \end{align*}
+as well as the convex constraint sets
 \begin{align*}
-\Ccal_k &= \left\{ \gamma\in\IR^{n_1 n_2} \mid A_k \gamma = \mu^k \right\}
+\Ccal_k &= \left\{ \gamma\in\IR^{n_1 n_2} \mid A_k \gamma = \mu^k \right\}.
 \end{align*}
-Kullback-Leibler divergence:
+Given a cost matrix $c$ where $c_{ij}$ represents the cost of moving mass from $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the following minimization problem:
+$$\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_+} \langle c,\gamma\rangle:=\sum_{i,j}c_{i,j}\gamma_{i,j}.$$
+This is a linear programing problem whose resolution is intractable for large scale problems.
+
+\subsection{Regularized optimal transport}
+
+In \cite{cuturi13}, it has been proposed to regularize this problem by adding a strictly convex entropy regularization:
+\begin{equation}\label{ROT}
+\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_{+*}}K^\epsilon(\gamma) := \scal{c}{\gamma} 
++ \epsilon \sum_{i,j} \gamma_{i,j} (\log(\gamma_{i,j}) - 1)
+,\end{equation}
+with $\epsilon>0$. Considering, the Kullback-Leibler divergence:
 \[ 
-KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}
+KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j},
 \]
-Bregman projection onto $\Ccal$:
+it was shown in \cite{benamou15}  that the unique minimizer of problem \eqref{ROT} writes
+\begin{equation}\label{eq:reg_ot_pb}
+\gamma^* = \argmin_{\Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma^)= P_{\Ccal_1 \cap \Ccal_2} (e^{-c/\epsilon}),
+\end{equation}
+where $P_{\Ccal}$ is the  Bregman projection onto $\Ccal$ defined as:
 \[
-P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} KL(\gamma,\xi)
+P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} KL(\gamma,\xi).
 \]
-\[\argmin_{\gamma \in \Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma) = P_{\Ccal_1 \cap \Ccal_2} (e^{-c/\epsilon})\]
-Optimal transport plan: Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$.
+The regularized  optimal transport plan $\gamma^*$ is thus the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$.
 
 \subsection{Sinkhorn's algorithm}
-
+Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
+Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} can be considered to compute the regularized  plan:
 \begin{align*}
 \gamma^0 &= e^{-c/\epsilon} &
-\gamma^{l+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))
+\gamma^{l+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l)),
 \end{align*}
-Bregman's result: iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
+which checks 
+$\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
 
-Hence this algorithm is correct:
-\[\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*\]
-
-\subsection{Numerical version}
-
-Projections actually correspond to diagonal scalings.
+In the discrete setting,  projections actually correspond to diagonal scalings.
 \begin{align*}
@@
 \end{align*}
-(pointwise division).
+where the division is considered pointwise. 
+To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n2}$ and realize
 
-In practice, simply store $(a^l, b^l)$:
-\[
-\gamma^l = \diag(a^l) \gamma^0 \diag{b^l}
-\]
 \begin{align*}
@@
 \end{align*}
-
-Efficient implementation: matrix operations on GPU.
+where the transport matrix can be recovered as :
+$\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$ 
+Efficient parallel computations can be considered \cite{cuturi13} and one can even reach real time computation for large scale problem for certain class of cost matrices $c$ \cite{Solomon2015}. 
+For small values of the parameter $\epsilon$, numerical issues can arise and a stabilization of the algorithm is necessary \cite{2016arXiv160705816C}.
+The convergence of the process can nevertheless be very slow  in the setting $\epsilon\to 0$ .
+\subsection{Contributions}
+In this paper, we consider an over relaxation scheme  designed to accelerate  the Sinkhorn's algorithm.
+We first present and show the convergence of our algorithm in section 2.
+In section 3, we analyze the local convergence rate of the algorithm to justify the acceleration.
+We finally demonstrate numerically  in Section 4 the good behavior of our method, where larger accelerations are observed for decreasing values of $\epsilon$.
 
@@
 	F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma)) = 
-	\scal{\mu}{\varphi_\omega \left(\frac{A_1 \gamma}{\mu}\right)},
+	\scal{\mu^k}{\varphi_\omega \left(\frac{A_k \gamma}{\mu^k}\right)},
 	\end{equation}
@@
 \end{theorem}
-{\color{red} $\longrightarrow$ Has to calculate 3 projections sometimes.
+{\color{red} 
+	In practice the test would be: check whether the $\omega$-relaxed projection of $\zeta$ on $\Ccal_1$ decreases the value of
+	$F$; if so, calculate $P_{\Ccal_1}^\theta(\zeta)$ (don't update $\zeta$ yet) and check that its $\omega$-relaxed projection on $\Ccal_2$ decreases the value of $F$ again. \\
+	$\longrightarrow$ Has to calculate 3 projections sometimes.
 This result is weaker than the one from my report, but easier to prove.
@@
 	F(P^\theta_{\Ccal_k}(\gamma)) - F(P^\omega_{\Ccal_k}(\gamma))
-	= \scal{\mu}{(\varphi_\omega - \varphi_\theta) \left( \frac{A_k \gamma}{\mu} \right) } .
+	= \scal{\mu^k}{(\varphi_\omega - \varphi_\theta) \left( \frac{A_k \gamma}{\mu^k} \right) } .
 	\]
@@
 		 \[
-		 F(T_1(\zeta)) = F(\zeta).
+		 F(P_{\Ccal_2}^\theta(P_{\Ccal_1}^\theta(\zeta))) = F(\zeta).
 		 \]

### 2017-10-09 paper.tex
Test commit

@@
 \subsection{Contributions}
-In this paper, we consider an over relaxation scheme  designed to accelerate  the Sinkhorn's algorithm.
+In this paper, we consider an over-relaxation scheme  designed to accelerate  the Sinkhorn's algorithm.
 We first present and show the convergence of our algorithm in section 2.
@@
 where the logarithm is taken coordinate-wise.
-Note that $P^0 = Id$, $P^1 = P$, and $P^2$ is an involution (this is non-trivial, and uses the fact that $\Ccal_k$ is an affine subspace).
+Note that $P_{\Ccal_k}^0 = Id$, $P_{\Ccal_k}^1 = P_{\Ccal_k}$, and $P_{\Ccal_k}^2$ is an involution (this is non-trivial, and uses the fact that $\Ccal_k$ is an affine subspace).
 

### 2017-10-09 paper.tex
Test commit

@@
 \end{align*}
-as well as the convex constraint sets
+as well as thenlinear  constraint sets
 \begin{align*}
@@
 \end{align*}
-Given a cost matrix $c$ where $c_{ij}$ represents the cost of moving mass from $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the following minimization problem:
+Given a cost matrix $c$ where $c_{ij}$ represents the cost of moving mass $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the following minimization problem:
 $$\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_+} \langle c,\gamma\rangle:=\sum_{i,j}c_{i,j}\gamma_{i,j}.$$

### 2017-10-09 paper.tex
corrections intro

@@
 \end{align*}
-Given a cost matrix $c$ where $c_{ij}$ represents the cost of moving mass $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the following minimization problem:
+Given a cost matrix $c$, where $c_{ij}$ represents the cost of moving mass $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the following minimization problem:
 $$\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_+} \langle c,\gamma\rangle:=\sum_{i,j}c_{i,j}\gamma_{i,j}.$$
@@
 \min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_{+*}}K^\epsilon(\gamma) := \scal{c}{\gamma} 
-+ \epsilon \sum_{i,j} \gamma_{i,j} (\log(\gamma_{i,j}) - 1)
++ \epsilon KL(\gamma,\mathbf{1})
 ,\end{equation}
-with $\epsilon>0$. Considering, the Kullback-Leibler divergence:
+with $\epsilon>0$, $\mathbf{1}$ is the matrix of size $n_1\times n_2$ full of ones and the Kullback-Leibler divergence is:
 \[ 
-KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j},
+KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
 \]
-it was shown in \cite{benamou15}  that the unique minimizer of problem \eqref{ROT} writes
+It was shown in \cite{benamou15}  that the unique minimizer of problem \eqref{ROT} writes
 \begin{equation}\label{eq:reg_ot_pb}
-\gamma^* = \argmin_{\Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma^)= P_{\Ccal_1 \cap \Ccal_2} (e^{-c/\epsilon}),
+\gamma^* = \argmin_{\Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma)= P_{\Ccal_1 \cap \Ccal_2} (e^{-c/\epsilon}),
 \end{equation}
@@
 where the division is considered pointwise. 
-To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n2}$ and realize
+To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and realize
 

### 2017-10-10 paper.tex
Suite corrections

@@
 with $\epsilon>0$, $\mathbf{1}$ is the matrix of size $n_1\times n_2$ full of ones and the Kullback-Leibler divergence is:
-\[ 
+\begin{equation}\label{KL}
 KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
-\]
+\end{equation}
 It was shown in \cite{benamou15}  that the unique minimizer of problem \eqref{ROT} writes
@@
 In the discrete setting,  projections actually correspond to diagonal scalings.
-\begin{align*}
+\begin{align}\label{scaling}
 P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &
@@
 P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &
-b &= \frac{\mu^2}{A_2 \gamma}
-\end{align*}
+b &= \frac{\mu^2}{A_2 \gamma}\nonumber
+\end{align}
 where the division is considered pointwise. 
@@
 where the logarithm is taken coordinate-wise.
-Note that $P_{\Ccal_k}^0 = Id$, $P_{\Ccal_k}^1 = P_{\Ccal_k}$, and $P_{\Ccal_k}^2$ is an involution (this is non-trivial, and uses the fact that $\Ccal_k$ is an affine subspace).
+Note that $P_{\Ccal_k}^0 = Id$, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ gives the Sinkhorn algorithm, and $P_{\Ccal_k}^2$ is an involution (this is non-trivial, and uses the fact that $\Ccal_k$ is an affine subspace).
 
@@
 \end{lemma}
+\begin{proof}
+From \eqref{KL}, we have $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$.
+The result can then be deduced from relations \eqref{eq:def_or_proj} and \eqref{scaling}.
+\end{proof}
 Thus, calculating the decrease in value of the function $F$ is relatively cheap and straightforward.
-
+In Figure \ref{phi_omega}, we illustrate the functions  $\varphi_\omega$ for several values of $\omega$.
+Notice that for the Sinkhorn algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always non negative.
+\begin{figure}[ht!]
+\begin{center}
+\includegraphics[width=11cm]{phi_omega.png}
+\caption{\label{phi_omega} Plot of functions $\varphi_\omega(x)$ for different values $1\leq \omega\leq 2$ and   $x\in[10^{-2}, 10^1]$.}
+\end{center}
+\end{figure}
 \subsection{New algorithm}

### 2017-10-10 paper.tex
Suite corrections

@@
 
+As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  realizes alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
+The idea developped in this paper is to realize over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c)
+
+\begin{figure}[ht!]
+\begin{center}
+\begin{tabular}{ccc}
+%\includegraphics[width=3cm]{}&\includegraphics[width=3cm]{}&\includegraphics[width=3cm]{}\\
+&TODO: mettre les figures du beamer\\
+&en remplacant les $x$ par des $\gamma$\\
+(a)&(b)&(c)
+\end{tabular}
+\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the Sinkhorn algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
+In order to accelerate the convergence rate, over-relaxed projections are here considered (c).}
+\end{center}
+\end{figure}
 \subsection{Over-relaxed projections}
@@
 
-\begin{lemma}
+\begin{lemma}[{\color{red} Mettre la ref d'un papier de Marco}]
 	\label{lemma:trivial_intersection}

### 2017-10-10 paper.tex
Suite corrections

@@
 \begin{proof}[Proof of the theorem]
-	(New proof, using the Lyapunov function. Much shorter and cleaner than the one from my report.)
 	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^l)$ are thus diagonally similar to $\gamma^0$:

### 2017-10-10 paper.tex
Suite intro

@@
 \section{Introduction}
-BLABLA to write
+Optimal Transport is an efficient and flexible tool to compare two probability distributions which has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
+The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn's algorithm \cite{sinkhorn64}   scaling with high dimensional data.
+Regularized optimal transport have thus been intensively used  in  Machine Learning with applications such as   Geodesic PCA \cite{seguy2015principal}, domain adaptation \cite{2015arXiv150700504C}, data fitting \cite{2015arXiv150605439F},  training of Boltzmann Machine \cite{NIPS2016_6248}  or dictionary learning \cite{Rolet2016,2017arXiv170801955S}.
 
+The computation of optimal transport between two data relies on the estimation of an optimal transport matrix which entries represent the quantity of mass transported between  data locations. 
+Regularization of optimal transport with strictly convex regularizers \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
+Hence, for particular purposes such as color interpolation \cite{Rabin2015} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
+In this setting,  the regularized transport problem can be ill-conditioned. This is the issue  we want to tackle here.
+Before going into further details, we now briefly introduce the main notations and concepts used all along this article.
+
+\subsection{Discrete Optimal Transport}
 We consider two discrete probability measures $\mu^k \in \IR_{+*}^{n_k}$.
@@
 \end{align*}
-as well as thenlinear  constraint sets
+as well as the linear  constraint sets
 \begin{align*}
@@
 For small values of the parameter $\epsilon$, numerical issues can arise and a stabilization of the algorithm is necessary \cite{2016arXiv160705816C}.
-The convergence of the process can nevertheless be very slow  in the setting $\epsilon\to 0$ .
-\subsection{Contributions}
-In this paper, we consider an over-relaxation scheme  designed to accelerate  the Sinkhorn's algorithm.
+The convergence of the process can nevertheless be very slow  in the setting $\epsilon\to 0$.
+\subsection{Related works }
+The introduction of relaxation variables through heavy ball approaches \cite{POLYAK19641} has recently gained in popularity  to speed up the convergence of algorithms optimizing convex \cite{2014arXiv1412.7457G} or non convex \cite{Zavriev1993,2016arXiv160609070O} problems.
+Such schemes have also been empirically considered to accelerate Sinkhorn's algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
+The convergence of these algorithms is nevertheless not studied yet in the context of regularized optimal transport.
+
+\subsection{Overview and contributions}
+In this paper, we consider an over-relaxation scheme designed to accelerate   Sinkhorn's algorithm.
 We first present and show the convergence of our algorithm in section 2.

### 2017-10-10 paper.tex
Suite intro

@@
 \texttt{alexis.thibault@ens.fr}
+ \And
+L\'enaic CHIZAT\\
+\'Ecole Normale Sup\'erieure\\ Paris Dauphine (PSL research University)\\
+Paris, France\\
+\texttt{lenaic.chizat@ens.fr}
+ \AND
+Charles DOSSAL\\
+INSA Toulouse\\
+Institut de Math\'ematiques de Toulouse\\
+Toulouse,  France\\
+\texttt{dossal@insa-toulouse.fr}
+\And 
+Nicolas PAPADAKIS\\
+CNRS, Institut de Math\'ematiques de Bordeaux\\
+Talence, France\\
+\texttt{alexis.thibault@ens.fr}
 } 
-
 % The \author macro works with any number of authors. There are two commands
@@
 Regularization of optimal transport with strictly convex regularizers \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
-Hence, for particular purposes such as color interpolation \cite{Rabin2015} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
+Hence, for particular purposes such as color interpolation \cite{Rabin2014} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
 In this setting,  the regularized transport problem can be ill-conditioned. This is the issue  we want to tackle here.

### 2017-10-10 paper.tex
Suite intro

@@
 \end{align*}
-Given a cost matrix $c$, where $c_{ij}$ represents the cost of moving mass $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the following minimization problem:
+Given a cost matrix $c$, where $c_{ij}$ represents the cost of moving mass $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the estimation of an optimal transport matrix $\gamma$ solution of:
 $$\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_+} \langle c,\gamma\rangle:=\sum_{i,j}c_{i,j}\gamma_{i,j}.$$
@@
 Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
-Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} can be considered to compute the regularized  plan:
+Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} can be considered to compute the regularized  transport matrix:
 \begin{align*}

### 2017-10-11 paper.tex
Suite intro

@@
 Talence, France\\
-\texttt{alexis.thibault@ens.fr}
+\texttt{nicolas.papadakis@math.u-bordeaux.fr}
 } 
@@
 \end{equation}
-It was shown in \cite{benamou15}  that the unique minimizer of problem \eqref{ROT} writes
+It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$:
 \begin{equation}\label{eq:reg_ot_pb}
@@
 \]
-The regularized  optimal transport plan $\gamma^*$ is thus the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$.
 
@@
 Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
-Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} can be considered to compute the regularized  transport matrix:
+Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} that realizes alternate Bregman projections, can be considered to compute the regularized  transport matrix:
 \begin{align*}
@@
 \end{align*}
-which checks 
+and we have 
 $\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
@@
 where the division is considered pointwise. 
-To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and realize
-
+To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and iterates
 \begin{align*}
@@
 
-\begin{lemma}[{\color{red} Mettre la ref d'un papier de Marco}]
+\begin{lemma}[\cite{cuturi13}]
 	\label{lemma:trivial_intersection}

### 2017-10-11 paper.tex
Suite intro

@@
 
+{\color{red} \begin{proposition} 
+State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
+\end{proposition}}
 \subsection{Lyapunov function}
@@
 \end{figure}
-\subsection{New algorithm}
+\subsection{Proposed algorithm}
 
@@
 	P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l)) & \text{if } \gamma \in U \\
-	P_{\Ccal_2}(P_{\Ccal_1})(\gamma^k) & \text{otherwise.}
+	P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l)) & \text{otherwise.}
 	\end{cases}
@@
 \end{remark}
+{\color{blue} It could be good to show on typical experiments how many times the test fails.
+If the test never fails after a few iterations, we could justify that doing a few Sinkhorn iteration and then always the over-relaxed algorithm is fine in practice.}
 
@@
 \section*{Acknowledgments}
+This study has been carried out with financial support from the French State, managed by the French National Research Agency (ANR) in the frame of the  GOTMI project (ANR-16-CE33-0010-01).
 

### 2017-10-11 paper.tex
Suite intro

@@
 \usepackage{amsthm}
-\usepackage{mathtools}
+\usepackage{mathtools,enumitem}
 
@@
 	Let us denote, for convenience, $T_1 = P_{\Ccal_2}^\theta \circ P_{\Ccal_1}^\theta$ and $T_2 = P_{\Ccal_2} \circ P_{\Ccal_1}$ the two operators used in the iteration, and $T$ the compound operator, so that $T_{|U} \equiv T_{1|U}$ and $T_{|U^c} \equiv T_{2|U^c}$.
-	\begin{itemize}
-		\item If $\zeta \in U$, then since $U$ is open, all $(\gamma^{k_l})$ are eventually in $U$ as well.
+	\begin{enumerate}[align=right,labelsep=10pt,leftmargin=0.5cm]
+		\item[$\bullet$] If $\zeta \in U$, then since $U$ is open, all $(\gamma^{k_l})$ are eventually in $U$ as well.
 		Thus, for any large enough $l$, $ T(\gamma^{k_l}) = T_1(\gamma^{k_l}) $. By continuity of $T_1$ on $U$, since $(\gamma^{k_l})$ converges to $\zeta$, 
@@
 		 \] Since $\zeta$ is inside the set $U$, which ensures strict decrease of the Lyapunov function, this is absurd.
-		 \item If $\zeta \in U^c$ and an infinite number of $(\gamma^{k_l})$ are in $U^c$, then the same argument of continuity applied to $T_2$ yields
+		\item[$\bullet$] If $\zeta \in U^c$ and an infinite number of $(\gamma^{k_l})$ are in $U^c$, then the same argument of continuity applied to $T_2$ yields
 		 \[
@@
 
-		 \item If $\zeta \in U^c$ and all but finitely many $\gamma^{k_l}$ are in $U$, then $\zeta$ is on the border $\partial U$. In particular it belongs to $\bar{U}$, and therefore
+		 \item[$\bullet$] If $\zeta \in U^c$ and all but finitely many $\gamma^{k_l}$ are in $U$, then $\zeta$ is on the border $\partial U$. In particular it belongs to $\bar{U}$, and therefore
 		 \[
@@
 		 Thus both (\ref{eq:ineq1}) and (\ref{eq:ineq2}) fall in the case of equality, which means $\zeta \in \Ccal_1$ and $P^\theta_{\Ccal_1}(\zeta) = \zeta \in \Ccal_2$.
-	\end{itemize}
+	\end{enumerate}
 \end{proof}

### 2017-10-11 paper.tex
Suite intro

@@
 
-{\color{red} \begin{proposition} 
+{\color{blue} \begin{proposition} 
 State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
@@
 
+\section{Conclusion}
+
+Cheaper test, how chosing $\omega$.
+
+{\color{blue}
+From experiments, it seems possible to design a good guess for $\omega$ from $c$ and $\epsilon$ only. For $K=e^{-c/\epsilon}$, let us  define:
+    $T=(diag(K\mathbf{1}))^{-1}K(diag(K^T\mathbf{1}))^{-1}K^T$. The matrix $T$ is positive semidefinite and checks
+$\rho(T)=1$. Moreover  $T$ has  only one  eigenvalue $1$ associated to the eigenvector $\mathbf{1}$.
+    Denoting as $\lambda$ the second larger eigenvalue of $T$, one can set 
+    $\omega=\frac{2}{\lambda}(1-\sqrt{1-\lambda})$.}
+
 \section*{Acknowledgments}

### 2017-10-11 paper.tex
Pb dual

--- 2017-10-11 paper.tex
@@
 \section{Local rate of convergence ?}
+Some material for adapting Lenaic results:
+
+The Sinkhorn algorithm can be interpreted as an alternate maximisation algorithm on the dual of the regularized optimal transport problem.
+The dual problem of \eqref{ROT} is:
+\begin{equation}\label{DROT}\max_{\alpha,\beta}\langle \alpha,\mu^1\rangle+\langle \beta,\mu^2\rangle-\epsilon\sum_{i,j}e^{(\alpha_i+\beta_j-c_{i,j})/\epsilon}\end{equation}
+It is concave and continuously differentiable for all $(a,b)\in\IR^{n_1+n_2}$.
+Alternate maximization then converges and we recover, for $a_i=e^{\alpha_i/\epsilon}$, $b_j=e^{\beta_j/\epsilon}$ and $\gamma^0_{i,j}=e^{-c_{i,j}/\epsilon}$:
+\begin{align*}
+a^{l+1} &= \frac{\mu^1}{\gamma^0 b^l} &
+b^{l+1} &= \frac{\mu^2}{^t \gamma^0 a^{l+1}} 
+\end{align*}
+
+If $(\alpha^*,\beta^*)$ is a solution of \eqref{DROT}, then all the 
+other optimum can be written as $(\alpha^*+\kappa,\beta^*-\kappa)$, which means that the problem admits a unique solution, up to a translation with constant $\kappa\in\IR$.
+
+The Hessian matrix of the functional in \eqref{DROT} at point $(\alpha,\beta)$ reads, for $a_i=e^{\alpha_i/\epsilon}>0$ and $b_j=e^{\beta_j/\epsilon}>0$:
+
+$$H(a,b)=\frac1\epsilon\begin{bmatrix}
+\diag(a)\diag(\gamma^0 b)&\diag(a)\gamma^0 \diag(b)\\
+\diag(b){^t\gamma^0} \diag(a)&\diag(^t\gamma^0a)\diag(b)
+\end{bmatrix}.$$
+Notice that
+
+$$H(a^*,b^*)=\frac1\epsilon\begin{bmatrix}
+\diag(\mu^1)&\gamma^*\\
+^t\gamma^*&\diag(\mu^2)
+\end{bmatrix}.$$
+
+The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
+Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\gamma^0(diag(^t\gamma^*))^{-1} {^t\gamma^0}$ is positive semi definite.
+Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.
+All the other eigenvalues belongs to the set $[0;1[$.
+
+From experiments, it seems possible to design a good guess for $\omega$ from $c$ and $\epsilon$ only. 
+    Denoting as $\lambda$ the second larger eigenvalue of $T(1,1)$, one can set 
+    $\omega=\frac{2}{\lambda}(1-\sqrt{1-\lambda})$.
 
@@
 {\color{red} TODO}
+
+\section{Conclusion}
+
+Cheaper test, how chosing $\omega$.
 

### 2017-10-13 paper.tex
Pb dual
update

@@
 
+%
 \author{
@@
 	Given lemma \ref{lemma:lyapunov_decrease}, it can be seen that $\gamma^* \in U$. Therefore, iterations eventually all use the over-relaxed operators.
+
 \end{remark}
-{\color{blue} It could be good to show on typical experiments how many times the test fails.
-If the test never fails after a few iterations, we could justify that doing a few Sinkhorn iteration and then always the over-relaxed algorithm is fine in practice.}
+{\color{blue} As the inequality is strict in U, $\gamma^*\notin U$.}
 
@@
 \section{Local rate of convergence ?}
-Some material for adapting Lenaic results:
+Some material to adapt Lenaic results:
 
@@
 
-\section{Conclusion}
+\section{Conclusion and perspectives}
 

### 2017-10-13 paper.tex
Schemas sur_relaxation

--- 2017-10-11 paper.tex
@@
 
-As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  realizes alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
+As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
 The idea developped in this paper is to realize over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c)
@@
 \begin{tabular}{ccc}
-%\includegraphics[width=3cm]{}&\includegraphics[width=3cm]{}&\includegraphics[width=3cm]{}\\
-&TODO: mettre les figures du beamer\\
-&en remplacant les $x$ par des $\gamma$\\
+\includegraphics[height=4cm]{schema_a.png}&\includegraphics[height=4cm]{schema_b.png}&\includegraphics[height=4cm]{schema_c.png}\\
 (a)&(b)&(c)

### 2017-10-13 paper.tex

@@
 Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
-Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} that realizes alternate Bregman projections, can be considered to compute the regularized  transport matrix:
+Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
 \begin{align*}
@@
 As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
-The idea developped in this paper is to realize over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c)
+The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c)
 

### 2017-10-13 paper.tex
update

--- 2017-10-13 paper.tex
@@
 \usepackage{mathtools,enumitem}
-
+\usepackage{algorithm}
+\usepackage{algorithmic}
 %\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
@@
 P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &
-a &=  \frac{\mu^1}{A_1 \gamma} \\
+a &=  {\mu^1}\oslash{A_1 \gamma} \\
 P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &
-b &= \frac{\mu^2}{A_2 \gamma}\nonumber
+b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
 \end{align}
-where the division is considered pointwise. 
+where $\oslash$ is the pointwise division. 
 To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and iterates
 \begin{align*}
-a^{l+1} &= \frac{\mu^1}{\gamma^0 b^l} &
-b^{l+1} &= \frac{\mu^2}{^t \gamma^0 a^{l+1}} 
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} 
 \end{align*}
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:algo}
 	Let $\theta$ and $\omega$ be such that $1\le \theta < \omega < 2$. Let $U$ be the open set defined by:
@@
 \begin{align*}
-a^{l+1} &= \frac{\mu^1}{\gamma^0 b^l} &
-b^{l+1} &= \frac{\mu^2}{^t \gamma^0 a^{l+1}} 
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} 
 \end{align*}
@@
 \section{Experimental results}
+
+In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for realizing the over-relaxed iterations with parameter $\omega$.
+The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
+
+\renewcommand{\algorithmiccomment}[1]{(#1)}
+\begin{algorithm}
+\caption{Over-relaxed Sinkhorn algorithm}
+\label{SOR}
+\begin{algorithmic}
+\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
+\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
+\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
+\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
+\STATE $\bar a=a^{1-\omega}\otimes \tilde a^\omega$
+\STATE $\tilde b=\mu_2\oslash (^t\gamma^0 \bar a)$
+\IF{ $\langle \mu^1,\phi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\phi_\omega(b\oslash\tilde b)\rangle>0$}
+\STATE  $a=\bar a$ and $b=b^{1-\omega}\otimes \tilde b^\omega$
+\COMMENT{Over-relaxed Sinkhorn iteration}
+\ELSE
+\STATE $a=\mu_1\oslash (\gamma^0 b)$ and $b=\mu_2\oslash (^t\gamma^0 a)$  
+\COMMENT{Sinkhorn iteration}
+\ENDIF
+\ENDWHILE
+ \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
+\end{algorithmic}
+\end{algorithm}
+
 {\color{red} TODO}
@@
 \section{Conclusion and perspectives}
+Contributions: 
+cheap test, preliminary work for convergence results. Huge acceleration in practice.
 
-Cheaper test, how chosing $\omega$.
+Perspectives:
+global convergence, how chosing $\omega$?, intensive numerical study with respect to other accelerarions methods\cite{}.
 

### 2017-10-13 paper.tex
Merged

@@
 
-As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  realizes alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
-The idea developped in this paper is to realize over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c)
+As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
+The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c)
 

### 2017-10-13 paper.tex
Algo

@@
 \end{remark}
-{\color{blue} As the inequality is strict in U, $\gamma^*\notin U$.}
+{\color{blue} Comment: As the inequality is strict in U, $\gamma^*\notin U$.}
 
@@
 
-In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for realizing the over-relaxed iterations with parameter $\omega$.
+In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for realizing the over-relaxed iterations with parameter $\omega$.
 The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
 
-\renewcommand{\algorithmiccomment}[1]{(#1)}
+\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
 \begin{algorithm}
@@
 \WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
-\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
-\STATE $\bar a=a^{1-\omega}\otimes \tilde a^\omega$
-\STATE $\tilde b=\mu_2\oslash (^t\gamma^0 \bar a)$
-\IF{ $\langle \mu^1,\phi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\phi_\omega(b\oslash\tilde b)\rangle>0$}
-\STATE  $a=\bar a$ and $b=b^{1-\omega}\otimes \tilde b^\omega$
-\COMMENT{Over-relaxed Sinkhorn iteration}
+\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
+\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
+\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
+\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
 \ELSE
-\STATE $a=\mu_1\oslash (\gamma^0 b)$ and $b=\mu_2\oslash (^t\gamma^0 a)$  
-\COMMENT{Sinkhorn iteration}
+\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
 \ENDIF

### 2017-10-13 paper.tex
Algo

@@
 Perspectives:
-global convergence, how chosing $\omega$?, intensive numerical study with respect to other accelerarions methods\cite{}.
+global convergence, how chosing $\omega$?, intensive numerical study with respect to other accelerarions methods\cite{2016arXiv160604133S,2017arXiv170509634A}.
 

### 2017-10-16 paper.tex
Algo
Proof of lemma:KL_compact

@@
 \begin{lemma} \label{lemma:KL_compact}
-	For any $M \in \IR_+^*$, the sublevel set $\left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
+	For any $M \in \IR_+^*$, the sublevel set $K := \left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
 \end{lemma}
 \begin{proof}
-	{\color{red} TODO}
+	First note that K is closed in $\IR_{+*}^{n_1 n_2}$, by continuity of $F$.
+	For $\gamma \in K$, one has:
+	\begin{align*}
+	F(\gamma) &= \scal{\gamma^*}{\log \gamma^* - 1} + \scal{\gamma}{1} - \scal{\gamma^*}{\log \gamma}\\
+	\sum_{i,j} \gamma_{i,j} - \gamma^*_{i,j} \log \gamma_{i,j} &\le M - \scal{\gamma^*}{\log \gamma^* - 1} .
+	\end{align*}
+	By concavity, the logarithm admits a simple upper bound: $\log \gamma_{i,j} \le \frac{\gamma_{i,j}}{e}$. Moreover, since $\gamma^*$ is a solution of the problem, each of its coordinates is at most 1.
+	\begin{align*}
+	\sum_{i,j} \gamma_{i,j} &\le \frac{M - \scal{\gamma^*}{\log \gamma^*-1}}{1-\frac{1}{e}} =: M'.
+	\end{align*}
+	Therefore each coordinate $\gamma_{i,j}$ is bounded above by $M'$.
+
+	Let us show that the coordinates are bounded below by a strictly positive number; from which it can be deduced that $K$ is bounded and closed in $\IR^{n_1 n_2}$, and therefore compact.
+	\begin{align*}
+	F(\gamma)
+	&\ge \scal{\gamma^*}{\log \gamma^*} - \sum_{i,j} \gamma^*_{i,j} \log(\gamma_{i,j}) - 1\\
+	&\ge \scal{\gamma^*}{\log \gamma^*} - \gamma^*_{i_0,j_0} \log(\gamma_{i_0,j_0}) - 1 - \log M',
+	\end{align*}
+	where the last equation derives from $\log(y_{i,j}) \le \log M'$, and $\sum_{(i,j)\neq(i_0,j_0)} \gamma^*_{i,j} \le 1$.
+	Finally:
+	\[
+	\log(\gamma_{i_0,j_0}) \ge \frac{\scal{\gamma^*}{\log \gamma^*} - M - \log M' - 1}{\gamma^*_{i_0,j_0}},
+	\]
+	and $K$ is thus compact.
 \end{proof}
 
-Note that it is not necessary to know $\gamma^*$ in order to calculate the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$, as shown by the following lemma.
+Note that the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$ may be calculated without knowing $\gamma^*$, as shown by the following lemma.
 \begin{lemma}\label{lemma:lyapunov_decrease}
@@
 In Figure \ref{phi_omega}, we illustrate the functions  $\varphi_\omega$ for several values of $\omega$.
-Notice that for the Sinkhorn algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always non negative.
+Notice that for the Sinkhorn algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative.
+For other values $1\le\omega<2$, it is nonnegative for arguments close to 1.
 \begin{figure}[ht!]
@@
 
-In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for realizing the over-relaxed iterations with parameter $\omega$.
+In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
 The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
@@
 \STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
+\ENDIF
+\ENDWHILE
+ \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Over-relaxed Sinkhorn algorithm 2}
+\label{SOR}
+\begin{algorithmic}
+\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
+\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
+\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
+\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
+\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
+\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
+\ELSE
+\STATE $a=\tilde a$,\COMMENT{Sinkhorn iteration}
+\ENDIF
+\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a)$
+\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
+\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
+\ELSE
+\STATE $b=\tilde b$,\COMMENT{Sinkhorn iteration}
 \ENDIF

### 2017-10-16 paper.tex
Algo

@@
 \end{proof}
-Thus, calculating the decrease in value of the function $F$ is relatively cheap and straightforward.
+Thus,  the decrease in value of the function $F$ for a over-relaxed projection associated to $\Ccal_k$ is cheap to estimate, since its computational cost is  linear with respect to the dimension of data $\mu^k$.
 In Figure \ref{phi_omega}, we illustrate the functions  $\varphi_\omega$ for several values of $\omega$.

### 2017-10-16 paper.tex
Schemas tikz

@@
 
+\usepackage{tikz}
 \usepackage{cite}
@@
 
-The computation of optimal transport between two data relies on the estimation of an optimal transport matrix which entries represent the quantity of mass transported between  data locations. 
+The computation of optimal transport between two data relies on the estimation of an optimal transport matrix, the entries of which represent the quantity of mass transported between  data locations. 
 Regularization of optimal transport with strictly convex regularizers \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
@@
 $$\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_+} \langle c,\gamma\rangle:=\sum_{i,j}c_{i,j}\gamma_{i,j}.$$
-This is a linear programing problem whose resolution is intractable for large scale problems.
+This is a linear programing problem whose resolution is unmanageable for large scale problems.
 
@@
 $\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$ 
-Efficient parallel computations can be considered \cite{cuturi13} and one can even reach real time computation for large scale problem for certain class of cost matrices $c$ \cite{Solomon2015}. 
+Efficient parallel computations can be considered \cite{cuturi13} and one can even reach real-time computation for large scale problem for certain class of cost matrices $c$ \cite{Solomon2015}. 
 For small values of the parameter $\epsilon$, numerical issues can arise and a stabilization of the algorithm is necessary \cite{2016arXiv160705816C}.
@@
 \begin{tabular}{ccc}
-\includegraphics[height=4cm]{schema_a.png}&\includegraphics[height=4cm]{schema_b.png}&\includegraphics[height=4cm]{schema_c.png}\\
+\input{schema_a.tex}&
+\input{schema_b.tex}&
+\input{schema_c.tex}\\
 (a)&(b)&(c)
@@
 It shall be used as a Lyapunov function, so that $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+
+We argue that it makes sense to consider this Lyapunov function.
+Obviously, it is closely related to the notion of Bregman projection that is used throughout the algorithm.
+But it is also easy to calculate its decrease, as shown in lemma \ref{lemma:lyapunov_decrease}.
+Moreover, in practice it allows the use of a wide range of over-relaxation parameters.
 

### 2017-10-16 schema_overrelaxation.py

@@ -1,0 +1,85 @@
+import matplotlib.pyplot as plt
+
+def normalize(x,y):
+    n = (x**2+y**2)**.5
+    return x/n,y/n
+
+"""
+Projects the point (gx,gy) onto the line passing through (0,0) with direction (dx,dy)
+"""
+def project(gx,gy,dx,dy):
+    # Supposes that ||dx,dy|| = 1
+    scal = gx*dx+gy*dy
+    return scal*dx,scal*dy
+
+alpha = 1.7
+def overrelaxed(gx,gy,dx,dy,alpha):
+    gx2,gy2 = project(gx,gy,dx,dy)
+    return alpha*gx2+(1-alpha)*gx, alpha*gy2+(1-alpha)*gy
+
+# gamma^epsilon = 0
+# lines going out from it in directions :
+x1,y1 = normalize(0,1)
+x2,y2 = normalize(-0.4,1.5)
+
+
+ax = plt.axes()
+pict = "\\begin{tikzpicture}\n"
+gx,gy = -0.5,4
+ax.plot([gx],[gy],'o',color=(0.5,0,0.5),label="")
+pict += "\\fill (%f,%f) circle (2pt) node[above] (gamma0) {$\\gamma^0$};\n" % (gx,gy)
+pict += "\\fill (0,0) circle (2pt) node[right] {$\\gamma^*$};\n"
+ax.text(gx,gy,"$\\gamma^0$",fontsize=25)
+xx,yy = [],[]
+pict += "\\draw[dashed] "
+for _ in range(100):
+    xx.append(gx)
+    yy.append(gy)
+    new_gx,new_gy = overrelaxed(gx,gy,x1,y1,alpha)
+    pict += "(%f,%f) -- (%f,%f)\n" % (gx,gy,new_gx,new_gy)
+    xx.append(new_gx)
+    yy.append(new_gy)
+    gx,gy = overrelaxed(new_gx,new_gy,x2,y2,alpha)
+    pict += "(%f,%f) -- (%f,%f)\n" % (gx,gy,new_gx,new_gy)
+pict += ";\n"
+
+def constraint(x1,y1):
+    cx,cy = [-5*x1,10*x1],[-5*y1,10*y1]
+    maxy = max(yy)
+    for i in range(2):
+        if cy[i] > maxy:
+            ratio = 1.1*maxy/cy[i]
+            cy[i] *= ratio
[…]

### 2017-10-17 paper.tex
Algo

@@
 The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
-Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\gamma^0(diag(^t\gamma^*))^{-1} {^t\gamma^0}$ is positive semi definite.
+Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite.
 Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.

### 2017-10-17 paper.tex
Algo

--- 2017-10-16 paper.tex
@@
 The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
-Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\gamma^0(diag(^t\gamma^*))^{-1} {^t\gamma^0}$ is positive semi definite.
+Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite for $a,b>0$.
 Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.

### 2017-10-17 paper.tex
Algo

--- 2017-10-16 paper.tex
@@
 The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
-Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\gamma^0(diag(^t\gamma^*))^{-1} {^t\gamma^0}$ is positive semi definite.
+Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite for $a_i,b_j>0$.
 Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.

### 2017-10-17 paper.tex
Theoreme continu

@@
 \begin{theorem}\label{thm:algo}
-	Let $\theta$ and $\omega$ be such that $1\le \theta < \omega < 2$. Let $U$ be the open set defined by:
-	\begin{equation}\label{eq:open_set_U}
-	U = \left\{
-	\gamma \in \IR_{+*}^{nm} \mid
-	F(P^\omega_{\Ccal_1}(\gamma)) < F(\gamma)
-	\text{ and }
-	F(P^\omega_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma))) < F(P^\theta_{\Ccal_1}(\gamma))
-	\right\}.
-	\end{equation}
-	Set the initial value $\gamma^0 = e^{-c/\epsilon}$, and iterate:
+	Take two continuous functions $\theta_1,\theta_2$ such that
+	\[
+	\forall \gamma \in \IR_{+*}^{n_1 n_2},\quad
+	F(P_{\Ccal_k}^{\theta_k(\gamma)}(\gamma)) \le F(\gamma) ,
+	\]
+	where the inequality is strict whenever $\gamma \notin \Ccal_k$.
+	Consider the sequence defined by:
 	\begin{align*}
-	\gamma^{l+1} =
-	\begin{cases}
-	P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l)) & \text{if } \gamma \in U \\
-	P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l)) & \text{otherwise.}
-	\end{cases}
+	\gamma^0 &= e^{-c/\epsilon} \\
+	\tilde{\gamma}^{l+1} &= P_{\Ccal_1}^{\theta_1(\gamma^l)}(\gamma^l) \\
+	\gamma^{l+1} &= P_{\Ccal_2}^{\theta_2(\tilde{\gamma}^{l+1})}(\tilde{\gamma}^{l+1}).
 	\end{align*}
-	Then the iterates $(\gamma^l)$ converge to $\gamma^*$.
+	Then the sequence $(\gamma^l)$ converges to $\gamma^*$.
 \end{theorem}
 {\color{red} 
-	In practice the test would be: check whether the $\omega$-relaxed projection of $\zeta$ on $\Ccal_1$ decreases the value of
-	$F$; if so, calculate $P_{\Ccal_1}^\theta(\zeta)$ (don't update $\zeta$ yet) and check that its $\omega$-relaxed projection on $\Ccal_2$ decreases the value of $F$ again. \\
-	$\longrightarrow$ Has to calculate 3 projections sometimes.
-This result is weaker than the one from my report, but easier to prove.
-Maybe we could avoid having $U$, and rather perform the test independently on each variable ?
-(Or we could find a cheaper test).}
+TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
 \begin{remark}
@@
 
-	The set $U$ is constructed in such a way that, by applying lemma \ref{lemma:F_P_theta} several times, one ensures that the sequence $(F(\gamma^l))$ is nonincreasing.
-Moreover, it is strictly decreasing as long as $\gamma^l \not \in \Ccal_1 \cap \Ccal_2$.
-	Hence, 
-	\[
-	F(\gamma^l) \le F(\gamma^0).
-	\]
-	By lemma \ref{lemma:KL_compact}, the sequence $(\gamma^l)$ is thus precompact.
-Let us show that any accumulation point $\zeta$ belongs to $\Ccal_1 \cap \Ccal_2$.
-Since $S$ is closed in $\IR_{+*}^{n_1 n_2}$, by lemma \ref{lemma:trivial_intersection} this would mean that $\zeta = \gamma^*$, and would be sufficient to prove the convergence.
+	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^l))$ is non-increasing. Hence $(\gamma^l)$ is precompact.
 
-	Let $(\gamma^{k_l})$ be a subsequence of $(\gamma^l)$ that converges to $\zeta$.
-	Let us denote, for convenience, $T_1 = P_{\Ccal_2}^\theta \circ P_{\Ccal_1}^\theta$ and $T_2 = P_{\Ccal_2} \circ P_{\Ccal_1}$ the two operators used in the iteration, and $T$ the compound operator, so that $T_{|U} \equiv T_{1|U}$ and $T_{|U^c} \equiv T_{2|U^c}$.
-	\begin{enumerate}[align=right,labelsep=10pt,leftmargin=0.5cm]
-		\item[$\bullet$] If $\zeta \in U$, then since $U$ is open, all $(\gamma^{k_l})$ are eventually in $U$ as well.
-		Thus, for any large enough $l$, $ T(\gamma^{k_l}) = T_1(\gamma^{k_l}) $. By continuity of $T_1$ on $U$, since $(\gamma^{k_l})$ converges to $\zeta$, 
-		\[T(\zeta) = \lim_{l\rightarrow \infty} T(\gamma^{k_l}).\] Thus
-		 \[F(T(\zeta)) = \lim_{l \rightarrow \infty} F(T(\gamma^{k_l})) \ge \lim_{l\rightarrow \infty} F(\gamma^{k_{l+1}}) .\]
-		 \[
-		 F(T(\zeta)) \ge F(\zeta)
-		 \] Since $\zeta$ is inside the set $U$, which ensures strict decrease of the Lyapunov function, this is absurd.
-		\item[$\bullet$] If $\zeta \in U^c$ and an infinite number of $(\gamma^{k_l})$ are in $U^c$, then the same argument of continuity applied to $T_2$ yields
-		 \[
-		 F(T(\zeta)) = F(\zeta).
-		 \]
-		 The function $\varphi_1$ from lemma \ref{lemma:lyapunov_decrease} is strictly positive except at $x=1$.
-Since the case of equality occurs for both projections, this ensures that $\zeta \in \Ccal_1 \cap \Ccal_2$.
-
-		 \item[$\bullet$] If $\zeta \in U^c$ and all but finitely many $\gamma^{k_l}$ are in $U$, then $\zeta$ is on the border $\partial U$. In particular it belongs to $\bar{U}$, and therefore
-		 \[
-		 F(P^\omega_{\Ccal_1}(\zeta)) \le F(\zeta) \quad
-		 \text{ and } \quad
-		 F(P^\omega_{\Ccal_2}(P^\theta_{\Ccal_1}(\zeta))) \le F(P^\theta_{\Ccal_1}(\zeta)).
-		 \]
-		 By lemma \ref{lemma:F_P_theta}, the inequality \begin{equation} \label{eq:ineq1}
-		 F(P_{\Ccal_1}^\theta(\zeta)) \le F(P_{\Ccal_1}^\omega(\zeta)) \end{equation}
-		 is valid; it is strict unless $\zeta \in \Ccal_1$.
-		 The same lemma yields
-		 \begin{equation} \label{eq:ineq2}
-		 F(P_{\Ccal_2}^\theta(P_{\Ccal_1}^\theta(\zeta)) \le F(P_{\Ccal_2}^\omega(P_{\Ccal_1}^\theta(\zeta)), \end{equation}
-		 with a strict inequality unless $P_{\Ccal_1}^\theta(\zeta) \in \Ccal_2$.
-		 To sum up, we have the series of inequalities:
-		 \[
-		 F(\zeta) 
-		 \overset{\zeta \in \bar{U}}{\ge}
-		 F(P^\omega_{\Ccal_1}(\zeta))
-		 \overset{\text{(\ref{eq:ineq1})}}{\ge}
-		 F(P^\theta_{\Ccal_1}(\zeta))
-		 \overset{\zeta \in U}{\ge}
-		 F(P_{\Ccal_2}^\omega(P_{\Ccal_1}^\theta(\zeta)))
-		 \overset{\text{(\ref{eq:ineq2})}}{\ge}
-		 F(P_{\Ccal_2}^\theta(P_{\Ccal_1}^\theta(\zeta)))
-		 \]
-
-		 Yet, by continuity of $T_1 = P_{\Ccal_2}^\theta \circ P_{\Ccal_1}^\theta $ on $\bar{U}$, we have
-		 \[
-		 F(P_{\Ccal_2}^\theta(P_{\Ccal_1}^\theta(\zeta))) = F(\zeta).
-		 \]
-		 Thus both (\ref{eq:ineq1}) and (\ref{eq:ineq2}) fall in the case of equality, which means $\zeta \in \Ccal_1$ and $P^\theta_{\Ccal_1}(\zeta) = \zeta \in \Ccal_2$.
-	\end{enumerate}
+	If $\xi$ is a cluster point of $(\gamma^l)$, let us define
+	\begin{align*}
+	\tilde{\xi} &= P_{\Ccal_1}^{\theta_1(\xi)}(\xi) \\
+	\xi' &= P_{\Ccal_2}^{\theta_2(\tilde{\xi})}(\tilde{\xi}).
+	\end{align*}
+	Then by continuity of all the applications, $F(\xi) = F(\tilde{\xi}) = F(\xi')$.
+	From the hypothesis made on $\theta_1$ and $\theta_2$ can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
+Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
+	By lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
 \end{proof}

### 2017-10-17 paper.tex
Proof of the continuous theorem

@@
 TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
-\begin{remark}
-	Given lemma \ref{lemma:lyapunov_decrease}, it can be seen that $\gamma^* \in U$. Therefore, iterations eventually all use the over-relaxed operators.
-
-\end{remark}
-{\color{blue} Comment: As the inequality is strict in U, $\gamma^*\notin U$.}
 
@@
 \end{lemma}
+
+\begin{proof}[Proof of the theorem]
+	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^l)$ are thus diagonally similar to $\gamma^0$:
+	\[
+	\forall l\ge0,\quad \gamma^l \in S
+	\]
+
+	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^l))$ is non-increasing. Hence $(\gamma^l)$ is precompact.
+
+	If $\xi$ is a cluster point of $(\gamma^l)$, let us define
+	\begin{align*}
+	\tilde{\xi} &= P_{\Ccal_1}^{\theta_1(\xi)}(\xi) \\
+	\xi' &= P_{\Ccal_2}^{\theta_2(\tilde{\xi})}(\tilde{\xi}).
+	\end{align*}
+	Then by continuity of all the applications, $F(\xi) = F(\tilde{\xi}) = F(\xi')$.
+	From the hypothesis made on $\theta_1$ and $\theta_2$ can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
+Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
+	By lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
+\end{proof}
+
+To construct the functions $\theta_k$ needed by the theorem, we
 
@@
 	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_1 \gamma = \mu$.
-\end{proof}
-
-\begin{proof}[Proof of the theorem]
-	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^l)$ are thus diagonally similar to $\gamma^0$:
-	\[
-	\forall l\ge0,\quad \gamma^l \in S
-	\]
-
-	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^l))$ is non-increasing. Hence $(\gamma^l)$ is precompact.
-
-	If $\xi$ is a cluster point of $(\gamma^l)$, let us define
-	\begin{align*}
-	\tilde{\xi} &= P_{\Ccal_1}^{\theta_1(\xi)}(\xi) \\
-	\xi' &= P_{\Ccal_2}^{\theta_2(\tilde{\xi})}(\tilde{\xi}).
-	\end{align*}
-	Then by continuity of all the applications, $F(\xi) = F(\tilde{\xi}) = F(\xi')$.
-	From the hypothesis made on $\theta_1$ and $\theta_2$ can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
-Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
-	By lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
 \end{proof}

### 2017-10-17 paper.tex
Revert to test

--- 2017-10-17 paper.tex
@@
 \begin{remark}
-	Given lemma \ref{lemma:lyapunov_decrease}, it can be seen that $\gamma^* \in U$. Therefore, iterations eventually all use the over-relaxed operators.
-
+	Given lemma \ref{lemma:lyapunov_decrease}, it can be seen that a neighborhood of $\gamma^*$ is in $U$. Therefore, iterations eventually all use the over-relaxed operators.
 \end{remark}
-{\color{blue} Comment: As the inequality is strict in U, $\gamma^*\notin U$.}
 

### 2017-10-18 paper.tex
Adaptative theta

--- 2017-10-17 paper.tex
@@
 As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
-The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c)
+The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
 
@@
 	Take two continuous functions $\theta_1,\theta_2$ such that
-	\[
+	\begin{equation}\label{eq:cond_theta_k}
 	\forall \gamma \in \IR_{+*}^{n_1 n_2},\quad
 	F(P_{\Ccal_k}^{\theta_k(\gamma)}(\gamma)) \le F(\gamma) ,
-	\]
+	\end{equation}
 	where the inequality is strict whenever $\gamma \notin \Ccal_k$.
@@
 \end{theorem}
-{\color{red} 
-TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
 
@@
 	Then by continuity of all the applications, $F(\xi) = F(\tilde{\xi}) = F(\xi')$.
-	From the hypothesis made on $\theta_1$ and $\theta_2$ can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
+	From the hypothesis made on $\theta_1$ and $\theta_2$, can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
 Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
@@
 
-To construct the functions $\theta_k$ needed by the theorem, we
+{\color{red} 
+	TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
+To construct the functions $\theta_k$ needed by the theorem, we need the following lemma.
 
@@
 	\]
-	Moreover, $t \in [1,\infty) \mapsto \varphi_t(x)$ is nonincreasing:
-	\[
-	\frac{d}{dt} \varphi_t(x) = \log x (x^{1-t} - 1).
-	\]
+	Moreover, the function that maps $t \in [1,\infty)$ to $\varphi_t(x)$ is non-increasing:
+	$\frac{d}{dt} \varphi_t(x) = \log x (x^{1-t} - 1).$
 	For $x\neq 1$, it is even strictly decreasing.
-	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_1 \gamma = \mu$.
+	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_k \gamma = \mu$.
+\end{proof}
+
+\begin{proposition}
+	We argue that a good choice for the functions $\theta_k$ may be constructed as follows. Pick a target parameter $\theta_0 \in [1;2)$, and a small $\delta>0$.
+	Define the function $\theta_k^*$ as:
+	\begin{equation}
+	\theta_k^*(\gamma) = \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min \frac{A_k \gamma}{\mu^k}\right) \ge 0 \right\} .
+	\end{equation}
+	Then the function
+	\begin{equation}
+	\label{eq:theta_k}
+	\theta_k(\gamma) = \min(\max(1,\theta_k^*(\gamma)-\delta),\theta_0)
+	\end{equation}
+	is continuous, and verifies condition (\ref{eq:cond_theta_k}).
+\end{proposition}
+This construction has several advantages.
+One is that it allows to choose arbitrarily the parameter $\theta_0$ that will be used asymptotically. A heuristic for this choice is proposed in section \ref{section:local}.
+But it is also an easy approach to having an adaptive method, as the approximation of $\theta_k^*$ has a negligible cost (it only depends on the smallest value of $\frac{A_k \gamma}{\mu^k}$, and the function to invert is decreasing and concave).
+We will show later on that this is often an excellent choice in practice.
+\begin{proof}
+	Since the partial derivative of $\frac{\partial\varphi_\omega(x)}{\partial\omega}$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\theta_k^*$.
+	The function $\theta_k^*(\gamma)$ is constructed so that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
+	Therefore, by lemma \ref{lemma:F_P_theta}, using parameter $\theta_k^*(\gamma)-\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
+	It is no problem to constrain it to $[1,\theta_0]$, as using 1 always validates the inequality (\ref{eq:cond_theta_k}), and using $\theta_0$ when it is smaller than $\theta_k^*(\gamma)$ does as well.
 \end{proof}
@@
 \section{Local rate of convergence ?}
+\label{section:local}
 Some material to adapt Lenaic results:

### 2017-10-18 paper.tex
Algo

@@
 
-{\color{red} 
-	TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
+%{\color{red} 
+%	TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
 To construct the functions $\theta_k$ needed by the theorem, we need the following lemma.
@@
 
-In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
-The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
+%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
+%The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
 
 \renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
+%\begin{algorithm}
+%\caption{Over-relaxed Sinkhorn algorithm}
+%\label{SOR}
+%\begin{algorithmic}
+%\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
+%\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
+%\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
+%\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
+%\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
+%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
+%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
+%\ELSE
+%\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
+%\ENDIF
+%\ENDWHILE
+% \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
+%\end{algorithmic}
+%\end{algorithm}
+In order to summarize the whole algorithm, we first reformulate the functions $\theta_k(\gamma)$ defined in \eqref{eq:theta_k} as $\theta_k(\gamma)=\Theta(\frac{A_k \gamma}{\mu^k})$ where 
+$ \Theta(u) = \min(\max(1,\Theta^*(u)-\delta),\theta_0)$ and  $\Theta^*(u)=\sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u \right) \ge 0 \right\}$.
+
 \begin{algorithm}
-\caption{Over-relaxed Sinkhorn algorithm}
+\caption{Over-relaxed Sinkhorn algorithm }
 \label{SOR}
@@
 \REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
-\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
-\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
-\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
-\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
-\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
-\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
-\ELSE
-\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
-\ENDIF
-\ENDWHILE
- \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{Over-relaxed Sinkhorn algorithm 2}
-\label{SOR}
-\begin{algorithmic}
-\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
-\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
+\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\theta_0\in[1;2)$ and $\eta>0$
 \WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
 \STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
-\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
-\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
-\ELSE
-\STATE $a=\tilde a$,\COMMENT{Sinkhorn iteration}
-\ENDIF
+\STATE $\omega=\Theta(a\oslash\tilde a)$
+%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
+\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$%\COMMENT{Over-relaxed Sinkhorn iteration}
+%\ELSE
+%\STATE $a=\tilde a$,\COMMENT{Sinkhorn iteration}
+%\ENDIF
 \STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a)$
-\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
-\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
-\ELSE
-\STATE $b=\tilde b$,\COMMENT{Sinkhorn iteration}
-\ENDIF
+\STATE $\omega=\Theta(b\oslash\tilde b)$
+%\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
+\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Over-relaxed Sinkhorn iteration}
+%\ELSE
+%\STATE $b=\tilde b$,\COMMENT{Sinkhorn iteration}
+%\ENDIF
 \ENDWHILE

### 2017-10-18 paper.tex
Algo

@@
 One is that it allows to choose arbitrarily the parameter $\theta_0$ that will be used asymptotically. A heuristic for this choice is proposed in section \ref{section:local}.
-But it is also an easy approach to having an adaptive method, as the approximation of $\theta_k^*$ has a negligible cost (it only depends on the smallest value of $\frac{A_k \gamma}{\mu^k}$, and the function to invert is decreasing and concave).
+But it is also an easy approach to having an adaptive method, as the approximation of $\theta_k^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $\frac{A_k \gamma}{\mu^k}$, and the corresponding function to invert is decreasing and concave).
 We will show later on that this is often an excellent choice in practice.

### 2017-10-18 paper.tex
Notation Theta

@@
 	We argue that a good choice for the functions $\theta_k$ may be constructed as follows. Pick a target parameter $\theta_0 \in [1;2)$, and a small $\delta>0$.
-	Define the function $\theta_k^*$ as:
-	\begin{equation}
-	\theta_k^*(\gamma) = \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min \frac{A_k \gamma}{\mu^k}\right) \ge 0 \right\} .
-	\end{equation}
+	Define the functions $\Theta^*$ and $\Theta$ as:
+	\begin{align}
+	\Theta^*(u) &= \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u\right) \ge 0 \right\} ,\\
+	\label{eq:Theta}
+	\Theta(u) &= \min(\max(1,\Theta^*(u)-\delta),\theta_0),
+	\end{align}
+	where $\min u$ denotes the smallest coordinate of the vector $u$.
 	Then the function
@@
 	\label{eq:theta_k}
-	\theta_k(\gamma) = \min(\max(1,\theta_k^*(\gamma)-\delta),\theta_0)
+	\theta_k(\gamma) =\Theta\left (\frac{A_k \gamma}{\mu^k}\right)
 	\end{equation}
@@
 \begin{proof}
-	Since the partial derivative of $\frac{\partial\varphi_\omega(x)}{\partial\omega}$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\theta_k^*$.
-	The function $\theta_k^*(\gamma)$ is constructed so that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
-	Therefore, by lemma \ref{lemma:F_P_theta}, using parameter $\theta_k^*(\gamma)-\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
-	It is no problem to constrain it to $[1,\theta_0]$, as using 1 always validates the inequality (\ref{eq:cond_theta_k}), and using $\theta_0$ when it is smaller than $\theta_k^*(\gamma)$ does as well.
+	Since the partial derivative of $\frac{\partial\varphi_\omega(x)}{\partial\omega}$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
+	The function $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
+	Therefore, by lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
+	It is no problem to constrain it to $[1,\theta_0]$, as using 1 always verifies the inequality (\ref{eq:cond_theta_k}), and using $\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
 \end{proof}
@@
 %\end{algorithm}
-In order to summarize the whole algorithm, we first reformulate the functions $\theta_k(\gamma)$ defined in \eqref{eq:theta_k} as $\theta_k(\gamma)=\Theta(\frac{A_k \gamma}{\mu^k})$ where 
-$ \Theta(u) = \min(\max(1,\Theta^*(u)-\delta),\theta_0)$ and  $\Theta^*(u)=\sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u \right) \ge 0 \right\}$.
+
+This algorithm uses the function $\Theta$ defined in (\ref{eq:Theta}).
 

### 2017-10-18 paper.tex
Abstract

@@
 In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
+In this paper, we propose a simple method for establishing the global convergence of the over-relaxed Sinkhorn's algorithm, by ensuring the decrease of a Lyapunov function at each step.
+An inexpensive adaptive choice of over-relaxation parameter based on the Lyapunov function is constructed.
+We also suggest a heuristic to choose a suitable asymptotic over-relaxation parameter.
 \end{abstract}
@@
 where $\oslash$ is the pointwise division. 
-To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and iterates
+To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and to iterate
 \begin{align*}

### 2017-10-18 paper.tex
Sinkhorn -> Sinkhorn--Knopp

@@
 
-\title{Over-relaxed Sinkhorn's Algorithm for Regularized Optimal Transport}
+\title{Over-relaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport}
 
@@
 This article describes a method for quickly calculating the solution to the regularized optimal transport problem.
-It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn's algorithm).
+It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn--Knopp algorithm).
 The idea is to over-relax the Bregman projection operators, allowing for faster convergence.
 In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
-In this paper, we propose a simple method for establishing the global convergence of the over-relaxed Sinkhorn's algorithm, by ensuring the decrease of a Lyapunov function at each step.
+In this paper, we propose a simple method for establishing the global convergence of the over-relaxed Sinkhorn--Knopp algorithm, by ensuring the decrease of a Lyapunov function at each step.
 An inexpensive adaptive choice of over-relaxation parameter based on the Lyapunov function is constructed.
@@
 Optimal Transport is an efficient and flexible tool to compare two probability distributions which has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
-The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn's algorithm \cite{sinkhorn64}   scaling with high dimensional data.
+The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64}   scaling with high dimensional data.
 Regularized optimal transport have thus been intensively used  in  Machine Learning with applications such as   Geodesic PCA \cite{seguy2015principal}, domain adaptation \cite{2015arXiv150700504C}, data fitting \cite{2015arXiv150605439F},  training of Boltzmann Machine \cite{NIPS2016_6248}  or dictionary learning \cite{Rolet2016,2017arXiv170801955S}.
@@
 
-\subsection{Sinkhorn's algorithm}
+\subsection{The Sinkhorn--Knopp algorithm}
 Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
-Hence, the so-called Sinkhorn algorithm \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
+Hence, the so-called Sinkhorn--Knopp algorithm \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
 \begin{align*}
@@
 The introduction of relaxation variables through heavy ball approaches \cite{POLYAK19641} has recently gained in popularity  to speed up the convergence of algorithms optimizing convex \cite{2014arXiv1412.7457G} or non convex \cite{Zavriev1993,2016arXiv160609070O} problems.
-Such schemes have also been empirically considered to accelerate Sinkhorn's algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
+Such schemes have also been empirically considered to accelerate the SK algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
 The convergence of these algorithms is nevertheless not studied yet in the context of regularized optimal transport.
@@
 \subsection{Overview and contributions}
-In this paper, we consider an over-relaxation scheme designed to accelerate   Sinkhorn's algorithm.
+In this paper, we consider an over-relaxation scheme designed to accelerate the Sinkhorn--Knopp algorithm.
 We first present and show the convergence of our algorithm in section 2.
@@
 
-\section{Over-relaxed Sinkhorn's algorithm}
+\section{Over-relaxed Sinkhorn--Knopp algorithm}
 
-As illustrated in Figure \ref{alternate_projections} (a-b), the Sinkhorn algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
+As illustrated in Figure \ref{alternate_projections} (a-b), the SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
 The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
@@
 \end{tabular}
-\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the Sinkhorn algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
+\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
 In order to accelerate the convergence rate, over-relaxed projections are here considered (c).}
@@
 where the logarithm is taken coordinate-wise.
-Note that $P_{\Ccal_k}^0 = Id$, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ gives the Sinkhorn algorithm, and $P_{\Ccal_k}^2$ is an involution (this is non-trivial, and uses the fact that $\Ccal_k$ is an affine subspace).
+Note that $P_{\Ccal_k}^0$ is the identity, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ is the standard Bregman projection, and $P_{\Ccal_k}^2$ is an involution (this uses the fact that $\Ccal_k$ is an affine subspace).
 
@@
 In Figure \ref{phi_omega}, we illustrate the functions  $\varphi_\omega$ for several values of $\omega$.
-Notice that for the Sinkhorn algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative.
-For other values $1\le\omega<2$, it is nonnegative for arguments close to 1.
+Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative. For other values $1\le\omega<2$, it is nonnegative for arguments close to 1.
 \begin{figure}[ht!]
@@
 
-The Sinkhorn algorithm can be interpreted as an alternate maximisation algorithm on the dual of the regularized optimal transport problem.
+The SK algorithm can be interpreted as an alternate maximisation algorithm on the dual of the regularized optimal transport problem.
 The dual problem of \eqref{ROT} is:
@@
 %\begin{algorithm}
-%\caption{Over-relaxed Sinkhorn algorithm}
+%\caption{Over-relaxed SK algorithm}
 %\label{SOR}
@@
 %\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
-%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
+%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed SK iteration}
 %\ELSE
@@
 \begin{algorithm}
-\caption{Over-relaxed Sinkhorn algorithm }
+\caption{Over-relaxed SK algorithm }
 \label{SOR}
@@
 %\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
-\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$%\COMMENT{Over-relaxed Sinkhorn iteration}
+\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$%\COMMENT{Over-relaxed SK iteration}
 %\ELSE
-%\STATE $a=\tilde a$,\COMMENT{Sinkhorn iteration}
+%\STATE $a=\tilde a$,\COMMENT{SK iteration}
 %\ENDIF
@@
 %\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
-\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Over-relaxed Sinkhorn iteration}
+\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Over-relaxed SK iteration}
 %\ELSE
-%\STATE $b=\tilde b$,\COMMENT{Sinkhorn iteration}
+%\STATE $b=\tilde b$,\COMMENT{SK iteration}
 %\ENDIF

### 2017-10-19 paper.tex
Algo
Details

@@
 Hence, for particular purposes such as color interpolation \cite{Rabin2014} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
-In this setting,  the regularized transport problem can be ill-conditioned. This is the issue  we want to tackle here.
+In this setting,  the regularized transport problem can be ill-conditioned, rendering the Sinkhorn--Knopp algorithm significantly slower. This is the issue  we want to tackle here.
 Before going into further details, we now briefly introduce the main notations and concepts used all along this article.
@@
 where $P_{\Ccal}$ is the  Bregman projection onto $\Ccal$ defined as:
-\[
+\begin{equation}\label{eq:P_C}
 P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} KL(\gamma,\xi).
-\]
+\end{equation}
 
@@
 
-In the discrete setting,  projections actually correspond to diagonal scalings.
+In the discrete setting,  projections actually correspond to diagonal scalings. This can be seen by writing the minimality conditions for the projection, given equations (\ref{KL}) and (\ref{eq:P_C}).
 \begin{align}\label{scaling}
@@
 
-As illustrated in Figure \ref{alternate_projections} (a-b), the SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
+As illustrated in Figure \ref{alternate_projections} (a-b), the SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$, as the problem becomes ill-conditioned.
 The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
@@
 where the logarithm is taken coordinate-wise.
-Note that $P_{\Ccal_k}^0$ is the identity, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ is the standard Bregman projection, and $P_{\Ccal_k}^2$ is an involution (this uses the fact that $\Ccal_k$ is an affine subspace).
+Note that $P_{\Ccal_k}^0$ is the identity, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ is the standard Bregman projection, and $P_{\Ccal_k}^2$ is an involution (recall that $\Ccal_k$ is an affine subspace).
 
@@
 But it is also easy to calculate its decrease, as shown in lemma \ref{lemma:lyapunov_decrease}.
-Moreover, in practice it allows the use of a wide range of over-relaxation parameters.
+Moreover, as we will see later, it allows the use of a wide range of over-relaxation parameters in practice.
 
@@
 	Therefore each coordinate $\gamma_{i,j}$ is bounded above by $M'$.
-
-	Let us show that the coordinates are bounded below by a strictly positive number; from which it can be deduced that $K$ is bounded and closed in $\IR^{n_1 n_2}$, and therefore compact.
+	Let us show that the coordinates are bounded below by a strictly positive number; from which can be deduced that $K$ is bounded and closed in $\IR^{n_1 n_2}$, and therefore compact.
 	\begin{align*}
@@
 	\end{align*}
-	where the last equation derives from $\log(y_{i,j}) \le \log M'$, and $\sum_{(i,j)\neq(i_0,j_0)} \gamma^*_{i,j} \le 1$.
+	where the last equation derives from $\log(\gamma_{i,j}) \le \log M'$, and $\sum_{(i,j)\neq(i_0,j_0)} \gamma^*_{i,j} \le 1$.
 	Finally:
@@
 \begin{lemma}\label{lemma:lyapunov_decrease}
-	Take $\gamma$ in $\IR^{mn}_{+*}$. The decrease in value of the Lyapunov function can be calculated with the following formula:
+	Take $\gamma$ in $\IR^{mn}_{+*}$. The decrease of the Lyapunov function when applying an over-relaxed projection can be calculated with the following formula:
 	\begin{equation} \label{eq:kl_diff_scal}
@@
 \end{proof}
-Thus,  the decrease in value of the function $F$ for a over-relaxed projection associated to $\Ccal_k$ is cheap to estimate, since its computational cost is  linear with respect to the dimension of data $\mu^k$.
+Thus, the decrease of the Lyapunov function $F$ for an over-relaxed projection associated to $\Ccal_k$ is cheap to estimate, since its computational cost is  linear with respect to the dimension of data $\mu^k$.
 In Figure \ref{phi_omega}, we illustrate the functions  $\varphi_\omega$ for several values of $\omega$.
-Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative. For other values $1\le\omega<2$, it is nonnegative for arguments close to 1.
+Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always non-negative. For other values $1\le\omega<2$, it is non-negative for arguments close to 1.
 \begin{figure}[ht!]
@@
 
-\begin{lemma}[\cite{cuturi13}]
+\begin{lemma}[]
 	\label{lemma:trivial_intersection}
@@
 \end{lemma}
+We refer to \cite{cuturi13} for the proof of this well-known result, which is used in the proof of the theorem.
 
@@
 	For $x\neq 1$, it is even strictly decreasing.
-	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_k \gamma = \mu$.
+	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_k \gamma = \mu^k$.
 \end{proof}
@@
 	\begin{align}
+	\label{eq:Theta_opt}
 	\Theta^*(u) &= \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u\right) \ge 0 \right\} ,\\
@@
 %\end{algorithm}
+Add algorithm to solve \eqref{eq:Theta_opt}?
 
@@
 \section{Conclusion and perspectives}
+
+We have presented a generalization of the Sinkhorn--Knopp algorithm based on over-relaxed projections, with a criterion for global convergence based on a Lyapunov function.
+The proposed Lyapunov function allows for an adaptive choice of over-relaxation parameter at a low computational price.
+In practice, this algorithm converges way faster than Sinkhorn--Knopp; moreover, the acceleration rate is especially impressive for small regularization $\epsilon$.
+
 Contributions: 
@@
 global convergence, how chosing $\omega$?, intensive numerical study with respect to other accelerarions methods\cite{2016arXiv160604133S,2017arXiv170509634A}.
+
+Links with monotone Fista? (that choose Fista if energy decreased and forward backward otherwise...)
 

### 2017-10-19 backup_191017/paper.tex
Len first contrib

--- 2017-10-18 paper.tex
@@
 \DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\KL}{KL}
 \newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
@@
 Hence, for particular purposes such as color interpolation \cite{Rabin2014} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
-In this setting,  the regularized transport problem can be ill-conditioned. This is the issue  we want to tackle here.
+In this setting,  the regularized transport problem can be ill-conditioned and Sinkhorn's algorithm converges slowly. This is the issue  we want to tackle here.
 Before going into further details, we now briefly introduce the main notations and concepts used all along this article.
@@
 \end{align*}
-as well as the linear  constraint sets
+as well as the affine constraint sets
 \begin{align*}
@@
 $$\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_+} \langle c,\gamma\rangle:=\sum_{i,j}c_{i,j}\gamma_{i,j}.$$
-This is a linear programing problem whose resolution is unmanageable for large scale problems.
+This is a linear programming problem whose resolution becomes intractable for large problems.
 
@@
 \begin{equation}\label{ROT}
-\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_{+*}}K^\epsilon(\gamma) := \scal{c}{\gamma} 
-+ \epsilon KL(\gamma,\mathbf{1})
+\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_{+}}K^\epsilon(\gamma) := \scal{c}{\gamma} 
++ \epsilon \KL(\gamma,\mathbf{1})
 ,\end{equation}
-with $\epsilon>0$, $\mathbf{1}$ is the matrix of size $n_1\times n_2$ full of ones and the Kullback-Leibler divergence is:
+with $\epsilon>0$, $\mathbf{1}$ is the matrix of size $n_1\times n_2$ full of ones and the Kullback-Leibler divergence is
 \begin{equation}\label{KL}
-KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
+\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}
 \end{equation}
-It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$:
+with the convention $0 \log 0= 0$.
+It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ (here and in the sequel, exponentiation is meant entrywise) onto $\Ccal_1 \cap \Ccal_2$:
 \begin{equation}\label{eq:reg_ot_pb}
@@
 \end{equation}
-where $P_{\Ccal}$ is the  Bregman projection onto $\Ccal$ defined as:
+where $P_{\Ccal}$ is the  Bregman projection onto $\Ccal$ defined as
 \[
-P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} KL(\gamma,\xi).
+P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi).
 \]
@@
 Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
-Hence, the so-called Sinkhorn--Knopp algorithm \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
+Hence, the so-called Sinkhorn--Knopp algorithm (SK) \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
 \begin{align*}
@@
 $\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
-
-In the discrete setting,  projections actually correspond to diagonal scalings.
+%
+In the discrete setting, these projections correspond to diagonal scalings of the input:
 \begin{align}\label{scaling}
-P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &
+P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
 a &=  {\mu^1}\oslash{A_1 \gamma} \\
-P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &
+P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
 b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
@@
 where $\oslash$ is the pointwise division. 
-To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and to iterate
+To compute numerically the solution one simply has to store $(a^l, b^l)\in\IR^{n_1}\times \IR^{n_2}$ and to iterate
+\begin{align*}
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} .
+\end{align*}
+We then have $\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$
+
+Another way to interpret the SK algorithm is as an alternate maximization algorithm on the dual of the regularized optimal transport problem. The dual problem of \eqref{ROT} is:
+\begin{equation}\label{DROT}
+\max_{\alpha,\beta}\; \langle \alpha,\mu^1\rangle+\langle \beta,\mu^2\rangle-\epsilon\sum_{i,j}e^{(\alpha_i+\beta_j-c_{i,j})/\epsilon},
+\end{equation}
+it is concave and continuously differentiable.
+Alternate maximization then converges and we recover, for $a_i=e^{\alpha_i/\epsilon}$, $b_j=e^{\beta_j/\epsilon}$ and $\gamma^0_{i,j}=e^{-c_{i,j}/\epsilon}$:
 \begin{align*}
@@
 \end{align*}
-where the transport matrix can be recovered as :
-$\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$ 
+
 Efficient parallel computations can be considered \cite{cuturi13} and one can even reach real-time computation for large scale problem for certain class of cost matrices $c$ \cite{Solomon2015}. 
@@
 The convergence of the process can nevertheless be very slow  in the setting $\epsilon\to 0$.
-\subsection{Related works }
-The introduction of relaxation variables through heavy ball approaches \cite{POLYAK19641} has recently gained in popularity  to speed up the convergence of algorithms optimizing convex \cite{2014arXiv1412.7457G} or non convex \cite{Zavriev1993,2016arXiv160609070O} problems.
-Such schemes have also been empirically considered to accelerate the SK algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
-The convergence of these algorithms is nevertheless not studied yet in the context of regularized optimal transport.
 
@@
 
-\section{Over-relaxed Sinkhorn--Knopp algorithm}
+\subsection{Related works }
+The introduction of relaxation variables through heavy ball approaches \cite{POLYAK19641} has recently gained in popularity  to speed up the convergence of algorithms optimizing convex \cite{2014arXiv1412.7457G} or non convex \cite{Zavriev1993,2016arXiv160609070O} problems.
+Such schemes have also been empirically considered to accelerate the SK algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
+The convergence of these algorithms is nevertheless not studied yet in the context of regularized optimal transport.
+
+\section{Overrelaxed Sinkhorn--Knopp algorithm}
 
@@
 \begin{equation}\label{eq:lyapunov_function}
-F(\gamma) = KL(\gamma^*, \gamma)
+F(\gamma) = \KL(\gamma^*, \gamma)
 \end{equation}
@@
 \begin{lemma} \label{lemma:KL_compact}
-	For any $M \in \IR_+^*$, the sublevel set $K := \left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
+	For any $M \in \IR_+^*$, the sublevel set $\left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
 \end{lemma}
+\begin{proof}
+The fact that the Kullback-Leibler divergence is jointly lower semicontinuous implies in particular that $K$ is closed.
+Moreover, $K\subset \mathbb{R}^{n_1\times n_2}_+$ is bounded because $F$ is the sum of nonnegative, coercive functions of each coordinate.
+\end{proof}
+
+\iffalse
 \begin{proof}
@@
 \end{proof}
+\fi
 
@@
 
+Links with monotone Fista? (that choose Fista if energy decreased and forward backward otherwise...)
+
 \section*{Acknowledgments}

### 2017-10-19 paper.tex
Len first contrib

--- 2017-10-18 paper.tex
@@
 \usepackage{algorithmic}
+\usepackage{subcaption}
 %\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
@@
 
-\title{Over-relaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport}
+\title{Overrelaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport}
 
@@
  \And
-L\'enaic CHIZAT\\
+L\'ena\"ic CHIZAT\\
 \'Ecole Normale Sup\'erieure\\ Paris Dauphine (PSL research University)\\
@@
 \begin{abstract}
-This article describes a method for quickly calculating the solution to the regularized optimal transport problem.
+This article describes a method for quickly computing the solution to the regularized optimal transport problem.
 It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn--Knopp algorithm).
-The idea is to over-relax the Bregman projection operators, allowing for faster convergence.
+The idea is to overrelax the Bregman projection operators, allowing for faster convergence.
 In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
-In this paper, we propose a simple method for establishing the global convergence of the over-relaxed Sinkhorn--Knopp algorithm, by ensuring the decrease of a Lyapunov function at each step.
-An inexpensive adaptive choice of over-relaxation parameter based on the Lyapunov function is constructed.
-We also suggest a heuristic to choose a suitable asymptotic over-relaxation parameter.
+We propose a simple method for establishing global convergence by ensuring the decrease of a Lyapunov function at each step.
+An adaptive choice of overrelaxation parameter based on the Lyapunov function is constructed.
+We also suggest a heuristic to choose a suitable asymptotic over-relaxation parameter, based on a local convergence analysis.
+Our numerical experiments show a gain in convergence speed by an order of magnitude in certain regimes.
 \end{abstract}
@@
 \section{Introduction}
-Optimal Transport is an efficient and flexible tool to compare two probability distributions which has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
+Optimal Transport (OT) is an efficient and flexible tool to compare two probability distributions which has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
 The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64}   scaling with high dimensional data.
@@
 Hence, for particular purposes such as color interpolation \cite{Rabin2014} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
-In this setting,  the regularized transport problem can be ill-conditioned. This is the issue  we want to tackle here.
+In this setting,  the regularized transport problem can be ill-conditioned and Sinkhorn's algorithm converges slowly. This is the issue  we want to tackle here.
 Before going into further details, we now briefly introduce the main notations and concepts used all along this article.
 
-\subsection{Discrete Optimal Transport}
+\subsection{Discrete optimal transport}
 We consider two discrete probability measures $\mu^k \in \IR_{+*}^{n_k}$.
@@
 \begin{equation}\label{ROT}
-\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_{+*}}K^\epsilon(\gamma) := \scal{c}{\gamma} 
-+ \epsilon KL(\gamma,\mathbf{1})
+\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_{+}}K^\epsilon(\gamma) \coloneqq \scal{c}{\gamma} 
++ \epsilon \KL(\gamma,\mathbf{1})
 ,\end{equation}
-with $\epsilon>0$, $\mathbf{1}$ is the matrix of size $n_1\times n_2$ full of ones and the Kullback-Leibler divergence is:
+with $\epsilon>0$, $\mathbf{1}$ is the matrix of size $n_1\times n_2$ full of ones and the Kullback-Leibler divergence is
 \begin{equation}\label{KL}
-KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
+\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}
 \end{equation}
-It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$:
+with the convention $0 \log 0= 0$.
+It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ (here and in the sequel, exponentiation is meant entrywise) onto $\Ccal_1 \cap \Ccal_2$:
 \begin{equation}\label{eq:reg_ot_pb}
@@
 \end{equation}
-where $P_{\Ccal}$ is the  Bregman projection onto $\Ccal$ defined as:
+where $P_{\Ccal}$ is the  Bregman projection onto $\Ccal$ defined as
 \[
-P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} KL(\gamma,\xi).
+P_{\Ccal}(\xi) \coloneqq \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi).
 \]
 
-\subsection{The Sinkhorn--Knopp algorithm}
+\subsection{Sinkhorn--Knopp algorithm}
 Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
-Hence, the so-called Sinkhorn--Knopp algorithm \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
+Hence, the so-called Sinkhorn--Knopp algorithm (SK) \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
 \begin{align*}
@@
 where $\oslash$ is the pointwise division. 
-To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and to iterate
+To compute numerically the solution one simply has to store $(a^l, b^l)\in\IR^{n_1}\times \IR^{n_2}$ and to iterate
+\begin{align*}
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} .
+\end{align*}
+We then have $\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$
+
+Another way to interpret the SK algorithm is as an alternate maximization algorithm on the dual of the regularized optimal transport problem. The dual problem of \eqref{ROT} is
+\begin{equation}\label{DROT}
+\max_{\substack{\alpha\in \IR^n\\\beta\in \IR^m}}\; E(\alpha,\beta) \coloneqq \langle \alpha,\mu^1\rangle+\langle \beta,\mu^2\rangle-\epsilon\sum_{i,j}e^{(\alpha_i+\beta_j-c_{i,j})/\epsilon}.
+\end{equation}
+The function $E$ is concave, continuously differentiable and admits a maximizer, so alternate maximization converges and we recover, for $a_i=e^{\alpha_i/\epsilon}$, $b_j=e^{\beta_j/\epsilon}$ and $\gamma^0_{i,j}=e^{-c_{i,j}/\epsilon}$:
 \begin{align*}
@@
 \end{align*}
-where the transport matrix can be recovered as :
-$\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$ 
+
 Efficient parallel computations can be considered \cite{cuturi13} and one can even reach real-time computation for large scale problem for certain class of cost matrices $c$ \cite{Solomon2015}. 
 For small values of the parameter $\epsilon$, numerical issues can arise and a stabilization of the algorithm is necessary \cite{2016arXiv160705816C}.
-The convergence of the process can nevertheless be very slow  in the setting $\epsilon\to 0$.
+The convergence of the process can nevertheless be very slow when $\epsilon$ is small.
+
+\subsection{Overview and contributions}
+In this paper, we consider an overrelaxation scheme designed to accelerate the Sinkhorn--Knopp algorithm.
+We first present and show the convergence of our algorithm in section 2.
+In section 3, we analyze the local convergence rate of the algorithm to justify the acceleration.
+We finally demonstrate numerically  in Section 4 the good behavior of our method, where larger accelerations are observed for decreasing values of $\epsilon$.
+
 \subsection{Related works }
@@
 
-\subsection{Overview and contributions}
-In this paper, we consider an over-relaxation scheme designed to accelerate the Sinkhorn--Knopp algorithm.
-We first present and show the convergence of our algorithm in section 2.
-In section 3, we analyze the local convergence rate of the algorithm to justify the acceleration.
-We finally demonstrate numerically  in Section 4 the good behavior of our method, where larger accelerations are observed for decreasing values of $\epsilon$.
-
-\section{Over-relaxed Sinkhorn--Knopp algorithm}
+\section{Overrelaxed Sinkhorn--Knopp algorithm}
 
 As illustrated in Figure \ref{alternate_projections} (a-b), the SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
-The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
+The idea developped in this paper is to perform overrelaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
 
+\begin{figure}[ht!]
+\centering
+\begin{minipage}[b]{.33\linewidth}
+   \centering
+   \input{schema_a.tex}
+                    \subcaption{}\label{fig:schema_a}
+\end{minipage}%
+\begin{minipage}[b]{.33\linewidth}
+   \centering
+   \input{schema_b.tex}
+                    \subcaption{}\label{fig:schema_b}
+\end{minipage}%
+\begin{minipage}[b]{.33\linewidth}
+   \centering
+   \input{schema_c.tex}
+                    \subcaption{}\label{fig:schema_c}
+\end{minipage}%
+\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
+Overrelaxed projection typically accelerate the convergence rate.}
+\end{figure}
+
+\iffalse
 \begin{figure}[ht!]
@@
 \end{figure}
-\subsection{Over-relaxed projections}
+\fi
 
-Let us define the $\omega$-over-relaxed projection operator by:
+\subsection{Overrelaxed projections}
+
+We define the $\omega$-overrelaxed projection operator $P^\omega_{\Ccal_k}$ as
 \begin{equation}\label{eq:def_or_proj}
@@
 where the logarithm is taken coordinate-wise.
-Note that $P_{\Ccal_k}^0$ is the identity, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ is the standard Bregman projection, and $P_{\Ccal_k}^2$ is an involution (this uses the fact that $\Ccal_k$ is an affine subspace).
+Note that $P_{\Ccal_k}^0$ is the identity, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ is the standard Bregman projection and $P_{\Ccal_k}^2$ is an involution (in particular because $\Ccal_k$ is an affine subspace).
+A naive algorithm would then consist in iteratively applying $P^\omega_{\Ccal_2}\circ P^\omega_{\Ccal_1}$ for some choice of $\omega$.
+This algorithm, while it may behave well in practice, has no convergence guarantee---it sometimes does not convergence even for reasonable values of $\omega$.
+Our goal in this section is to modify this algorithm to make it robust and garantee convergence.
+
+Duality gives another point of view on the iterative overrelaxed Bregman projections.
+They indeed correspond to a successive overelaxation (SOR) algorithm on the dual objective $E$.
+This can be seen by using the relationships gievn after equation~\eqref{DROT}:
+\begin{align}\label{SORdual}
+\alpha^{\ell+1} &= (1-\omega)\alpha^{\ell} + \omega \arg \max_\alpha E(\alpha,\beta^{\ell})\\
+\beta^{\ell + 1}&=(1-\omega) \beta^{\ell} + \omega \arg \max_\beta E(\alpha^{\ell+1},\beta).
+\end{align}
 
@@
 State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
+lenaic: what is exactly needed?
+I can think of it tomorow.
 \end{proposition}}
 \subsection{Lyapunov function}
-Let $\gamma^*$ denote the solution of the regularized OT problem.
-The function $F$ is defined as:
+Convergence of the successive overrelaxed projection cannot be guaranteed in general. In order to derive a robust algorithm with provable convergence, we introduce the Lyapunov function 
 \begin{equation}\label{eq:lyapunov_function}
-F(\gamma) = KL(\gamma^*, \gamma)
+F(\gamma) = \KL(\gamma^*, \gamma),
 \end{equation}
-It shall be used as a Lyapunov function, so that $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+where $\gamma^*$ denotes the solution of the regularized OT problem.
+We will use this function to enforce the strict descent criterion $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+%
+The choice of \eqref{eq:lyapunov_function} as a Lyapunov function is of course related to the fact that Bregman projections are used throughout the algorithm.
+Further, we will show (Lemma \ref{lemma:lyapunov_decrease}) that its decrease is very easy to compute and this descent criterion still allows enough freedom in the choice of the overrelaxation parameter.
 
-We argue that it makes sense to consider this Lyapunov function.
-Obviously, it is closely related to the notion of Bregman projection that is used throughout the algorithm.
-But it is also easy to calculate its decrease, as shown in lemma \ref{lemma:lyapunov_decrease}.
-Moreover, in practice it allows the use of a wide range of over-relaxation parameters.
+Crucial properties of this Lyapunov function are gathered in the next lemma.
+\begin{lemma} \label{lemma:lyapunov_decrease}
+	For any $M \in \IR_+^*$, the sublevel set $\left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
+%\end{lemma}
+%\begin{lemma}
+	Moreover, for $\gamma$ in $\IR^{mn}_{+*}$, the decrease of the Lyapunov function after an overrelaxed projection can be computed as
+	\begin{equation} \label{eq:kl_diff_scal}
+	F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma)) = 
+	\scal{\mu^k}{\varphi_\omega \left((A_k \gamma) \oslash \mu^k \right)},
+	\end{equation}
+	where
+	\begin{equation}
+	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega \log x
+	\end{equation}
+	is a real function, applied coordinate-wise.
+\end{lemma}
 
-\begin{lemma} \label{lemma:KL_compact}
-	For any $M \in \IR_+^*$, the sublevel set $K := \left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
-\end{lemma}
+\begin{proof}
+The fact that the Kullback-Leibler divergence is jointly lower semicontinuous implies in particular that $K$ is closed.
+Moreover, $K\subset \mathbb{R}^{n_1\times n_2}_+$ is bounded because $F$ is the sum of nonnegative, coercive functions of each component of its argument $\gamma$.
+%\end{proof}
+%\begin{proof}
+%
+Formula \eqref{eq:kl_diff_scal} comes the expression $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$ and the relations \eqref{eq:def_or_proj} and \eqref{scaling}.
+\end{proof}
+
+\iffalse
 \begin{proof}
@@
 \end{proof}
+\fi
 
-Note that the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$ may be calculated without knowing $\gamma^*$, as shown by the following lemma.
-\begin{lemma}\label{lemma:lyapunov_decrease}
-	Take $\gamma$ in $\IR^{mn}_{+*}$. The decrease in value of the Lyapunov function can be calculated with the following formula:
-	\begin{equation} \label{eq:kl_diff_scal}
-	F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma)) = 
-	\scal{\mu^k}{\varphi_\omega \left(\frac{A_k \gamma}{\mu^k}\right)},
-	\end{equation}
-	where
-	\begin{equation}
-	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega \log x
-	\end{equation}
-	is a real function, applied coordinate-wise.
-\end{lemma}
-\begin{proof}
-From \eqref{KL}, we have $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$.
-The result can then be deduced from relations \eqref{eq:def_or_proj} and \eqref{scaling}.
-\end{proof}
-Thus,  the decrease in value of the function $F$ for a over-relaxed projection associated to $\Ccal_k$ is cheap to estimate, since its computational cost is  linear with respect to the dimension of data $\mu^k$.
-In Figure \ref{phi_omega}, we illustrate the functions  $\varphi_\omega$ for several values of $\omega$.
-Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative. For other values $1\le\omega<2$, it is nonnegative for arguments close to 1.
+%Note that the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$ may be calculated without knowing $\gamma^*$, as shown by the following lemma.
+
+It follows from Lemma \ref{lemma:lyapunov_decrease} that the decrease of $F$ for an overrelaxed projection is very cheap to estimate, since its computational cost is linear with respect to the dimension of data $\mu^k$.
+In Figure \ref{phi_omega}, we display the function  $\varphi_\omega(x)$.
+Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative.
+For other values $1\le\omega<2$, it is nonnegative for $x$ close to 1.
+
 \begin{figure}[ht!]
 \begin{center}
-\includegraphics[width=11cm]{phi_omega.png}
-\caption{\label{phi_omega} Plot of functions $\varphi_\omega(x)$ for different values $1\leq \omega\leq 2$ and   $x\in[10^{-2}, 10^1]$.}
+\includegraphics[width=7cm]{images/cvgce_zone}
+\caption{\label{phi_omega} Value of $\varphi_\omega(x)$.
+The function is positive above the black line, negative below.
+For any relaxation parameter $\omega$ smaller than $2$, there exists a neighborhood of $1$ on which $\varphi_\omega(\cdot)$ is positive.}
+%\includegraphics[width=11cm]{phi_omega.png}
+%\caption{\label{phi_omega} Plot of functions $\varphi_\omega(x)$ for different values $1\leq \omega\leq 2$ and   $x\in[10^{-2}, 10^1]$.}
 \end{center}
 \end{figure}
+
 \subsection{Proposed algorithm}
-
+We first give a general convergence result that later serves as a basis to design an explicit algorithm.
 \begin{theorem}\label{thm:algo}
-	Take two continuous functions $\theta_1,\theta_2$ such that
+Let $\theta_1$ and $\theta_2$ be two continuous functions of $\gamma$ such that
 	\begin{equation}\label{eq:cond_theta_k}
@@
 	where the inequality is strict whenever $\gamma \notin \Ccal_k$.
-	Consider the sequence defined by:
+	Consider the sequence defined by $\gamma^0 = e^{-c/\epsilon}$ and
 	\begin{align*}
-	\gamma^0 &= e^{-c/\epsilon} \\
 	\tilde{\gamma}^{l+1} &= P_{\Ccal_1}^{\theta_1(\gamma^l)}(\gamma^l) \\
@@
 	\]
-
 	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^l))$ is non-increasing. Hence $(\gamma^l)$ is precompact.
-
+	%
 	If $\xi$ is a cluster point of $(\gamma^l)$, let us define
@@
 	\end{align*}
-	Then by continuity of all the applications, $F(\xi) = F(\tilde{\xi}) = F(\xi')$.
-	From the hypothesis made on $\theta_1$ and $\theta_2$, can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
+	Then by continuity of the applications, $F(\xi) = F(\tilde{\xi}) = F(\xi')$.
+	From the hypothesis made on $\theta_1$ and $\theta_2$, it can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
 Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
-	By lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
+	By Lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
 \end{proof}
@@
 %	TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
-To construct the functions $\theta_k$ needed by the theorem, we need the following lemma.
 
+To construct explicitly functions $\theta_k$ as needed in Theorem~\ref{thm:algo}, we need the following lemma.
 \begin{lemma}\label{lemma:F_P_theta}
@@
 	Moreover, the function that maps $t \in [1,\infty)$ to $\varphi_t(x)$ is non-increasing:
-	$\frac{d}{dt} \varphi_t(x) = \log x (x^{1-t} - 1).$
+	$\partial_t \varphi_t(x) =  (x^{1-t} - 1)\log x.$
 	For $x\neq 1$, it is even strictly decreasing.
-	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_k \gamma = \mu$.
+	Thus inequality \eqref{eq:F_P_theta} is valid, with equality \emph{iff} $A_k \gamma = \mu$.
 \end{proof}
 
-\begin{proposition}
-	We argue that a good choice for the functions $\theta_k$ may be constructed as follows. Pick a target parameter $\theta_0 \in [1;2)$, and a small $\delta>0$.
-	Define the functions $\Theta^*$ and $\Theta$ as:
+We now argue that a good choice for the functions $\theta_k$ may be constructed as follows.
+Pick a target parameter $\theta_0 \in [1;2)$ and a small security distance $\delta>0$.
+Define the functions $\Theta^*$ and $\Theta$ as
 	\begin{align}
+	\label{eq:Theta_opt}
 	\Theta^*(u) &= \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u\right) \ge 0 \right\} ,\\
@@
 	where $\min u$ denotes the smallest coordinate of the vector $u$.
-	Then the function
+
+\begin{proposition}\label{prop:thetachoice}
+	The function
 	\begin{equation}
 	\label{eq:theta_k}
-	\theta_k(\gamma) =\Theta\left (\frac{A_k \gamma}{\mu^k}\right)
+	\theta_k(\gamma) =\Theta\left ((A_k \gamma)\oslash \mu^k\right)
 	\end{equation}
-	is continuous, and verifies condition (\ref{eq:cond_theta_k}).
+	is continuous and verifies the descent condition (\ref{eq:cond_theta_k}).
 \end{proposition}
-This construction has several advantages.
-One is that it allows to choose arbitrarily the parameter $\theta_0$ that will be used asymptotically. A heuristic for this choice is proposed in section \ref{section:local}.
-But it is also an easy approach to having an adaptive method, as the approximation of $\theta_k^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $\frac{A_k \gamma}{\mu^k}$, and the corresponding function to invert is decreasing and concave).
-We will show later on that this is often an excellent choice in practice.
+This construction has several advantages:
+\begin{itemize}
+\item it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
+\item it is also an easy approach to having an adaptive method, as the approximation of $\Theta_k^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
+\item experiments show that this is often an excellent choice in practice.
+\end{itemize}
+
 \begin{proof}
-	Since the partial derivative of $\frac{\partial\varphi_\omega(x)}{\partial\omega}$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
-	The function $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
+Looking at Figure~\ref{phi_omega} can help understanding this proof.	Since the partial derivative of $\partial_\omega \varphi_\omega(x)$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
+	The function $\Theta^*\left((A_k \gamma)\oslash \mu^k)\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
 	Therefore, by lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
-	It is no problem to constrain it to $[1,\theta_0]$, as using 1 always verifies the inequality (\ref{eq:cond_theta_k}), and using $\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
+	Constraining the parameter to $[1,\theta_0]$ preserves this properties.% as using $\theta=1$ always satisfies inequality \eqref{eq:cond_theta_k} and using $\theta=\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
 \end{proof}
 
-\section{Local rate of convergence ?}
-\label{section:local}
-Some material to adapt Lenaic results:
-
-The SK algorithm can be interpreted as an alternate maximisation algorithm on the dual of the regularized optimal transport problem.
-The dual problem of \eqref{ROT} is:
-\begin{equation}\label{DROT}\max_{\alpha,\beta}\langle \alpha,\mu^1\rangle+\langle \beta,\mu^2\rangle-\epsilon\sum_{i,j}e^{(\alpha_i+\beta_j-c_{i,j})/\epsilon}\end{equation}
-It is concave and continuously differentiable for all $(a,b)\in\IR^{n_1+n_2}$.
-Alternate maximization then converges and we recover, for $a_i=e^{\alpha_i/\epsilon}$, $b_j=e^{\beta_j/\epsilon}$ and $\gamma^0_{i,j}=e^{-c_{i,j}/\epsilon}$:
-\begin{align*}
-a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
-b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} 
-\end{align*}
-
-If $(\alpha^*,\beta^*)$ is a solution of \eqref{DROT}, then all the 
-other optimum can be written as $(\alpha^*+\kappa,\beta^*-\kappa)$, which means that the problem admits a unique solution, up to a translation with constant $\kappa\in\IR$.
-
-The Hessian matrix of the functional in \eqref{DROT} at point $(\alpha,\beta)$ reads, for $a_i=e^{\alpha_i/\epsilon}>0$ and $b_j=e^{\beta_j/\epsilon}>0$:
-
-$$H(a,b)=\frac1\epsilon\begin{bmatrix}
-\diag(a)\diag(\gamma^0 b)&\diag(a)\gamma^0 \diag(b)\\
-\diag(b){^t\gamma^0} \diag(a)&\diag(^t\gamma^0a)\diag(b)
-\end{bmatrix}.$$
-Notice that
-
-$$H(a^*,b^*)=\frac1\epsilon\begin{bmatrix}
-\diag(\mu^1)&\gamma^*\\
-^t\gamma^*&\diag(\mu^2)
-\end{bmatrix}.$$
-
-The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
-Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite for $a_i,b_j>0$.
-Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.
-All the other eigenvalues belongs to the set $[0;1[$.
-
-From experiments, it seems possible to design a good guess for $\omega$ from $c$ and $\epsilon$ only. 
-    Denoting as $\lambda$ the second larger eigenvalue of $T(1,1)$, one can set 
-    $\omega=\frac{2}{\lambda}(1-\sqrt{1-\lambda})$.
-
-\section{Experimental results}
-
-%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
-%The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
-
-\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
-%\begin{algorithm}
-%\caption{Over-relaxed SK algorithm}
-%\label{SOR}
-%\begin{algorithmic}
-%\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
-%\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
-%\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
-%\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
-%\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
-%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
-%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed SK iteration}
-%\ELSE
-%\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
-%\ENDIF
-%\ENDWHILE
-% \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
-%\end{algorithmic}
-%\end{algorithm}
-
-This algorithm uses the function $\Theta$ defined in (\ref{eq:Theta}).
+The resulting algorithm, which is proved to be convergent by Theorem~\ref{thm:algo}, is written in pseudo-code in Algorithm \ref{SOR}.
+It uses the function $\Theta$ defined implicitely in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
+Note that with the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
 
 \begin{algorithm}
-\caption{Over-relaxed SK algorithm }
+\caption{Overrelaxed SK algorithm }
 \label{SOR}
@@
 
-{\color{red} TODO}
+\section{Acceleration of local convergence rate}
+\label{section:local}
+In order to theoretically jsutify the acceleration of convergence that is observed in practice, we now study the local convergence rate of the algorithm.
+%We do not claim that this simple overrelaxation is the most efficient method to accelerate local convergence of fixed points---adaptive methods with excellent properties exist e.g.~\cite{2016arXiv160604133S}--- but this local analysis gives a hint for what makes our scheme relevant is the conjunction of our global convergence guarantee with this local acceleration property.
+
+%The local behavior of the extrapolated is summarized in the next result, which gathers many results from the literature.
+\begin{proposition}\label{prop:local}
+Let $1-\eta$ be the local linear rate of convergence of the SK algorithm.
+By \cite{knight2008sinkhorn} we know that $\eta>0$.
+Then, for the optimal choice of parameter $\theta_0 = 2/(1+\sqrt{\eta})$, the overrelaxed projection algorithm converges locally linearly at a rate $(1-\sqrt{\eta})/(1+\sqrt{\eta})$.
+\end{proposition}
+
+Note that the local linear rate is also improved if $\theta_0$ is not the optimal choice, see \cite[chapter 5]{ciarlet1982introduction}.
+By remarking that for any choice of $\theta_0<2$ there is a neighborhood of $1$ on which $\varphi_{\theta_0}(\cdot)$ is nonnegative, one sees that this result applies in particular to our algorithm because eventually, for $\ell$ big enough, one has $\Theta(\gamma^\ell)=\theta_0$.
+\begin{corollary}
+The precedding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
+\end{corollary}
+
+\begin{proof}
+For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual} and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
+Notice that the dual problem~\eqref{DROT} has an invariant subspace due to the invariance by translation $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is strictly convex in any subspace which does not contain this line.
+Also, this invariance disappears in the associated primal iterates.
+
+The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative}, even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.
+
+Using standard arguments, we have that the local convergence rate of the SOR procedure on the dual objective $E$ is that of the SOR procedure applied to the quadratic form defined by its Hessian and the linear rate of \emph{semi-convergence} to the projection on the set of minimizers.
+\end{proof}
+
+%If $(\alpha^*,\beta^*)$ is a solution of \eqref{DROT}, then all the other optimum can be written as $(\alpha^*+\kappa,\beta^*-\kappa)$, which means that the problem admits a unique solution, up to a translation with constant $\kappa\in\IR$.
+
+\iffalse
+The Hessian matrix of the functional in \eqref{DROT} at point $(\alpha,\beta)$ reads, for $a_i=e^{\alpha_i/\epsilon}>0$ and $b_j=e^{\beta_j/\epsilon}>0$:
+
+$$H(a,b)=\frac1\epsilon\begin{bmatrix}
+\diag(a)\diag(\gamma^0 b)&\diag(a)\gamma^0 \diag(b)\\
+\diag(b){^t\gamma^0} \diag(a)&\diag(^t\gamma^0a)\diag(b)
+\end{bmatrix}.$$
+Notice that
+
+$$H(a^*,b^*)=\frac1\epsilon\begin{bmatrix}
+\diag(\mu^1)&\gamma^*\\
+^t\gamma^*&\diag(\mu^2)
+\end{bmatrix}.$$
+
+The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
+Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite for $a_i,b_j>0$.
+Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.
+All the other eigenvalues belongs to the set $[0;1[$.
+
+From experiments, it seems possible to design a good guess for $\omega$ from $c$ and $\epsilon$ only. 
+    Denoting as $\lambda$ the second larger eigenvalue of $T(1,1)$, one can set 
+    $\omega=\frac{2}{\lambda}(1-\sqrt{1-\lambda})$.
+
+    \fi
+
+\section{Experimental results}
+We compare Algorithm~\ref{SOR} to SK in two very different settings.
+In setting (a) we consider the domain $[0,1]$ discretized into $100$ samples and the squared Euclidean transport cost on this domain.
+The marginals are densities made of the sum of a base plateau of height $0.1$ and another plateau of heights and boundaries chosen uniformly in $[0,1]$, subsequently normalized.
+In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.%The interest of having two settings is that the SK algorithm is known to behave very differently as $\epsilon$ decreases for these two problems and
+
+Given an estimation of $1-\eta$, the local convergence rate of SK algorithm, we define $\theta_0$ as the optimal parameter as given in Proposition \ref{prop:local}.
+For estimating $\eta$, we follow two strategies.
+For strategy ``estimated'' (in blue on figure~\ref{fig:comparison}), $\eta$ is measured by looking at the local convergence rate of SK run on another random problem of the same setting and for the same value of $\epsilon$.
+For strategy ``measured'' (in orange on figure~\ref{fig:comparison}) the parameter is set using the local convergence rate of SK run on the same problem.
+Of course, the latter is unrealistic strategy but our experiment shows that the acceleration power is not so much impacted by estimating once for all the overrelaxation parameter for a class of problem with fixed regularization, as shown on~\ref{fig:comparison}.
+
+Figure~\ref{fig:comparison} displays the ratio of the number of iterations required to reach a precision of $10^{-6}$ on the dual variable $\alpha$ for SK algorithm and Algorithm \ref{SOR}.
+It is is worth noting that the complexity per iteration is the same, modulo negligible terms, so this ratio is also the runtime ratio (our algorithm can also be parallelized on GPUs just as SK algorithm).
+
+%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
+%The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
+
+\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
+%\begin{algorithm}
+%\caption{Over-relaxed SK algorithm}
+%\label{SOR}
+%\begin{algorithmic}
+%\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
+%\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
+%\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
+%\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
+%\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
+%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
+%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed SK iteration}
+%\ELSE
+%\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
+%\ENDIF
+%\ENDWHILE
+% \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
+%\end{algorithmic}
+%\end{algorithm}
+%Add algorithm to solve \eqref{eq:Theta_opt}?
+
+\begin{figure}
+\begin{minipage}[b]{.5\linewidth}
+   \centering
+   \includegraphics[scale=0.5]{images/speedratio_image}
+                    \subcaption{Quadratic cost, random marginals}\label{fig:compare_a}
+\end{minipage}%
+			\begin{minipage}[b]{.5\linewidth}
+                    \centering
+			\includegraphics[scale=0.5]{images/speedratio_ML}
+                    \subcaption{Random cost, uniform marginals}\label{fig:compare_b}
+\end{minipage}
+			\caption{\label{fig:comparison}Complexity ratio of SK algorithm and its accelerated version Algorithm~\ref{SOR}.}% (stopping criterion $\Vert u - u^* \Vert_\infty<10^{-6})$
+\end{figure}
 
 \section{Conclusion and perspectives}
-Contributions: 
-cheap test, preliminary work for convergence results. Huge acceleration in practice.
+SK algorithm is widely used to solve entropy regularized OT.
+The use of overrelaxed projections turns out to be a natural and simple idea to accelerate convergence while keeping many nice properties of this algorithm (first order, paralelizable, simple).
+We have proposed an algorithm that adaptively chooses the overrelaxation parameter so as to guarantee global convergence.
+The acceleration of the convergence speed is numerically impressive, in particular in low regularization regimes.
+It is theoretically supported in the local regime by the standard analysis of SOR iterations.
 
-Perspectives:
-global convergence, how chosing $\omega$?, intensive numerical study with respect to other accelerarions methods\cite{2016arXiv160604133S,2017arXiv170509634A}.
+This idea of overrelaxation can be generalized to solve more general problems such as barycenters, gradient flows, unbalanced OT~\cite[chap.
+4]{chizat2017these} but there is no systematic way to derive globally convergent algorithms.
+Our work as an effort in the direction of building and understanding the properties of robust first order algorithms for solving OT.
+More understanding is needed regarding SOR itself (global convergence speed, choice of $\theta_0$), but also its relation to other acceleration methods~\cite{2016arXiv160604133S,2017arXiv170509634A}.
+
+%Links with monotone Fista? (that choose Fista if energy decreased and forward backward otherwise...)
 

### 2017-10-19 paper.tex

@@
 \end{document}
+=======
+\documentclass{article} % For LaTeX2e
+\usepackage{nips14submit_e,times}
+\usepackage{hyperref}
+\usepackage{url}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{mathtools,enumitem}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
+
+\usepackage{tikz}
+\usepackage{cite}
+
+% Math operators
+\newcommand{\scal}[2]{\left\langle #1 , #2 \right\rangle}
+\DeclareMathOperator{\IR}{\mathbb{R}}
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator{\One}{\mathbbm{1}}
+\DeclareMathOperator{\Ccal}{\mathcal{C}}
+\DeclareMathOperator{\logsumexp}{logsumexp}
+\DeclareMathOperator{\diag}{diag}
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
+\renewcommand{\epsilon}{\varepsilon}
+
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{proposition}{Proposition}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\newtheorem{example}{Example}
+
+\title{Over-relaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport}
+
+%
+\author{
+Alexis THIBAULT\\
+\'Ecole Normale Sup\'erieure\\
+Paris, France\\
+\texttt{alexis.thibault@ens.fr}
+ \And
+L\'enaic CHIZAT\\
+\'Ecole Normale Sup\'erieure\\ Paris Dauphine (PSL research University)\\
+Paris, France\\
+\texttt{lenaic.chizat@ens.fr}
+ \AND
+Charles DOSSAL\\
+INSA Toulouse\\
+Institut de Math\'ematiques de Toulouse\\
+Toulouse,  France\\
+\texttt{dossal@insa-toulouse.fr}
+\And 
+Nicolas PAPADAKIS\\
+CNRS, Institut de Math\'ematiques de Bordeaux\\
+Talence, France\\
+\texttt{nicolas.papadakis@math.u-bordeaux.fr}
+} 
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
+%
+% Using \And between authors leaves it to \LaTeX{} to determine where to break
+% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
+% puts 3 of 4 authors names on the first line, and the last on the second
+% line, try using \AND instead of \And before the third author name.
+
+\newcommand{\fix}{\marginpar{FIX}}
+\newcommand{\new}{\marginpar{NEW}}
+
+\nipsfinalcopy % Uncomment for camera-ready version
+
+\begin{document}
+
+\maketitle
+
+\begin{abstract}
+This article describes a method for quickly calculating the solution to the regularized optimal transport problem.
+It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn--Knopp algorithm).
+The idea is to over-relax the Bregman projection operators, allowing for faster convergence.
+In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
+In this paper, we propose a simple method for establishing the global convergence of the over-relaxed Sinkhorn--Knopp algorithm, by ensuring the decrease of a Lyapunov function at each step.
+An inexpensive adaptive choice of over-relaxation parameter based on the Lyapunov function is constructed.
+We also suggest a heuristic to choose a suitable asymptotic over-relaxation parameter.
+\end{abstract}
+
+\section{Introduction}
+Optimal Transport is an efficient and flexible tool to compare two probability distributions which has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
+The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64}   scaling with high dimensional data.
+Regularized optimal transport have thus been intensively used  in  Machine Learning with applications such as   Geodesic PCA \cite{seguy2015principal}, domain adaptation \cite{2015arXiv150700504C}, data fitting \cite{2015arXiv150605439F},  training of Boltzmann Machine \cite{NIPS2016_6248}  or dictionary learning \cite{Rolet2016,2017arXiv170801955S}.
+
+The computation of optimal transport between two data relies on the estimation of an optimal transport matrix, the entries of which represent the quantity of mass transported between  data locations. 
+Regularization of optimal transport with strictly convex regularizers \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
+Hence, for particular purposes such as color interpolation \cite{Rabin2014} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
+In this setting,  the regularized transport problem can be ill-conditioned, rendering the Sinkhorn--Knopp algorithm significantly slower. This is the issue  we want to tackle here.
+Before going into further details, we now briefly introduce the main notations and concepts used all along this article.
+
+\subsection{Discrete Optimal Transport}
+We consider two discrete probability measures $\mu^k \in \IR_{+*}^{n_k}$.
+Let us define the two following linear operators
+\begin{align*}
+A_1 &: \begin{cases}
+\IR^{n_1 n_2} \rightarrow \IR^{n_1} \\
+(A_1 x)_i = \sum_j x_{i,j}
+\end{cases} &
+A_2 &: \begin{cases}
+\IR^{n_1 n_2} \rightarrow \IR^{n_2}\\
+(A_2 x)_j = \sum_i x_{i,j},
+\end{cases}
+\end{align*}
+as well as the linear  constraint sets
+\begin{align*}
+\Ccal_k &= \left\{ \gamma\in\IR^{n_1 n_2} \mid A_k \gamma = \mu^k \right\}.
+\end{align*}
+Given a cost matrix $c$, where $c_{ij}$ represents the cost of moving mass $\mu^1_i$ to $\mu^2_j$,  the optimal transport problem corresponds to the estimation of an optimal transport matrix $\gamma$ solution of:
+$$\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_+} \langle c,\gamma\rangle:=\sum_{i,j}c_{i,j}\gamma_{i,j}.$$
+This is a linear programing problem whose resolution is unmanageable for large scale problems.
+
+\subsection{Regularized optimal transport}
+
+In \cite{cuturi13}, it has been proposed to regularize this problem by adding a strictly convex entropy regularization:
+\begin{equation}\label{ROT}
+\min_{\gamma\in\Ccal_1\cap \Ccal_2\cap \IR^{n_1 n_2}_{+*}}K^\epsilon(\gamma) := \scal{c}{\gamma} 
++ \epsilon KL(\gamma,\mathbf{1})
+,\end{equation}
+with $\epsilon>0$, $\mathbf{1}$ is the matrix of size $n_1\times n_2$ full of ones and the Kullback-Leibler divergence is:
+\begin{equation}\label{KL}
+KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
+\end{equation}
+It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ onto $\Ccal_1 \cap \Ccal_2$:
+\begin{equation}\label{eq:reg_ot_pb}
+\gamma^* = \argmin_{\Ccal_1 \cap \Ccal_2} K^\epsilon(\gamma)= P_{\Ccal_1 \cap \Ccal_2} (e^{-c/\epsilon}),
+\end{equation}
+where $P_{\Ccal}$ is the  Bregman projection onto $\Ccal$ defined as:
+\begin{equation}\label{eq:P_C}
+P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} KL(\gamma,\xi).
+\end{equation}
+
+\subsection{The Sinkhorn--Knopp algorithm}
+Iterative Bregman projections onto $\Ccal_1$ and $\Ccal_2$ converge to a point in the intersection $\Ccal_1 \cap \Ccal_2$ \cite{bregman67}.
+Hence, the so-called Sinkhorn--Knopp algorithm \cite{sinkhorn64} that performs alternate Bregman projections, can be considered to compute the regularized  transport matrix:
+\begin{align*}
+\gamma^0 &= e^{-c/\epsilon} &
+\gamma^{l+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l)),
+\end{align*}
+and we have 
+$\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
+
+In the discrete setting,  projections actually correspond to diagonal scalings. This can be seen by writing the minimality conditions for the projection, given equations (\ref{KL}) and (\ref{eq:P_C}).
+\begin{align}\label{scaling}
+P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &
+a &=  {\mu^1}\oslash{A_1 \gamma} \\
+P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &
+b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
+\end{align}
+where $\oslash$ is the pointwise division. 
+To compute numerically the solution  one simply has to store $(a^l, b^l)\in\IR^{n_1+n_2}$ and to iterate
+\begin{align*}
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} 
+\end{align*}
+where the transport matrix can be recovered as :
+$\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$ 
+Efficient parallel computations can be considered \cite{cuturi13} and one can even reach real-time computation for large scale problem for certain class of cost matrices $c$ \cite{Solomon2015}. 
+For small values of the parameter $\epsilon$, numerical issues can arise and a stabilization of the algorithm is necessary \cite{2016arXiv160705816C}.
+The convergence of the process can nevertheless be very slow  in the setting $\epsilon\to 0$.
+\subsection{Related works }
+The introduction of relaxation variables through heavy ball approaches \cite{POLYAK19641} has recently gained in popularity  to speed up the convergence of algorithms optimizing convex \cite{2014arXiv1412.7457G} or non convex \cite{Zavriev1993,2016arXiv160609070O} problems.
+Such schemes have also been empirically considered to accelerate the SK algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
+The convergence of these algorithms is nevertheless not studied yet in the context of regularized optimal transport.
+
+\subsection{Overview and contributions}
+In this paper, we consider an over-relaxation scheme designed to accelerate the Sinkhorn--Knopp algorithm.
+We first present and show the convergence of our algorithm in section 2.
+In section 3, we analyze the local convergence rate of the algorithm to justify the acceleration.
+We finally demonstrate numerically  in Section 4 the good behavior of our method, where larger accelerations are observed for decreasing values of $\epsilon$.
+
+\section{Over-relaxed Sinkhorn--Knopp algorithm}
+
+As illustrated in Figure \ref{alternate_projections} (a-b), the SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$, as the problem becomes ill-conditioned.
+The idea developped in this paper is to perform over-relaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
+
+\begin{figure}[ht!]
+\begin{center}
+\begin{tabular}{ccc}
+\input{schema_a.tex}&
+\input{schema_b.tex}&
+\input{schema_c.tex}\\
+(a)&(b)&(c)
+\end{tabular}
+\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
+In order to accelerate the convergence rate, over-relaxed projections are here considered (c).}
+\end{center}
+\end{figure}
+\subsection{Over-relaxed projections}
+
+Let us define the $\omega$-over-relaxed projection operator by:
+\begin{equation}\label{eq:def_or_proj}
+\log P^\omega_{\Ccal_k}(\gamma) = (1-\omega) \log \gamma + \omega \log P_{\Ccal_k}(\gamma),
+\end{equation}
+where the logarithm is taken coordinate-wise.
+Note that $P_{\Ccal_k}^0$ is the identity, $P_{\Ccal_k}^1 = P_{\Ccal_k}$ is the standard Bregman projection, and $P_{\Ccal_k}^2$ is an involution (recall that $\Ccal_k$ is an affine subspace).
+
+{\color{blue} \begin{proposition} 
+State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
+\end{proposition}}
+\subsection{Lyapunov function}
+Let $\gamma^*$ denote the solution of the regularized OT problem.
+The function $F$ is defined as:
+\begin{equation}\label{eq:lyapunov_function}
+F(\gamma) = KL(\gamma^*, \gamma)
+\end{equation}
+It shall be used as a Lyapunov function, so that $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+
+We argue that it makes sense to consider this Lyapunov function.
+Obviously, it is closely related to the notion of Bregman projection that is used throughout the algorithm.
+But it is also easy to calculate its decrease, as shown in lemma \ref{lemma:lyapunov_decrease}.
+Moreover, as we will see later, it allows the use of a wide range of over-relaxation parameters in practice.
+
+\begin{lemma} \label{lemma:KL_compact}
+	For any $M \in \IR_+^*$, the sublevel set $K := \left\{ \gamma \mid F(\gamma) \le M \right\}$ is compact.
+\end{lemma}
+\begin{proof}
+	First note that K is closed in $\IR_{+*}^{n_1 n_2}$, by continuity of $F$.
+	For $\gamma \in K$, one has:
+	\begin{align*}
+	F(\gamma) &= \scal{\gamma^*}{\log \gamma^* - 1} + \scal{\gamma}{1} - \scal{\gamma^*}{\log \gamma}\\
+	\sum_{i,j} \gamma_{i,j} - \gamma^*_{i,j} \log \gamma_{i,j} &\le M - \scal{\gamma^*}{\log \gamma^* - 1} .
+	\end{align*}
+	By concavity, the logarithm admits a simple upper bound: $\log \gamma_{i,j} \le \frac{\gamma_{i,j}}{e}$. Moreover, since $\gamma^*$ is a solution of the problem, each of its coordinates is at most 1.
+	\begin{align*}
+	\sum_{i,j} \gamma_{i,j} &\le \frac{M - \scal{\gamma^*}{\log \gamma^*-1}}{1-\frac{1}{e}} =: M'.
+	\end{align*}
+	Therefore each coordinate $\gamma_{i,j}$ is bounded above by $M'$.
+	Let us show that the coordinates are bounded below by a strictly positive number; from which can be deduced that $K$ is bounded and closed in $\IR^{n_1 n_2}$, and therefore compact.
+	\begin{align*}
+	F(\gamma)
+	&\ge \scal{\gamma^*}{\log \gamma^*} - \sum_{i,j} \gamma^*_{i,j} \log(\gamma_{i,j}) - 1\\
+	&\ge \scal{\gamma^*}{\log \gamma^*} - \gamma^*_{i_0,j_0} \log(\gamma_{i_0,j_0}) - 1 - \log M',
+	\end{align*}
+	where the last equation derives from $\log(\gamma_{i,j}) \le \log M'$, and $\sum_{(i,j)\neq(i_0,j_0)} \gamma^*_{i,j} \le 1$.
+	Finally:
+	\[
+	\log(\gamma_{i_0,j_0}) \ge \frac{\scal{\gamma^*}{\log \gamma^*} - M - \log M' - 1}{\gamma^*_{i_0,j_0}},
+	\]
+	and $K$ is thus compact.
+\end{proof}
+
+Note that the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$ may be calculated without knowing $\gamma^*$, as shown by the following lemma.
+\begin{lemma}\label{lemma:lyapunov_decrease}
+	Take $\gamma$ in $\IR^{mn}_{+*}$. The decrease of the Lyapunov function when applying an over-relaxed projection can be calculated with the following formula:
+	\begin{equation} \label{eq:kl_diff_scal}
+	F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma)) = 
+	\scal{\mu^k}{\varphi_\omega \left(\frac{A_k \gamma}{\mu^k}\right)},
+	\end{equation}
+	where
+	\begin{equation}
+	\varphi_\omega(x) = x(1-x^{-\omega}) - \omega \log x
+	\end{equation}
+	is a real function, applied coordinate-wise.
+\end{lemma}
+\begin{proof}
+From \eqref{KL}, we have $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$.
+The result can then be deduced from relations \eqref{eq:def_or_proj} and \eqref{scaling}.
+\end{proof}
+Thus, the decrease of the Lyapunov function $F$ for an over-relaxed projection associated to $\Ccal_k$ is cheap to estimate, since its computational cost is  linear with respect to the dimension of data $\mu^k$.
+In Figure \ref{phi_omega}, we illustrate the functions  $\varphi_\omega$ for several values of $\omega$.
+Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always non-negative. For other values $1\le\omega<2$, it is non-negative for arguments close to 1.
+\begin{figure}[ht!]
+\begin{center}
+\includegraphics[width=11cm]{phi_omega.png}
+\caption{\label{phi_omega} Plot of functions $\varphi_\omega(x)$ for different values $1\leq \omega\leq 2$ and   $x\in[10^{-2}, 10^1]$.}
+\end{center}
+\end{figure}
+\subsection{Proposed algorithm}
+
+\begin{theorem}\label{thm:algo}
+	Take two continuous functions $\theta_1,\theta_2$ such that
+	\begin{equation}\label{eq:cond_theta_k}
+	\forall \gamma \in \IR_{+*}^{n_1 n_2},\quad
+	F(P_{\Ccal_k}^{\theta_k(\gamma)}(\gamma)) \le F(\gamma) ,
+	\end{equation}
+	where the inequality is strict whenever $\gamma \notin \Ccal_k$.
+	Consider the sequence defined by:
+	\begin{align*}
+	\gamma^0 &= e^{-c/\epsilon} \\
+	\tilde{\gamma}^{l+1} &= P_{\Ccal_1}^{\theta_1(\gamma^l)}(\gamma^l) \\
+	\gamma^{l+1} &= P_{\Ccal_2}^{\theta_2(\tilde{\gamma}^{l+1})}(\tilde{\gamma}^{l+1}).
+	\end{align*}
+	Then the sequence $(\gamma^l)$ converges to $\gamma^*$.
+\end{theorem}
+
+\begin{lemma}[]
+	\label{lemma:trivial_intersection}
+	Let us take $\gamma^0$ in $\IR_{+*}^{n_1 n_2}$,
+	and denote
+	\[
+	S = \left\{
+	\diag(a) \gamma^0 \diag(b),\quad
+	(a,b) \in \IR_{+*}^{n_1 + n_2}
+	\right\}
+	\]
+	the set of matrices that are diagonally similar to $\gamma^0$.
+	Then the set $S \cap \Ccal_1 \cap \Ccal_2$ contains exactly one element $\gamma^* = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0)$.
+\end{lemma}
+We refer to \cite{cuturi13} for the proof of this well-known result, which is used in the proof of the theorem.
+
+\begin{proof}[Proof of the theorem]
+	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^l)$ are thus diagonally similar to $\gamma^0$:
+	\[
+	\forall l\ge0,\quad \gamma^l \in S
+	\]
+
+	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^l))$ is non-increasing. Hence $(\gamma^l)$ is precompact.
+
+	If $\xi$ is a cluster point of $(\gamma^l)$, let us define
+	\begin{align*}
+	\tilde{\xi} &= P_{\Ccal_1}^{\theta_1(\xi)}(\xi) \\
+	\xi' &= P_{\Ccal_2}^{\theta_2(\tilde{\xi})}(\tilde{\xi}).
+	\end{align*}
+	Then by continuity of all the applications, $F(\xi) = F(\tilde{\xi}) = F(\xi')$.
+	From the hypothesis made on $\theta_1$ and $\theta_2$, can be deduced that $\xi$ is in $\Ccal_1$ and that $\tilde{\xi}$ is in $\Ccal_2$.
+Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
+	By lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
+\end{proof}
+
+%{\color{red} 
+%	TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
+To construct the functions $\theta_k$ needed by the theorem, we need the following lemma.
+
+\begin{lemma}\label{lemma:F_P_theta}
+	Let $1\le \theta < \omega$. Then, for any $\gamma \in \IR_{+*}^{nm}$, one has
+	\begin{equation}\label{eq:F_P_theta}
+	F(P^\theta_{\Ccal_k}(\gamma)) \le F(P^\omega_{\Ccal_k}(\gamma)).
+	\end{equation}
+	Moreover, equality occurs if and only if $\gamma \in \Ccal_k$.
+\end{lemma}
+\begin{proof}
+	Thanks to lemma \ref{lemma:lyapunov_decrease}, one knows that
+	\[
+	F(P^\theta_{\Ccal_k}(\gamma)) - F(P^\omega_{\Ccal_k}(\gamma))
+	= \scal{\mu^k}{(\varphi_\omega - \varphi_\theta) \left( \frac{A_k \gamma}{\mu^k} \right) } .
+	\]
+	Moreover, the function that maps $t \in [1,\infty)$ to $\varphi_t(x)$ is non-increasing:
+	$\frac{d}{dt} \varphi_t(x) = \log x (x^{1-t} - 1).$
+	For $x\neq 1$, it is even strictly decreasing.
+	Thus the inequality (\ref{eq:F_P_theta}) is valid, with equality \emph{iff} $A_k \gamma = \mu^k$.
+\end{proof}
+
+\begin{proposition}
+	We argue that a good choice for the functions $\theta_k$ may be constructed as follows. Pick a target parameter $\theta_0 \in [1;2)$, and a small $\delta>0$.
+	Define the functions $\Theta^*$ and $\Theta$ as:
+	\begin{align}
+	\label{eq:Theta_opt}
+	\Theta^*(u) &= \sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u\right) \ge 0 \right\} ,\\
+	\label{eq:Theta}
+	\Theta(u) &= \min(\max(1,\Theta^*(u)-\delta),\theta_0),
+	\end{align}
+	where $\min u$ denotes the smallest coordinate of the vector $u$.
+	Then the function
+	\begin{equation}
+	\label{eq:theta_k}
+	\theta_k(\gamma) =\Theta\left (\frac{A_k \gamma}{\mu^k}\right)
+	\end{equation}
+	is continuous, and verifies condition (\ref{eq:cond_theta_k}).
+\end{proposition}
+This construction has several advantages.
+One is that it allows to choose arbitrarily the parameter $\theta_0$ that will be used asymptotically. A heuristic for this choice is proposed in section \ref{section:local}.
+But it is also an easy approach to having an adaptive method, as the approximation of $\theta_k^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $\frac{A_k \gamma}{\mu^k}$, and the corresponding function to invert is decreasing and concave).
+We will show later on that this is often an excellent choice in practice.
+\begin{proof}
+	Since the partial derivative of $\frac{\partial\varphi_\omega(x)}{\partial\omega}$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
+	The function $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
+	Therefore, by lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
+	It is no problem to constrain it to $[1,\theta_0]$, as using 1 always verifies the inequality (\ref{eq:cond_theta_k}), and using $\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
+\end{proof}
+
+\section{Local rate of convergence ?}
+\label{section:local}
+Some material to adapt Lenaic results:
+
+The SK algorithm can be interpreted as an alternate maximisation algorithm on the dual of the regularized optimal transport problem.
+The dual problem of \eqref{ROT} is:
+\begin{equation}\label{DROT}\max_{\alpha,\beta}\langle \alpha,\mu^1\rangle+\langle \beta,\mu^2\rangle-\epsilon\sum_{i,j}e^{(\alpha_i+\beta_j-c_{i,j})/\epsilon}\end{equation}
+It is concave and continuously differentiable for all $(a,b)\in\IR^{n_1+n_2}$.
+Alternate maximization then converges and we recover, for $a_i=e^{\alpha_i/\epsilon}$, $b_j=e^{\beta_j/\epsilon}$ and $\gamma^0_{i,j}=e^{-c_{i,j}/\epsilon}$:
+\begin{align*}
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} 
+\end{align*}
+
+If $(\alpha^*,\beta^*)$ is a solution of \eqref{DROT}, then all the 
+other optimum can be written as $(\alpha^*+\kappa,\beta^*-\kappa)$, which means that the problem admits a unique solution, up to a translation with constant $\kappa\in\IR$.
+
+The Hessian matrix of the functional in \eqref{DROT} at point $(\alpha,\beta)$ reads, for $a_i=e^{\alpha_i/\epsilon}>0$ and $b_j=e^{\beta_j/\epsilon}>0$:
+
+$$H(a,b)=\frac1\epsilon\begin{bmatrix}
+\diag(a)\diag(\gamma^0 b)&\diag(a)\gamma^0 \diag(b)\\
+\diag(b){^t\gamma^0} \diag(a)&\diag(^t\gamma^0a)\diag(b)
+\end{bmatrix}.$$
+Notice that
+
+$$H(a^*,b^*)=\frac1\epsilon\begin{bmatrix}
+\diag(\mu^1)&\gamma^*\\
+^t\gamma^*&\diag(\mu^2)
+\end{bmatrix}.$$
+
+The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
+Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite for $a_i,b_j>0$.
+Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.
+All the other eigenvalues belongs to the set $[0;1[$.
+
+From experiments, it seems possible to design a good guess for $\omega$ from $c$ and $\epsilon$ only. 
+    Denoting as $\lambda$ the second larger eigenvalue of $T(1,1)$, one can set 
+    $\omega=\frac{2}{\lambda}(1-\sqrt{1-\lambda})$.
+
+\section{Experimental results}
+
+%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
+%The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
+
+\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
+%\begin{algorithm}
+%\caption{Over-relaxed SK algorithm}
+%\label{SOR}
+%\begin{algorithmic}
+%\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
+%\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
+%\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
+%\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
+%\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
+%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
+%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed SK iteration}
+%\ELSE
+%\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
+%\ENDIF
+%\ENDWHILE
+% \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
+%\end{algorithmic}
+%\end{algorithm}
+Add algorithm to solve \eqref{eq:Theta_opt}?
+
+This algorithm uses the function $\Theta$ defined in (\ref{eq:Theta}).
+
+\begin{algorithm}
+\caption{Over-relaxed SK algorithm }
+\label{SOR}
+\begin{algorithmic}
+\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
+\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\theta_0\in[1;2)$ and $\eta>0$
+\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
+\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
+\STATE $\omega=\Theta(a\oslash\tilde a)$
+%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
+\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$%\COMMENT{Over-relaxed SK iteration}
+%\ELSE
+%\STATE $a=\tilde a$,\COMMENT{SK iteration}
+%\ENDIF
+\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a)$
+\STATE $\omega=\Theta(b\oslash\tilde b)$
+%\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
+\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Over-relaxed SK iteration}
+%\ELSE
+%\STATE $b=\tilde b$,\COMMENT{SK iteration}
+%\ENDIF
+\ENDWHILE
+ \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
+\end{algorithmic}
+\end{algorithm}
+
+{\color{red} TODO}
+
+\section{Conclusion and perspectives}
+
+We have presented a generalization of the Sinkhorn--Knopp algorithm based on over-relaxed projections, with a criterion for global convergence based on a Lyapunov function.
+The proposed Lyapunov function allows for an adaptive choice of over-relaxation parameter at a low computational price.
+In practice, this algorithm converges way faster than Sinkhorn--Knopp; moreover, the acceleration rate is especially impressive for small regularization $\epsilon$.
+
+Contributions: 
+cheap test, preliminary work for convergence results. Huge acceleration in practice.
+
+Perspectives:
+global convergence, how chosing $\omega$?, intensive numerical study with respect to other accelerarions methods\cite{2016arXiv160604133S,2017arXiv170509634A}.
+
+Links with monotone Fista? (that choose Fista if energy decreased and forward backward otherwise...)
+
+\section*{Acknowledgments}
+This study has been carried out with financial support from the French State, managed by the French National Research Agency (ANR) in the frame of the  GOTMI project (ANR-16-CE33-0010-01).
+
+%\subsubsection*{References}
+
+\bibliographystyle{apalike}
+\bibliography{references}
+
+\end{document}

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 
-\title{Overrelaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport}
+\title{Over-relaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport}
 
@@
 It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn--Knopp algorithm).
-The idea is to overrelax the Bregman projection operators, allowing for faster convergence.
+The idea is to over-relax the Bregman projection operators, allowing for faster convergence.
 In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
 We propose a simple method for establishing global convergence by ensuring the decrease of a Lyapunov function at each step.
-An adaptive choice of overrelaxation parameter based on the Lyapunov function is constructed.
+An adaptive choice of over-relaxation parameter based on the Lyapunov function is constructed.
 We also suggest a heuristic to choose a suitable asymptotic over-relaxation parameter, based on a local convergence analysis.
@@
 As illustrated in Figure \ref{alternate_projections} (a-b), the SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
-The idea developped in this paper is to perform overrelaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
+The idea developed in this paper is to perform overrelaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
 
@@
 \caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
-In order to accelerate the convergence rate, over-relaxed projections are here considered (c).}
+In order to accelerate the convergence rate, overrelaxed projections are here considered (c).}
 \end{center}
@@
 This algorithm, while it may behave well in practice, has no convergence guarantee---it sometimes does not convergence even for reasonable values of $\omega$.
-Our goal in this section is to modify this algorithm to make it robust and garantee convergence.
+Our goal   is to modify this algorithm to make it robust and guarantee convergence.
 
@@
 They indeed correspond to a successive overelaxation (SOR) algorithm on the dual objective $E$.
-This can be seen by using the relationships gievn after equation~\eqref{DROT}:
+This can be seen by using the relationships given after equation~\eqref{DROT}:
 \begin{align}\label{SORdual}
@@
 
-{\color{blue} \begin{proposition} 
-State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
+%{\color{blue} \begin{proposition} 
+%State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
 lenaic: what is exactly needed?
 I can think of it tomorow.
-\end{proposition}}
+%\end{proposition}}
 \subsection{Lyapunov function}
@@
 	\]
-	Moreover, the function that maps $t \in [1,\infty)$ to $\varphi_t(x)$ is non-increasing:
+	The function that maps $t \in [1,\infty)$ to $\varphi_t(x)$ is non-increasing since
 	$\partial_t \varphi_t(x) =  (x^{1-t} - 1)\log x.$
 	For $x\neq 1$, it is even strictly decreasing.
-	Thus inequality \eqref{eq:F_P_theta} is valid, with equality \emph{iff} $A_k \gamma = \mu$.
+	Thus inequality \eqref{eq:F_P_theta} is valid, with equality \emph{iff} $A_k \gamma = \mu^k$.
 \end{proof}
@@
 \item it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
-\item it is also an easy approach to having an adaptive method, as the approximation of $\Theta_k^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
+\item it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
 \item experiments show that this is often an excellent choice in practice.
@@
 It uses the function $\Theta$ defined implicitely in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
-Note that with the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
+With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
 
@@
 \label{section:local}
-In order to theoretically jsutify the acceleration of convergence that is observed in practice, we now study the local convergence rate of the algorithm.
+In order to theoretically justify the acceleration of convergence that is observed in practice, we now study the local convergence rate of the algorithm.
 %We do not claim that this simple overrelaxation is the most efficient method to accelerate local convergence of fixed points---adaptive methods with excellent properties exist e.g.~\cite{2016arXiv160604133S}--- but this local analysis gives a hint for what makes our scheme relevant is the conjunction of our global convergence guarantee with this local acceleration property.
@@
 \begin{corollary}
-The precedding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
+The preceding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' over-relaxation parameter.
 \end{corollary}
-
 \begin{proof}
@@
 Of course, the latter is unrealistic strategy but our experiment shows that the acceleration power is not so much impacted by estimating once for all the overrelaxation parameter for a class of problem with fixed regularization, as shown on~\ref{fig:comparison}.
-
+\begin{figure}
+\begin{minipage}[b]{.5\linewidth}
+   \centering
+   \includegraphics[scale=0.6]{images/speedratio_image}
+                    \subcaption{Quadratic cost, random marginals}\label{fig:compare_a}
+\end{minipage}%
+			\begin{minipage}[b]{.5\linewidth}
+                    \centering
+			\includegraphics[scale=0.6]{images/speedratio_ML}
+                    \subcaption{Random cost, uniform marginals}\label{fig:compare_b}
+\end{minipage}
+			\caption{\label{fig:comparison}Speed ratio of SK algorithm and its accelerated version Algorithm~\ref{SOR} w.r.t parameter $\epsilon$.}% (stopping criterion $\Vert u - u^* \Vert_\infty<10^{-6})$
+\end{figure}
 Figure~\ref{fig:comparison} displays the ratio of the number of iterations required to reach a precision of $10^{-6}$ on the dual variable $\alpha$ for SK algorithm and Algorithm \ref{SOR}.
 It is is worth noting that the complexity per iteration is the same, modulo negligible terms, so this ratio is also the runtime ratio (our algorithm can also be parallelized on GPUs just as SK algorithm).
-
-%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
-%The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
-
-\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
+In both experimental settings, for low values of the regularization parameter $\epsilon$, the acceleration ratio is above $20$  with Algorithm \ref{SOR}.
+
+%\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
 %\begin{algorithm}
@@
 
-\begin{figure}
-\begin{minipage}[b]{.5\linewidth}
-   \centering
-   \includegraphics[scale=0.5]{images/speedratio_image}
-                    \subcaption{Quadratic cost, random marginals}\label{fig:compare_a}
-\end{minipage}%
-			\begin{minipage}[b]{.5\linewidth}
-                    \centering
-			\includegraphics[scale=0.5]{images/speedratio_ML}
-                    \subcaption{Random cost, uniform marginals}\label{fig:compare_b}
-\end{minipage}
-			\caption{\label{fig:comparison}Complexity ratio of SK algorithm and its accelerated version Algorithm~\ref{SOR}.}% (stopping criterion $\Vert u - u^* \Vert_\infty<10^{-6})$
-\end{figure}
-
 \section{Conclusion and perspectives}
@@
 4]{chizat2017these} but there is no systematic way to derive globally convergent algorithms.
-Our work as an effort in the direction of building and understanding the properties of robust first order algorithms for solving OT.
+Our work is an effort in the direction of building and understanding the properties of robust first order algorithms for solving OT.
 More understanding is needed regarding SOR itself (global convergence speed, choice of $\theta_0$), but also its relation to other acceleration methods~\cite{2016arXiv160604133S,2017arXiv170509634A}.

### 2017-10-20 paper.tex
Relecture Alexis

--- 2017-10-19 paper.tex
@@
 An adaptive choice of overrelaxation parameter based on the Lyapunov function is constructed.
-We also suggest a heuristic to choose a suitable asymptotic over-relaxation parameter, based on a local convergence analysis.
+We also suggest a heuristic to choose a suitable asymptotic overrelaxation parameter, based on a local convergence analysis.
 Our numerical experiments show a gain in convergence speed by an order of magnitude in certain regimes.
@@
 The computation of optimal transport between two data relies on the estimation of an optimal transport matrix, the entries of which represent the quantity of mass transported between  data locations. 
-Regularization of optimal transport with strictly convex regularizers \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
+Regularization of optimal transport with strictly convex regularization \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
 Hence, for particular purposes such as color interpolation \cite{Rabin2014} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
-In this setting,  the regularized transport problem can be ill-conditioned and Sinkhorn's algorithm converges slowly. This is the issue  we want to tackle here.
+In this setting,  the regularized transport problem can be ill-conditioned and the Sinkhorn--Knopp algorithm converges slowly. This is the issue  we want to tackle here.
 Before going into further details, we now briefly introduce the main notations and concepts used all along this article.
@@
 with the convention $0 \log 0= 0$.
-It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ (here and in the sequel, exponentiation is meant entrywise) onto $\Ccal_1 \cap \Ccal_2$:
+It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ (here and in the sequel, exponentiation is meant entry-wise) onto $\Ccal_1 \cap \Ccal_2$:
 \begin{equation}\label{eq:reg_ot_pb}
@@
 \gamma^0 &= e^{-c/\epsilon} &
-\gamma^{l+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l)),
+\gamma^{\ell+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^{\ell})),
 \end{align*}
 and we have 
-$\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
+$\lim_{l\rightarrow +\infty} \gamma^{\ell} = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
 %
@@
 where $\oslash$ is the pointwise division. 
-To compute numerically the solution one simply has to store $(a^l, b^l)\in\IR^{n_1}\times \IR^{n_2}$ and to iterate
+To compute numerically the solution one simply has to store $(a^{\ell}, b^{\ell})\in\IR^{n_1}\times \IR^{n_2}$ and to iterate
 \begin{align*}
-a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
-b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} .
+a^{\ell+1} &= {\mu^1}\oslash{\gamma^0 b^{\ell}} &
+b^{\ell+1} &= {\mu^2}\oslash{^t \gamma^0 a^{\ell+1}} .
 \end{align*}
-We then have $\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$
+We then have $\gamma^{\ell} = \diag(a^{\ell}) \gamma^0 \diag(b^{\ell}).$
 
@@
 \begin{align*}
-a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
-b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} 
+a^{\ell+1} &= {\mu^1}\oslash{\gamma^0 b^{\ell}} &
+b^{\ell+1} &= {\mu^2}\oslash{^t \gamma^0 a^{\ell+1}} 
 \end{align*}
@@
 \end{minipage}%
-\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
-Overrelaxed projection typically accelerate the convergence rate.}
+\caption{\label{alternate_projections} The trajectory of $\gamma^{\ell}$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
+Overrelaxed projections (c) typically accelerate the convergence rate.}
 \end{figure}
-
-\iffalse
-\begin{figure}[ht!]
-\begin{center}
-\begin{tabular}{ccc}
-\input{schema_a.tex}&
-\input{schema_b.tex}&
-\input{schema_c.tex}\\
-(a)&(b)&(c)
-\end{tabular}
-\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
-In order to accelerate the convergence rate, over-relaxed projections are here considered (c).}
-\end{center}
-\end{figure}
-\fi
 
@@
 A naive algorithm would then consist in iteratively applying $P^\omega_{\Ccal_2}\circ P^\omega_{\Ccal_1}$ for some choice of $\omega$.
-This algorithm, while it may behave well in practice, has no convergence guarantee---it sometimes does not convergence even for reasonable values of $\omega$.
-Our goal in this section is to modify this algorithm to make it robust and garantee convergence.
+While it often behaves well in practice, this algorithm may sometimes not converge even for reasonable values of $\omega$.
+Our goal in this section is to modify this algorithm to make it robust and to guarantee convergence.
 
@@
 
-{\color{blue} \begin{proposition} 
-State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
-lenaic: what is exactly needed?
-I can think of it tomorow.
-\end{proposition}}
 \subsection{Lyapunov function}
-Convergence of the successive overrelaxed projection cannot be guaranteed in general. In order to derive a robust algorithm with provable convergence, we introduce the Lyapunov function 
+Convergence of the successive overrelaxed projections is not guaranteed in general. In order to derive a robust algorithm with provable convergence, we introduce the Lyapunov function 
 \begin{equation}\label{eq:lyapunov_function}
@@
 where $\gamma^*$ denotes the solution of the regularized OT problem.
-We will use this function to enforce the strict descent criterion $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+We will use this function to enforce the strict descent criterion $F(\gamma^{\ell+1}) < F(\gamma^\ell)$ as long as the process has not converged.
 %
@@
 
-\iffalse
-\begin{proof}
-	First note that K is closed in $\IR_{+*}^{n_1 n_2}$, by continuity of $F$.
-	For $\gamma \in K$, one has:
-	\begin{align*}
-	F(\gamma) &= \scal{\gamma^*}{\log \gamma^* - 1} + \scal{\gamma}{1} - \scal{\gamma^*}{\log \gamma}\\
-	\sum_{i,j} \gamma_{i,j} - \gamma^*_{i,j} \log \gamma_{i,j} &\le M - \scal{\gamma^*}{\log \gamma^* - 1} .
-	\end{align*}
-	By concavity, the logarithm admits a simple upper bound: $\log \gamma_{i,j} \le \frac{\gamma_{i,j}}{e}$. Moreover, since $\gamma^*$ is a solution of the problem, each of its coordinates is at most 1.
-	\begin{align*}
-	\sum_{i,j} \gamma_{i,j} &\le \frac{M - \scal{\gamma^*}{\log \gamma^*-1}}{1-\frac{1}{e}} =: M'.
-	\end{align*}
-	Therefore each coordinate $\gamma_{i,j}$ is bounded above by $M'$.
-
-	Let us show that the coordinates are bounded below by a strictly positive number; from which it can be deduced that $K$ is bounded and closed in $\IR^{n_1 n_2}$, and therefore compact.
-	\begin{align*}
-	F(\gamma)
-	&\ge \scal{\gamma^*}{\log \gamma^*} - \sum_{i,j} \gamma^*_{i,j} \log(\gamma_{i,j}) - 1\\
-	&\ge \scal{\gamma^*}{\log \gamma^*} - \gamma^*_{i_0,j_0} \log(\gamma_{i_0,j_0}) - 1 - \log M',
-	\end{align*}
-	where the last equation derives from $\log(y_{i,j}) \le \log M'$, and $\sum_{(i,j)\neq(i_0,j_0)} \gamma^*_{i,j} \le 1$.
-	Finally:
-	\[
-	\log(\gamma_{i_0,j_0}) \ge \frac{\scal{\gamma^*}{\log \gamma^*} - M - \log M' - 1}{\gamma^*_{i_0,j_0}},
-	\]
-	and $K$ is thus compact.
-\end{proof}
-\fi
-
 %Note that the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$ may be calculated without knowing $\gamma^*$, as shown by the following lemma.
@@
 	\begin{align*}
-	\tilde{\gamma}^{l+1} &= P_{\Ccal_1}^{\theta_1(\gamma^l)}(\gamma^l) \\
-	\gamma^{l+1} &= P_{\Ccal_2}^{\theta_2(\tilde{\gamma}^{l+1})}(\tilde{\gamma}^{l+1}).
+	\tilde{\gamma}^{\ell+1} &= P_{\Ccal_1}^{\theta_1(\gamma^{\ell})}(\gamma^{\ell}) \\
+	\gamma^{\ell+1} &= P_{\Ccal_2}^{\theta_2(\tilde{\gamma}^{\ell+1})}(\tilde{\gamma}^{\ell+1}).
 	\end{align*}
-	Then the sequence $(\gamma^l)$ converges to $\gamma^*$.
+	Then the sequence $(\gamma^{\ell})$ converges to $\gamma^*$.
 \end{theorem}
 
-\begin{lemma}[\cite{cuturi13}]
+\begin{lemma}[]
 	\label{lemma:trivial_intersection}
@@
 \end{lemma}
+\begin{proof}
+	We refer to \cite{cuturi13} for a proof of this lemma.
+\end{proof}
 
 \begin{proof}[Proof of the theorem]
-	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^l)$ are thus diagonally similar to $\gamma^0$:
+	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^{\ell})$ are thus diagonally similar to $\gamma^0$:
 	\[
-	\forall l\ge0,\quad \gamma^l \in S
+	\forall \ell\ge0,\quad \gamma^{\ell} \in S
 	\]
-	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^l))$ is non-increasing. Hence $(\gamma^l)$ is precompact.
+	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^{\ell}))$ is non-increasing. Hence $(\gamma^{\ell})$ is precompact.
 	%
-	If $\xi$ is a cluster point of $(\gamma^l)$, let us define
+	If $\xi$ is a cluster point of $(\gamma^{\ell})$, let us define
 	\begin{align*}
@@
 Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
-	By Lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
+	By Lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^{\ell})$ converges to the solution.
 \end{proof}
@@
 \begin{lemma}\label{lemma:F_P_theta}
-	Let $1\le \theta < \omega$. Then, for any $\gamma \in \IR_{+*}^{nm}$, one has
+	Let $1\le \theta < \omega$. Then, for any $\gamma \in \IR_{+*}^{n_1 n_2}$, one has
 	\begin{equation}\label{eq:F_P_theta}
@@
 	\]
-	Moreover, the function that maps $t \in [1,\infty)$ to $\varphi_t(x)$ is non-increasing:
+	The function that maps $t \in [1,\infty)$ to $\varphi_t(x)$ is non-increasing since
 	$\partial_t \varphi_t(x) =  (x^{1-t} - 1)\log x.$
@@
 \begin{proof}
-Looking at Figure~\ref{phi_omega} can help understanding this proof.	Since the partial derivative of $\partial_\omega \varphi_\omega(x)$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
+Looking at Figure~\ref{phi_omega} can help understand this proof.	Since the partial derivative of $\partial_\omega \varphi_\omega(x)$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
 	The function $\Theta^*\left((A_k \gamma)\oslash \mu^k)\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
 	Therefore, by lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
-	Constraining the parameter to $[1,\theta_0]$ preserves this properties.% as using $\theta=1$ always satisfies inequality \eqref{eq:cond_theta_k} and using $\theta=\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
+	Constraining the parameter to $[1,\theta_0]$ preserves this property.% as using $\theta=1$ always satisfies inequality \eqref{eq:cond_theta_k} and using $\theta=\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
 \end{proof}
@@
 The resulting algorithm, which is proved to be convergent by Theorem~\ref{thm:algo}, is written in pseudo-code in Algorithm \ref{SOR}.
-It uses the function $\Theta$ defined implicitely in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
+It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
 Note that with the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
@@
 \STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\theta_0\in[1;2)$ and $\eta>0$
-\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
+\WHILE {$||a\otimes \gamma^0 b -  \mu_1||>\eta$}
 \STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
@@
 %\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
-\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$%\COMMENT{Over-relaxed SK iteration}
+\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$
 %\ELSE
@@
 %\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
-\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Over-relaxed SK iteration}
+\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Overrelaxed SK iteration}
 %\ELSE
@@
 \begin{corollary}
-The precedding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
+The preceding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
 \end{corollary}
@@
 \begin{proof}
-For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual} and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
+For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual} and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0 e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
 Notice that the dual problem~\eqref{DROT} has an invariant subspace due to the invariance by translation $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is strictly convex in any subspace which does not contain this line.
@@
 
-\iffalse
-The Hessian matrix of the functional in \eqref{DROT} at point $(\alpha,\beta)$ reads, for $a_i=e^{\alpha_i/\epsilon}>0$ and $b_j=e^{\beta_j/\epsilon}>0$:
-
-$$H(a,b)=\frac1\epsilon\begin{bmatrix}
-\diag(a)\diag(\gamma^0 b)&\diag(a)\gamma^0 \diag(b)\\
-\diag(b){^t\gamma^0} \diag(a)&\diag(^t\gamma^0a)\diag(b)
-\end{bmatrix}.$$
-Notice that
-
-$$H(a^*,b^*)=\frac1\epsilon\begin{bmatrix}
-\diag(\mu^1)&\gamma^*\\
-^t\gamma^*&\diag(\mu^2)
-\end{bmatrix}.$$
-
-The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
-Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite for $a_i,b_j>0$.
-Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.
-All the other eigenvalues belongs to the set $[0;1[$.
-
-From experiments, it seems possible to design a good guess for $\omega$ from $c$ and $\epsilon$ only. 
-    Denoting as $\lambda$ the second larger eigenvalue of $T(1,1)$, one can set 
-    $\omega=\frac{2}{\lambda}(1-\sqrt{1-\lambda})$.
-
-    \fi
-
 \section{Experimental results}
@@
 The marginals are densities made of the sum of a base plateau of height $0.1$ and another plateau of heights and boundaries chosen uniformly in $[0,1]$, subsequently normalized.
-In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.%The interest of having two settings is that the SK algorithm is known to behave very differently as $\epsilon$ decreases for these two problems and
+In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.
+%The interest of having two settings is that the SK algorithm is known to behave very differently as $\epsilon$ decreases for these two problems and
 
@@
 Figure~\ref{fig:comparison} displays the ratio of the number of iterations required to reach a precision of $10^{-6}$ on the dual variable $\alpha$ for SK algorithm and Algorithm \ref{SOR}.
-It is is worth noting that the complexity per iteration is the same, modulo negligible terms, so this ratio is also the runtime ratio (our algorithm can also be parallelized on GPUs just as SK algorithm).
+It is is worth noting that the complexity per iteration is the same, modulo negligible terms, so this ratio is also the runtime ratio.
+%(our algorithm can also be parallelized on GPUs just as SK algorithm).
 
-%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
+%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^{\ell})))-F(\gamma^{\ell})>0$ for performing the overrelaxed iterations with parameter $\omega$.
 %The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
@@
 %\begin{algorithm}
-%\caption{Over-relaxed SK algorithm}
+%\caption{Overrelaxed SK algorithm}
 %\label{SOR}
@@
 %\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
-%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed SK iteration}
+%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Overrelaxed SK iteration}
 %\ELSE
@@
 \section{Conclusion and perspectives}
-SK algorithm is widely used to solve entropy regularized OT.
-The use of overrelaxed projections turns out to be a natural and simple idea to accelerate convergence while keeping many nice properties of this algorithm (first order, paralelizable, simple).
+The SK algorithm is widely used to solve entropy regularized OT.
+The use of overrelaxed projections turns out to be a natural and simple idea to accelerate convergence while keeping many nice properties of this algorithm (first order, parallelizable, simple).
 We have proposed an algorithm that adaptively chooses the overrelaxation parameter so as to guarantee global convergence.
@@
 
-This idea of overrelaxation can be generalized to solve more general problems such as barycenters, gradient flows, unbalanced OT~\cite[chap.
+This idea of overrelaxation can be generalized to solve more general problems such as multi-marginal OT, barycenters, gradient flows, unbalanced OT~\cite[chap.
 4]{chizat2017these} but there is no systematic way to derive globally convergent algorithms.
-Our work as an effort in the direction of building and understanding the properties of robust first order algorithms for solving OT.
+Our work is an effort in the direction of building and understanding the properties of robust first order algorithms for solving OT.
 More understanding is needed regarding SOR itself (global convergence speed, choice of $\theta_0$), but also its relation to other acceleration methods~\cite{2016arXiv160604133S,2017arXiv170509634A}.

### 2017-10-20 paper.tex

@@
 It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn--Knopp algorithm).
-The idea is to overrelax the Bregman projection operators, allowing for faster convergence.
+The idea is to over-relax the Bregman projection operators, allowing for faster convergence.
 In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
@@
 
+%{\color{blue} \begin{proposition} 
+%State here that the fixed points of   $\gamma^{l+1} =P^\theta_{\Ccal_2}(P^\theta_{\Ccal_1}(\gamma^l))$ are the same than $\gamma^{l+1} =P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l))$?
+lenaic: what is exactly needed?
+I can think of it tomorow.
+%\end{proposition}}
 \subsection{Lyapunov function}
@@
 	For $x\neq 1$, it is even strictly decreasing.
-	Thus inequality \eqref{eq:F_P_theta} is valid, with equality \emph{iff} $A_k \gamma = \mu$.
+	Thus inequality \eqref{eq:F_P_theta} is valid, with equality \emph{iff} $A_k \gamma = \mu^k$.
 \end{proof}
@@
 \end{corollary}
-
 \begin{proof}
@@
 Of course, the latter is unrealistic strategy but our experiment shows that the acceleration power is not so much impacted by estimating once for all the overrelaxation parameter for a class of problem with fixed regularization, as shown on~\ref{fig:comparison}.
+\begin{figure}
+\begin{minipage}[b]{.5\linewidth}
+   \centering
+   \includegraphics[scale=0.6]{images/speedratio_image}
+                    \subcaption{Quadratic cost, random marginals}\label{fig:compare_a}
+\end{minipage}%
+			\begin{minipage}[b]{.5\linewidth}
+                    \centering
+			\includegraphics[scale=0.6]{images/speedratio_ML}
+                    \subcaption{Random cost, uniform marginals}\label{fig:compare_b}
+\end{minipage}
+			\caption{\label{fig:comparison}Speed ratio of SK algorithm and its accelerated version Algorithm~\ref{SOR} w.r.t parameter $\epsilon$.}% (stopping criterion $\Vert u - u^* \Vert_\infty<10^{-6})$
+\end{figure}
+Figure~\ref{fig:comparison} displays the ratio of the number of iterations required to reach a precision of $10^{-6}$ on the dual variable $\alpha$ for SK algorithm and Algorithm \ref{SOR}.
+It is is worth noting that the complexity per iteration is the same, modulo negligible terms, so this ratio is also the runtime ratio (our algorithm can also be parallelized on GPUs just as SK algorithm).
+In both experimental settings, for low values of the regularization parameter $\epsilon$, the acceleration ratio is above $20$  with Algorithm \ref{SOR}.
 
-Figure~\ref{fig:comparison} displays the ratio of the number of iterations required to reach a precision of $10^{-6}$ on the dual variable $\alpha$ for SK algorithm and Algorithm \ref{SOR}.
-It is is worth noting that the complexity per iteration is the same, modulo negligible terms, so this ratio is also the runtime ratio.
-%(our algorithm can also be parallelized on GPUs just as SK algorithm).
-
-%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^{\ell})))-F(\gamma^{\ell})>0$ for performing the overrelaxed iterations with parameter $\omega$.
-%The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).
-
-\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
+%\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
 %\begin{algorithm}
@@
 %Add algorithm to solve \eqref{eq:Theta_opt}?
-
-\begin{figure}
-\begin{minipage}[b]{.5\linewidth}
-   \centering
-   \includegraphics[scale=0.5]{images/speedratio_image}
-                    \subcaption{Quadratic cost, random marginals}\label{fig:compare_a}
-\end{minipage}%
-			\begin{minipage}[b]{.5\linewidth}
-                    \centering
-			\includegraphics[scale=0.5]{images/speedratio_ML}
-                    \subcaption{Random cost, uniform marginals}\label{fig:compare_b}
-\end{minipage}
-			\caption{\label{fig:comparison}Complexity ratio of SK algorithm and its accelerated version Algorithm~\ref{SOR}.}% (stopping criterion $\Vert u - u^* \Vert_\infty<10^{-6})$
-\end{figure}
 

### 2017-10-20 paper.tex

@@
 \begin{proof}
-For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual} and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0 e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
+For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual}  and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
 Notice that the dual problem~\eqref{DROT} has an invariant subspace due to the invariance by translation $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is strictly convex in any subspace which does not contain this line.
@@
 
-The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative}, even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.
+The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative} , even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.
 

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 We propose a simple method for establishing global convergence by ensuring the decrease of a Lyapunov function at each step.
-An adaptive choice of overrelaxation parameter based on the Lyapunov function is constructed.
-We also suggest a heuristic to choose a suitable asymptotic overrelaxation parameter, based on a local convergence analysis.
+An adaptive choice of over-relaxation parameter based on the Lyapunov function is constructed.
+We also suggest a heuristic to choose a suitable asymptotic over-relaxation parameter, based on a local convergence analysis.
 Our numerical experiments show a gain in convergence speed by an order of magnitude in certain regimes.
@@
 The computation of optimal transport between two data relies on the estimation of an optimal transport matrix, the entries of which represent the quantity of mass transported between  data locations. 
-Regularization of optimal transport with strictly convex regularization \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
+Regularization of optimal transport with strictly convex regularizers \cite{cuturi13, dessein2016}  nevertheless involves a spreading of the mass.
 Hence, for particular purposes such as color interpolation \cite{Rabin2014} or gradient flow \cite{2016arXiv160705816C}, it is  necessary  to consider very small regularization of the problem.
-In this setting,  the regularized transport problem can be ill-conditioned and the Sinkhorn--Knopp algorithm converges slowly. This is the issue  we want to tackle here.
+In this setting,  the regularized transport problem can be ill-conditioned and Sinkhorn's algorithm converges slowly. This is the issue  we want to tackle here.
 Before going into further details, we now briefly introduce the main notations and concepts used all along this article.
@@
 with the convention $0 \log 0= 0$.
-It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ (here and in the sequel, exponentiation is meant entry-wise) onto $\Ccal_1 \cap \Ccal_2$:
+It was shown in \cite{benamou15}  that the regularized optimal transport matrix $\gamma^*$, which is the unique minimizer of problem \eqref{ROT},  is the Bregman projection of $\gamma^0 = e^{-c/\epsilon}$ (here and in the sequel, exponentiation is meant entrywise) onto $\Ccal_1 \cap \Ccal_2$:
 \begin{equation}\label{eq:reg_ot_pb}
@@
 \gamma^0 &= e^{-c/\epsilon} &
-\gamma^{\ell+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^{\ell})),
+\gamma^{l+1} = P_{\Ccal_2}(P_{\Ccal_1}(\gamma^l)),
 \end{align*}
 and we have 
-$\lim_{l\rightarrow +\infty} \gamma^{\ell} = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
+$\lim_{l\rightarrow +\infty} \gamma^l = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0) = \gamma^*.$
 %
@@
 where $\oslash$ is the pointwise division. 
-To compute numerically the solution one simply has to store $(a^{\ell}, b^{\ell})\in\IR^{n_1}\times \IR^{n_2}$ and to iterate
+To compute numerically the solution one simply has to store $(a^l, b^l)\in\IR^{n_1}\times \IR^{n_2}$ and to iterate
 \begin{align*}
-a^{\ell+1} &= {\mu^1}\oslash{\gamma^0 b^{\ell}} &
-b^{\ell+1} &= {\mu^2}\oslash{^t \gamma^0 a^{\ell+1}} .
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} .
 \end{align*}
-We then have $\gamma^{\ell} = \diag(a^{\ell}) \gamma^0 \diag(b^{\ell}).$
+We then have $\gamma^l = \diag(a^l) \gamma^0 \diag(b^l).$
 
@@
 \begin{align*}
-a^{\ell+1} &= {\mu^1}\oslash{\gamma^0 b^{\ell}} &
-b^{\ell+1} &= {\mu^2}\oslash{^t \gamma^0 a^{\ell+1}} 
+a^{l+1} &= {\mu^1}\oslash{\gamma^0 b^l} &
+b^{l+1} &= {\mu^2}\oslash{^t \gamma^0 a^{l+1}} 
 \end{align*}
@@
 \end{minipage}%
-\caption{\label{alternate_projections} The trajectory of $\gamma^{\ell}$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
-Overrelaxed projections (c) typically accelerate the convergence rate.}
+\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
+Overrelaxed projection typically accelerate the convergence rate.}
 \end{figure}
+
+\iffalse
+\begin{figure}[ht!]
+\begin{center}
+\begin{tabular}{ccc}
+\input{schema_a.tex}&
+\input{schema_b.tex}&
+\input{schema_c.tex}\\
+(a)&(b)&(c)
+\end{tabular}
+\caption{\label{alternate_projections} The trajectory of $\gamma^l$ given by the SK algorithm is illustrated for decreasing values of $\epsilon$ in (a) and (b).
+In order to accelerate the convergence rate, overrelaxed projections are here considered (c).}
+\end{center}
+\end{figure}
+\fi
 
@@
 A naive algorithm would then consist in iteratively applying $P^\omega_{\Ccal_2}\circ P^\omega_{\Ccal_1}$ for some choice of $\omega$.
-While it often behaves well in practice, this algorithm may sometimes not converge even for reasonable values of $\omega$.
-Our goal in this section is to modify this algorithm to make it robust and to guarantee convergence.
+This algorithm, while it may behave well in practice, has no convergence guarantee---it sometimes does not convergence even for reasonable values of $\omega$.
+Our goal   is to modify this algorithm to make it robust and guarantee convergence.
 
@@
 \subsection{Lyapunov function}
-Convergence of the successive overrelaxed projections is not guaranteed in general. In order to derive a robust algorithm with provable convergence, we introduce the Lyapunov function 
+Convergence of the successive overrelaxed projection cannot be guaranteed in general. In order to derive a robust algorithm with provable convergence, we introduce the Lyapunov function 
 \begin{equation}\label{eq:lyapunov_function}
@@
 where $\gamma^*$ denotes the solution of the regularized OT problem.
-We will use this function to enforce the strict descent criterion $F(\gamma^{\ell+1}) < F(\gamma^\ell)$ as long as the process has not converged.
+We will use this function to enforce the strict descent criterion $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
 %
@@
 
+\iffalse
+\begin{proof}
+	First note that K is closed in $\IR_{+*}^{n_1 n_2}$, by continuity of $F$.
+	For $\gamma \in K$, one has:
+	\begin{align*}
+	F(\gamma) &= \scal{\gamma^*}{\log \gamma^* - 1} + \scal{\gamma}{1} - \scal{\gamma^*}{\log \gamma}\\
+	\sum_{i,j} \gamma_{i,j} - \gamma^*_{i,j} \log \gamma_{i,j} &\le M - \scal{\gamma^*}{\log \gamma^* - 1} .
+	\end{align*}
+	By concavity, the logarithm admits a simple upper bound: $\log \gamma_{i,j} \le \frac{\gamma_{i,j}}{e}$. Moreover, since $\gamma^*$ is a solution of the problem, each of its coordinates is at most 1.
+	\begin{align*}
+	\sum_{i,j} \gamma_{i,j} &\le \frac{M - \scal{\gamma^*}{\log \gamma^*-1}}{1-\frac{1}{e}} =: M'.
+	\end{align*}
+	Therefore each coordinate $\gamma_{i,j}$ is bounded above by $M'$.
+
+	Let us show that the coordinates are bounded below by a strictly positive number; from which it can be deduced that $K$ is bounded and closed in $\IR^{n_1 n_2}$, and therefore compact.
+	\begin{align*}
+	F(\gamma)
+	&\ge \scal{\gamma^*}{\log \gamma^*} - \sum_{i,j} \gamma^*_{i,j} \log(\gamma_{i,j}) - 1\\
+	&\ge \scal{\gamma^*}{\log \gamma^*} - \gamma^*_{i_0,j_0} \log(\gamma_{i_0,j_0}) - 1 - \log M',
+	\end{align*}
+	where the last equation derives from $\log(y_{i,j}) \le \log M'$, and $\sum_{(i,j)\neq(i_0,j_0)} \gamma^*_{i,j} \le 1$.
+	Finally:
+	\[
+	\log(\gamma_{i_0,j_0}) \ge \frac{\scal{\gamma^*}{\log \gamma^*} - M - \log M' - 1}{\gamma^*_{i_0,j_0}},
+	\]
+	and $K$ is thus compact.
+\end{proof}
+\fi
+
 %Note that the difference $F(\gamma) - F(P^\omega_{\Ccal_k}(\gamma))$ may be calculated without knowing $\gamma^*$, as shown by the following lemma.
@@
 	\begin{align*}
-	\tilde{\gamma}^{\ell+1} &= P_{\Ccal_1}^{\theta_1(\gamma^{\ell})}(\gamma^{\ell}) \\
-	\gamma^{\ell+1} &= P_{\Ccal_2}^{\theta_2(\tilde{\gamma}^{\ell+1})}(\tilde{\gamma}^{\ell+1}).
+	\tilde{\gamma}^{l+1} &= P_{\Ccal_1}^{\theta_1(\gamma^l)}(\gamma^l) \\
+	\gamma^{l+1} &= P_{\Ccal_2}^{\theta_2(\tilde{\gamma}^{l+1})}(\tilde{\gamma}^{l+1}).
 	\end{align*}
-	Then the sequence $(\gamma^{\ell})$ converges to $\gamma^*$.
+	Then the sequence $(\gamma^l)$ converges to $\gamma^*$.
 \end{theorem}
 
-\begin{lemma}[]
+\begin{lemma}[\cite{cuturi13}]
 	\label{lemma:trivial_intersection}
@@
 \end{lemma}
-\begin{proof}
-	We refer to \cite{cuturi13} for a proof of this lemma.
-\end{proof}
 
 \begin{proof}[Proof of the theorem]
-	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^{\ell})$ are thus diagonally similar to $\gamma^0$:
+	First of all, notice that the operators $P_{\Ccal_k}^\theta$ apply a scaling to lines or columns of matrices. All $(\gamma^l)$ are thus diagonally similar to $\gamma^0$:
 	\[
-	\forall \ell\ge0,\quad \gamma^{\ell} \in S
+	\forall l\ge0,\quad \gamma^l \in S
 	\]
-	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^{\ell}))$ is non-increasing. Hence $(\gamma^{\ell})$ is precompact.
+	By construction of the functions $\theta_k$, the sequence of values of the Lyapunov function $(F(\gamma^l))$ is non-increasing. Hence $(\gamma^l)$ is precompact.
 	%
-	If $\xi$ is a cluster point of $(\gamma^{\ell})$, let us define
+	If $\xi$ is a cluster point of $(\gamma^l)$, let us define
 	\begin{align*}
@@
 Therefore $\xi' = \tilde{\xi} = \xi$ is in the intersection $\Ccal_1 \cap \Ccal_2$.
-	By Lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^{\ell})$ converges to the solution.
+	By Lemma \ref{lemma:trivial_intersection}, $\xi = \gamma^*$, and the whole sequence $(\gamma^l)$ converges to the solution.
 \end{proof}
@@
 \begin{lemma}\label{lemma:F_P_theta}
-	Let $1\le \theta < \omega$. Then, for any $\gamma \in \IR_{+*}^{n_1 n_2}$, one has
+	Let $1\le \theta < \omega$. Then, for any $\gamma \in \IR_{+*}^{nm}$, one has
 	\begin{equation}\label{eq:F_P_theta}
@@
 \begin{proof}
-Looking at Figure~\ref{phi_omega} can help understand this proof.	Since the partial derivative of $\partial_\omega \varphi_\omega(x)$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
+Looking at Figure~\ref{phi_omega} can help understanding this proof.	Since the partial derivative of $\partial_\omega \varphi_\omega(x)$ is nonzero for any $x<1$, the implicit function theorem proves the continuity of $\Theta^*$.
 	The function $\Theta^*\left((A_k \gamma)\oslash \mu^k)\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
 	Therefore, by lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
-	Constraining the parameter to $[1,\theta_0]$ preserves this property.% as using $\theta=1$ always satisfies inequality \eqref{eq:cond_theta_k} and using $\theta=\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
+	Constraining the parameter to $[1,\theta_0]$ preserves this properties.% as using $\theta=1$ always satisfies inequality \eqref{eq:cond_theta_k} and using $\theta=\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
 \end{proof}
@@
 The resulting algorithm, which is proved to be convergent by Theorem~\ref{thm:algo}, is written in pseudo-code in Algorithm \ref{SOR}.
-It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
-Note that with the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
+It uses the function $\Theta$ defined implicitely in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
+With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
 
@@
 \STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\theta_0\in[1;2)$ and $\eta>0$
-\WHILE {$||a\otimes \gamma^0 b -  \mu_1||>\eta$}
+\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
 \STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
@@
 %\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
-\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$
+\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$%\COMMENT{Over-relaxed SK iteration}
 %\ELSE
@@
 %\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
-\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Overrelaxed SK iteration}
+\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Over-relaxed SK iteration}
 %\ELSE
@@
 \begin{corollary}
-The preceding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
+The preceding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' over-relaxation parameter.
 \end{corollary}
@@
 
+\iffalse
+The Hessian matrix of the functional in \eqref{DROT} at point $(\alpha,\beta)$ reads, for $a_i=e^{\alpha_i/\epsilon}>0$ and $b_j=e^{\beta_j/\epsilon}>0$:
+
+$$H(a,b)=\frac1\epsilon\begin{bmatrix}
+\diag(a)\diag(\gamma^0 b)&\diag(a)\gamma^0 \diag(b)\\
+\diag(b){^t\gamma^0} \diag(a)&\diag(^t\gamma^0a)\diag(b)
+\end{bmatrix}.$$
+Notice that
+
+$$H(a^*,b^*)=\frac1\epsilon\begin{bmatrix}
+\diag(\mu^1)&\gamma^*\\
+^t\gamma^*&\diag(\mu^2)
+\end{bmatrix}.$$
+
+The matrices $H(a,b)$  are positive semi-definite. They have only one $0$ eigenvalue associated to the eigenvetor $^t[\mathbf{1},-\mathbf{1}]$. 
+Using Schur complement properties, one can then show that any matrix $T(a,b)=(\diag(a)\diag(\gamma^0 b))^{-1}\diag(a)\gamma^0\diag(b)(\diag(^t\gamma^0a)\diag(b))^{-1} \diag(b){^t\gamma^0}\diag(a)=(\diag(\gamma^0 b))^{-1}\gamma^0(\diag(^t\gamma^0a))^{-1} \diag(b){^t\gamma^0}\diag(a)$ is positive semi definite for $a_i,b_j>0$.
+Moreover, we can show that $T$ has a unique largest eigenvalue that is $1$ and is associated to the eigenvector $\mathbf{1}$.
+All the other eigenvalues belongs to the set $[0;1[$.
+
+From experiments, it seems possible to design a good guess for $\omega$ from $c$ and $\epsilon$ only. 
+    Denoting as $\lambda$ the second larger eigenvalue of $T(1,1)$, one can set 
+    $\omega=\frac{2}{\lambda}(1-\sqrt{1-\lambda})$.
+
+    \fi
+
 \section{Experimental results}
@@
 The marginals are densities made of the sum of a base plateau of height $0.1$ and another plateau of heights and boundaries chosen uniformly in $[0,1]$, subsequently normalized.
-In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.
-%The interest of having two settings is that the SK algorithm is known to behave very differently as $\epsilon$ decreases for these two problems and
+In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.%The interest of having two settings is that the SK algorithm is known to behave very differently as $\epsilon$ decreases for these two problems and
 
@@
 %\begin{algorithm}
-%\caption{Overrelaxed SK algorithm}
+%\caption{Over-relaxed SK algorithm}
 %\label{SOR}
@@
 %\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
-%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Overrelaxed SK iteration}
+%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed SK iteration}
 %\ELSE
@@
 \section{Conclusion and perspectives}
-The SK algorithm is widely used to solve entropy regularized OT.
-The use of overrelaxed projections turns out to be a natural and simple idea to accelerate convergence while keeping many nice properties of this algorithm (first order, parallelizable, simple).
+SK algorithm is widely used to solve entropy regularized OT.
+The use of overrelaxed projections turns out to be a natural and simple idea to accelerate convergence while keeping many nice properties of this algorithm (first order, paralelizable, simple).
 We have proposed an algorithm that adaptively chooses the overrelaxation parameter so as to guarantee global convergence.
@@
 
-This idea of overrelaxation can be generalized to solve more general problems such as multi-marginal OT, barycenters, gradient flows, unbalanced OT~\cite[chap.
+This idea of overrelaxation can be generalized to solve more general problems such as barycenters, gradient flows, unbalanced OT~\cite[chap.
 4]{chizat2017these} but there is no systematic way to derive globally convergent algorithms.

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 INSA Toulouse\\
-Institut de Math\'ematiques de Toulouse\\
+%Institut de Math\'ematiques de Toulouse\\
 Toulouse,  France\\
@@
 We will use this function to enforce the strict descent criterion $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
-%
+
 The choice of \eqref{eq:lyapunov_function} as a Lyapunov function is of course related to the fact that Bregman projections are used throughout the algorithm.
-Further, we will show (Lemma \ref{lemma:lyapunov_decrease}) that its decrease is very easy to compute and this descent criterion still allows enough freedom in the choice of the overrelaxation parameter.
+Further, we will show in Lemma \ref{lemma:lyapunov_decrease} that its decrease is very easy to compute and this descent criterion still allows enough freedom in the choice of the overrelaxation parameter.
 
@@
 \end{figure}
+
 Figure~\ref{fig:comparison} displays the ratio of the number of iterations required to reach a precision of $10^{-6}$ on the dual variable $\alpha$ for SK algorithm and Algorithm \ref{SOR}.

### 2017-10-20 paper.tex
Relecture Nicolas

--- 2017-10-20 paper.tex
@@
 \end{equation}
+
 where $\gamma^*$ denotes the solution of the regularized OT problem.
-We will use this function to enforce the strict descent criterion $F(\gamma^{\ell+1}) < F(\gamma^\ell)$ as long as the process has not converged.
-%
+We will use this function to enforce the strict descent criterion $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
+
 The choice of \eqref{eq:lyapunov_function} as a Lyapunov function is of course related to the fact that Bregman projections are used throughout the algorithm.
-Further, we will show (Lemma \ref{lemma:lyapunov_decrease}) that its decrease is very easy to compute and this descent criterion still allows enough freedom in the choice of the overrelaxation parameter.
+Further, we will show in Lemma \ref{lemma:lyapunov_decrease} that its decrease is very easy to compute and this descent criterion still allows enough freedom in the choice of the overrelaxation parameter.
 

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 where $\gamma^*$ denotes the solution of the regularized OT problem.
-We will use this function to enforce the strict descent criterion $F(\gamma^{k+1}) < F(\gamma^k)$ as long as the process has not converged.
-
+We will use this function to enforce the strict descent criterion $F(\gamma^{\ell+1}) < F(\gamma^\ell)$ as long as the process has not converged.
+%
 The choice of \eqref{eq:lyapunov_function} as a Lyapunov function is of course related to the fact that Bregman projections are used throughout the algorithm.
-Further, we will show in Lemma \ref{lemma:lyapunov_decrease} that its decrease is very easy to compute and this descent criterion still allows enough freedom in the choice of the overrelaxation parameter.
+Further, we will show (Lemma \ref{lemma:lyapunov_decrease}) that its decrease is very easy to compute and this descent criterion still allows enough freedom in the choice of the overrelaxation parameter.
 

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 It generalizes and improves upon the widely-used iterative Bregman projections algorithm (or Sinkhorn--Knopp algorithm).
-The idea is to over-relax the Bregman projection operators, allowing for faster convergence.
+The idea is to overrelax the Bregman projection operators, allowing for faster convergence.
 In practice this corresponds to elevating the diagonal scaling factors to a given power, at each step of the algorithm.
@@
 \end{equation}
-
 where $\gamma^*$ denotes the solution of the regularized OT problem.
 We will use this function to enforce the strict descent criterion $F(\gamma^{\ell+1}) < F(\gamma^\ell)$ as long as the process has not converged.
-%
+
 The choice of \eqref{eq:lyapunov_function} as a Lyapunov function is of course related to the fact that Bregman projections are used throughout the algorithm.

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 %\begin{lemma}
-	Moreover, for $\gamma$ in $\IR^{mn}_{+*}$, the decrease of the Lyapunov function after an overrelaxed projection can be computed as
+	Moreover, for any $\gamma$ in $\IR^{mn}_{+*}$, the decrease of the Lyapunov function after an overrelaxed projection can be computed as
 	\begin{equation} \label{eq:kl_diff_scal}
@@
 %\begin{proof}
-%
+
 Formula \eqref{eq:kl_diff_scal} comes the expression $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$ and the relations \eqref{eq:def_or_proj} and \eqref{scaling}.
@@
 In Figure \ref{phi_omega}, we display the function  $\varphi_\omega(x)$.
-Notice that for the SK algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative.
+Notice that for the Sinkhorn--Knopp algorithm, which corresponds to $\omega=1$, the function $\varphi_\omega$ is always nonnegative.
 For other values $1\le\omega<2$, it is nonnegative for $x$ close to 1.
@@
 It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
-Note that with the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
+With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
 

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 \begin{center}
-\includegraphics[width=7cm]{images/cvgce_zone}
+\includegraphics[width=9cm]{images/cvgce_zone}
 \caption{\label{phi_omega} Value of $\varphi_\omega(x)$.
@@
 \end{proposition}
-This construction has several advantages:
-\begin{itemize}
-\item it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
-\item it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
-\item experiments show that this is often an excellent choice in practice.
-\end{itemize}
+This construction, is often an excellent choice in practice, has several advantages:
+\begin{enumerate}[align=right,labelsep=10pt,leftmargin=0.5cm]
+\item[$\bullet$] it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
+\item[$\bullet$] it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
+%\item[$\bullet$] experiments show that this is often an excellent choice in practice.
+\end{enumerate}
 
@@
 It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
-With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
-
+With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.is \vspace{-0.1cm}
 \begin{algorithm}
@@
  \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
-\end{algorithmic}
+\end{algorithmic}\vspace*{-0.1cm}
 \end{algorithm}
@@
 \begin{proof}
-For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual}  and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
+For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual}  and we will rely  on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
 Notice that the dual problem~\eqref{DROT} has an invariant subspace due to the invariance by translation $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is strictly convex in any subspace which does not contain this line.
@@
 
-The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative} , even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.
+The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative}, even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.
+In this context, when alternate maximization converges linearly at the rate $1-\eta$,  the local linear convergence rate of the SOR method is known to be $(1-\sqrt{\eta})/(1+\sqrt{\eta})$ ~\cite[Chapter 3]{chizat2017these}.
 

### 2017-10-20 paper.tex
Relecture Nicolas
reshaped local analysis

--- 2017-10-20 paper.tex
@@
 INSA Toulouse\\
-%Institut de Math\'ematiques de Toulouse\\
+Institut de Math\'ematiques de Toulouse\\
 Toulouse,  France\\
@@
 We will use this function to enforce the strict descent criterion $F(\gamma^{\ell+1}) < F(\gamma^\ell)$ as long as the process has not converged.
-
+%
 The choice of \eqref{eq:lyapunov_function} as a Lyapunov function is of course related to the fact that Bregman projections are used throughout the algorithm.
@@
 \item it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
-\item it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
+\item it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method);
 \item experiments show that this is often an excellent choice in practice.
@@
 	The function $\Theta^*\left((A_k \gamma)\oslash \mu^k)\right)$ is such that every term in relation (\ref{eq:kl_diff_scal}) is non-negative.
-	Therefore, by lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
+	Therefore, by Lemma \ref{lemma:F_P_theta}, using this parameter minus $\delta$ ensures the strong decrease (\ref{eq:cond_theta_k}) of the Lyapunov function.
 	Constraining the parameter to $[1,\theta_0]$ preserves this property.% as using $\theta=1$ always satisfies inequality \eqref{eq:cond_theta_k} and using $\theta=\theta_0$ when it is smaller than $\Theta^*\left(\frac{A_k \gamma}{\mu^k}\right)$ does as well.
@@
 \label{section:local}
-In order to theoretically justify the acceleration of convergence that is observed in practice, we now study the local convergence rate of the algorithm.
-%We do not claim that this simple overrelaxation is the most efficient method to accelerate local convergence of fixed points---adaptive methods with excellent properties exist e.g.~\cite{2016arXiv160604133S}--- but this local analysis gives a hint for what makes our scheme relevant is the conjunction of our global convergence guarantee with this local acceleration property.
+In order to justify the acceleration of convergence that is observed in practice, we now study the local convergence rate of the overrelaxed algorithm, which follows from the classical convergence analysis of the linear SOR method.
+Our result involves the second largest eigenvalue of the matrix
+ \[
+M_1= \diag(1\oslash \mu^1) \, \gamma^* \, \diag(1\oslash \mu^2)\, ^t\gamma^*
+\]
+where $\gamma^*$ is the solution to the regularized OT problem.
+This matrix corresponds to the linearized dual SK iteration at the maximizer.
+We denote this second largest eigenvalue by $1-\eta$, which satisfies $\eta>0$ (see \cite{knight2008sinkhorn}).
 
-%The local behavior of the extrapolated is summarized in the next result, which gathers many results from the literature.
 \begin{proposition}\label{prop:local}
-Let $1-\eta$ be the local linear rate of convergence of the SK algorithm.
+The SK algorithm converges locally at a linear rate $1-\eta$.
+For the optimal choice of extrapolation parameter $\theta^* = 2/(1+\sqrt{\eta})$, the overrelaxed projection algorithm converges locally linearly at a rate $(1-\sqrt{\eta})/(1+\sqrt{\eta})$.
+More generally, the local convergence of the overrelaxed algorithm is guaranteed as long as $\theta \in {]0,2[}$ and the linear rate is given on Figure~\ref{fig:eigtransform} as a function of $1-\eta$ and $\theta$.
+\end{proposition}
+
+\begin{proof}
+In this proof, we focus on the dual problem and we recall the relationship $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ between the iterates of the overrelaxed projection algorithm $\gamma^\ell$ and the  iterates $(\alpha^\ell,\beta^\ell)$ of the SOR algorithm on the dual problem~\eqref{SORdual}, initialized with $(\alpha^0,\beta^0)=(0,0)$.
+%First notice that if $\gamma^\ell \to \gamma^*$, then $\KL(\gamma^\ell,\gamma^0)\to\KL(\gamma^*,\gamma^0)$, which is equivalent to $E(\alpha^\ell,\beta^\ell)$ converges to the maximum of $E$.
+The dual problem~\eqref{DROT} is invariant by translations of the form $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is otherwise strictly convex on any subspace which does not contain the line $\IR(\mathbf{1},-\mathbf{1})$ (this invariance cancels in the corresponding primal iterates).
+In order to deal with this invariance, consider the subspace $S$ of pairs of dual variables $(\alpha,\beta)$ that satisfy $\alpha_1=0$, let $\pi_S$ be the (non orthogonal) projection on $S$ of kernel $(\mathbf{1},-\mathbf{1})$ and let $(\alpha^*,\beta^*)\in S$ be a dual maximizer.
+
+% 
+Since the SOR iteration it smooth, its local convergence is characterized by the local convergence of its linearized version, which corresponds to the SOR method applied to maximize the quadratic expansion of the dual objective $E$ at $(\alpha^*,\beta^*)$.
+This defines an affine map $M_\theta : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ whose spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative} (see also \cite[chapter 4]{chizat2017these} for the specific case of convex minimization).
+Specifically, in the non strictly concave case~\cite{hadjidimos1985optimization}, we have that the operator $\pi_S\circ M_1^\ell$ converges at the linear rate $1-\eta$ towards the projector on $(\alpha^*,\beta^*)$ and that the convergence of $\pi_S\circ M_\theta^\ell$ is guaranteed for $\theta \in ]0,2[$, with the rate
+\[
+f(\theta,\eta) = \begin{cases}
+\theta-1 & \text{if $\theta$}\\
+\frac12 \theta^2(1-\eta) - (\theta-1) +\frac12 \sqrt{(1-\eta)\theta^2(\theta^2(1-\eta)-4(\theta-1))} & \text{otherwise.}
+\end{cases}
+\]
+where $\theta^*=2/(1+\sqrt{\eta})$ is the optimal parameter, for which $f(\theta,\eta)=(1-\sqrt{\eta})/(1+\sqrt{\eta})$. The function $f$ is plotted in Figure \ref{fig:eigtransform}.
+
+To switch from these dual convergence results to primal results, remark that $\gamma^\ell \to \gamma^*$ implies $\KL(\gamma^\ell,\gamma^0)\to \KL(\gamma^*,\gamma^0)$ which in turn implies $E(\alpha^\ell,\beta^\ell) \to \max E$ so invoking the partial strict concavity of $E$, $\pi_S(\alpha^\ell,\beta^\ell)\to (\alpha^*,\beta^*)$.
+The converse implication is direct.
+We conclude by noting that the fact that $\pi_S(\alpha^\ell, \beta^\ell)$ converges at a linear rate implies the same rate on $\gamma^\ell$, thanks to the relationship between the iterates.
+%Using standard arguments, we have that the local convergence rate of the SOR procedure on the dual objective $E$ is that of the SOR procedure applied to the quadratic form defined by its Hessian and the linear rate of \emph{semi-convergence} to the projection on the set of minimizers.
+\end{proof}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.65]{images/eigen_transform}
+\caption{\label{fig:eigtransform} Local linear rate of convergence of the overrelaxed algorithm as a function of $1-\eta$, the local convergence rate of SK algorithm and $\theta$ the overrelaxation parameter.
+(plain curve) the original rate is recovered for $\theta=1$.
+(dashed curve) optimal overrelaxation parameter $\theta^*$.}
+\end{figure}
+
+%\begin{proposition}%\label{prop:local}
+%Let $1-\eta$ be the (worst case) local linear rate of convergence of the SK algorithm.
 By \cite{knight2008sinkhorn} we know that $\eta>0$.
 Then, for the optimal choice of parameter $\theta_0 = 2/(1+\sqrt{\eta})$, the overrelaxed projection algorithm converges locally linearly at a rate $(1-\sqrt{\eta})/(1+\sqrt{\eta})$.
-\end{proposition}
+%\end{proposition}
 
-Note that the local linear rate is also improved if $\theta_0$ is not the optimal choice, see \cite[chapter 5]{ciarlet1982introduction}.
-By remarking that for any choice of $\theta_0<2$ there is a neighborhood of $1$ on which $\varphi_{\theta_0}(\cdot)$ is nonnegative, one sees that this result applies in particular to our algorithm because eventually, for $\ell$ big enough, one has $\Theta(\gamma^\ell)=\theta_0$.
+%Note that the local linear rate is also improved if $\theta_0$ is not the optimal choice, see \cite[chapter 5]{ciarlet1982introduction}, \cite[chapter 4]{chizat2017these}.
+
 \begin{corollary}
-The preceding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
+The previous local convergence analysis applies to Algorithm \ref{thm:algo}, where the local convergence rate is governed by the choice of the target extrapolation parameter $\theta_0$.
 \end{corollary}
 \begin{proof}
-For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual}  and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$.
-Notice that the dual problem~\eqref{DROT} has an invariant subspace due to the invariance by translation $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is strictly convex in any subspace which does not contain this line.
-Also, this invariance disappears in the associated primal iterates.
-
-The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative} , even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.
-
-Using standard arguments, we have that the local convergence rate of the SOR procedure on the dual objective $E$ is that of the SOR procedure applied to the quadratic form defined by its Hessian and the linear rate of \emph{semi-convergence} to the projection on the set of minimizers.
+What we need to show is that eventually one always has $\Theta(\gamma^\ell)=\theta_0$.
+This can be seen from the quadratic Taylor expansion $\varphi_{\theta_0}(1+z)=z^2(\theta_0 -\theta_0^2/2)+ o(z^2)$, which shows that for any choice of $\theta_0<2$ there is a neighborhood of $1$ on which $\varphi_{\theta_0}(\cdot)$ is nonnegative.
 \end{proof}
@@
 \section{Experimental results}
-We compare Algorithm~\ref{SOR} to SK in two very different settings.
+We compare Algorithm~\ref{SOR} to SK algorithm in two very different settings.
 In setting (a) we consider the domain $[0,1]$ discretized into $100$ samples and the squared Euclidean transport cost on this domain.

### 2017-10-20 paper.tex

@@
 We will use this function to enforce the strict descent criterion $F(\gamma^{\ell+1}) < F(\gamma^\ell)$ as long as the process has not converged.
-%
+
 The choice of \eqref{eq:lyapunov_function} as a Lyapunov function is of course related to the fact that Bregman projections are used throughout the algorithm.
@@
 \end{proposition}
-This construction has several advantages:
-\begin{itemize}
-\item it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
-\item it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only require to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method);
-\item experiments show that this is often an excellent choice in practice.
-\end{itemize}
+
+This construction, is often an excellent choice in practice, has several advantages:
+\begin{enumerate}[align=right,labelsep=10pt,leftmargin=0.5cm]
+\item[$\bullet$] it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
+\item[$\bullet$] it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only requires to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
+%\item[$\bullet$] experiments show that this is often an excellent choice in practice.
+\end{enumerate}
 
@@
 It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
-Note that with the choice $\theta_0=1$, one recovers exactly the original SK algorithm.
-
+With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.is \vspace{-0.1cm}
 \begin{algorithm}
@@
 4]{chizat2017these} but there is no systematic way to derive globally convergent algorithms.
-Our work is an effort in the direction of building and understanding the properties of robust first order algorithms for solving OT.
+Our work is a step in the direction of building and understanding the properties of robust first order algorithms for solving OT.
 More understanding is needed regarding SOR itself (global convergence speed, choice of $\theta_0$), but also its relation to other acceleration methods~\cite{2016arXiv160604133S,2017arXiv170509634A}.

### 2017-10-20 paper.tex
small modifs

@@
 \section{Introduction}
-Optimal Transport (OT) is an efficient and flexible tool to compare two probability distributions which has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
+Optimal Transport is an efficient and flexible tool to compare two probability distributions which has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
 The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64}   scaling with high dimensional data.
@@
 \end{equation}
-The function $E$ is concave, continuously differentiable and admits a maximizer, so alternate maximization converges and we recover, for $a_i=e^{\alpha_i/\epsilon}$, $b_j=e^{\beta_j/\epsilon}$ and $\gamma^0_{i,j}=e^{-c_{i,j}/\epsilon}$:
-\begin{align*}
-a^{\ell+1} &= {\mu^1}\oslash{\gamma^0 b^{\ell}} &
-b^{\ell+1} &= {\mu^2}\oslash{^t \gamma^0 a^{\ell+1}} 
-\end{align*}
+The function $E$ is concave, continuously differentiable and admits a maximizer, so alternate maximization converges and we recover SK algorithm by posing for $a_i=e^{\alpha_i/\epsilon}$, $b_j=e^{\beta_j/\epsilon}$ and $\gamma^0_{i,j}=e^{-c_{i,j}/\epsilon}$.
+%\begin{align*}
+%a^{\ell+1} &= {\mu^1}\oslash{\gamma^0 b^{\ell}} &
+%b^{\ell+1} &= {\mu^2}\oslash{^t \gamma^0 a^{\ell+1}} 
+%\end{align*}
 
@@
 In this paper, we consider an overrelaxation scheme designed to accelerate the Sinkhorn--Knopp algorithm.
-We first present and show the convergence of our algorithm in section 2.
-In section 3, we analyze the local convergence rate of the algorithm to justify the acceleration.
+We first present and show the convergence of our algorithm in Section 2.
+In Section 3, we analyze the local convergence rate of the algorithm to justify the acceleration.
 We finally demonstrate numerically  in Section 4 the good behavior of our method, where larger accelerations are observed for decreasing values of $\epsilon$.
@@
 The introduction of relaxation variables through heavy ball approaches \cite{POLYAK19641} has recently gained in popularity  to speed up the convergence of algorithms optimizing convex \cite{2014arXiv1412.7457G} or non convex \cite{Zavriev1993,2016arXiv160609070O} problems.
-Such schemes have also been empirically considered to accelerate the SK algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
-The convergence of these algorithms is nevertheless not studied yet in the context of regularized optimal transport.
+Our specific approach is very much related to the SOR algorithm~\cite{young2014iterative}, which is a classical way to solve linear systems.
+Similar schemes have been empirically considered to accelerate the SK algorithm  in \cite{peyre2016quantum,2017arXiv170801955S}.
+The convergence of these algorithms has nevertheless not been studied yet in the context of regularized optimal transport.
 
@@
 
-As illustrated in Figure \ref{alternate_projections} (a-b), the SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
+As illustrated in Figure \ref{alternate_projections} (a-b), SK algorithm, that  performs alternate Bregman projections onto the affine sets $\Ccal_1$ and $\Ccal_2$, can be very slow when $\epsilon\to 0$.
 The idea developed in this paper is to perform overrelaxed projections in order to accelerate the process, as displayed in Figure \ref{alternate_projections} (c).
@@
 
-Duality gives another point of view on the iterative overrelaxed Bregman projections.
-They indeed correspond to a successive overelaxation (SOR) algorithm on the dual objective $E$.
-This can be seen by using the relationships given after equation~\eqref{DROT}:
+Duality gives another point of view on the iterative overrelaxed Bregman projections: they indeed correspond to a successive overrelaxation (SOR) algorithm on the dual objective $E$.
+This is a procedure which, starting from $(\alpha^0,\beta^0)=(\mathbf{0},\mathbf{0})$, defines for $\ell \in \mathbb{N}^*$,
 \begin{align}\label{SORdual}
@@
 \end{align} 
-
+This can be seen by using the relationships given after equation~\eqref{DROT}.
 %{\color{blue} \begin{proposition} 
@@
 \begin{center}
-\includegraphics[width=9cm]{images/cvgce_zone}
+\includegraphics[width=8cm]{images/cvgce_zone}
 \caption{\label{phi_omega} Value of $\varphi_\omega(x)$.
-The function is positive above the black line, negative below.
+The function is positive above the red line, negative below.
 For any relaxation parameter $\omega$ smaller than $2$, there exists a neighborhood of $1$ on which $\varphi_\omega(\cdot)$ is positive.}
@@
 
-To construct explicitly functions $\theta_k$ as needed in Theorem~\ref{thm:algo}, we need the following lemma.
+We can construct explicitly functions $\theta_k$ as needed in Theorem~\ref{thm:algo} using the following lemma.
 \begin{lemma}\label{lemma:F_P_theta}
@@
 \begin{proof}
-	Thanks to lemma \ref{lemma:lyapunov_decrease}, one knows that
+	Thanks to Lemma \ref{lemma:lyapunov_decrease}, one knows that
 	\[
@@
 \end{proposition}
-
-This construction, is often an excellent choice in practice, has several advantages:
-\begin{enumerate}[align=right,labelsep=10pt,leftmargin=0.5cm]
-\item[$\bullet$] it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in section \ref{section:local});
-\item[$\bullet$] it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only requires to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
-%\item[$\bullet$] experiments show that this is often an excellent choice in practice.
-\end{enumerate}
-
 \begin{proof}
@@
 
+This construction, which is often an excellent choice in practice, has several advantages:
+\begin{enumerate}[align=right,labelsep=10pt,leftmargin=0.5cm]
+\item[$\bullet$] it allows to choose arbitrarily the parameter $\theta_0$ that will be used eventually when the algorithm is close to convergence (we motivate what are good choices for $\theta_0$ in Section \ref{section:local});
+\item[$\bullet$] it is also an easy approach to having an adaptive method, as the approximation of $\Theta^*$ has a negligible cost (it only requires to solve a one dimensional problem that depends on the smallest value of $(A_k \gamma)\oslash \mu^k$, which can be done in a few iterations of Newton's method).
+%\item[$\bullet$] experiments show that this is often an excellent choice in practice.
+\end{enumerate}
+
 The resulting algorithm, which is proved to be convergent by Theorem~\ref{thm:algo}, is written in pseudo-code in Algorithm \ref{SOR}.
-It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which decreasing is as can be seen on Figure~\ref{phi_omega}.
+It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which is decreasing is as can be seen on Figure~\ref{phi_omega}.
 With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.is \vspace{-0.1cm}
@@
 Our result involves the second largest eigenvalue of the matrix
- \[
+\begin{equation}\label{eq:M1}
 M_1= \diag(1\oslash \mu^1) \, \gamma^* \, \diag(1\oslash \mu^2)\, ^t\gamma^*
-\]
-where $\gamma^*$ is the solution to the regularized OT problem.
-This matrix corresponds to the linearized dual SK iteration at the maximizer.
-We denote this second largest eigenvalue by $1-\eta$, which satisfies $\eta>0$ (see \cite{knight2008sinkhorn}).
+\end{equation}
+where $\gamma^*$ is the solution to the regularized OT problem. We denote this second largest eigenvalue by $1-\eta$, which satisfies $\eta>0$ (see \cite{knight2008sinkhorn}).
 
@@
 % 
-Since the SOR iteration it smooth, its local convergence is characterized by the local convergence of its linearized version, which corresponds to the SOR method applied to maximize the quadratic expansion of the dual objective $E$ at $(\alpha^*,\beta^*)$.
+Since the SOR iteration followed by $\pi_S$ is a smooth map, its local convergence properties are characterized by the local convergence of its linearization, which corresponds to the SOR method applied to the maximization of the quadratic Taylor expansion of the dual objective $E$ at $(\alpha^*,\beta^*)$.
 This defines an affine map $M_\theta : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ whose spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative} (see also \cite[chapter 4]{chizat2017these} for the specific case of convex minimization).
+For the case $\theta=1$, we obtain the matrix $M_1$ defined in \eqref{eq:M1}.
 Specifically, in the non strictly concave case~\cite{hadjidimos1985optimization}, we have that the operator $\pi_S\circ M_1^\ell$ converges at the linear rate $1-\eta$ towards the projector on $(\alpha^*,\beta^*)$ and that the convergence of $\pi_S\circ M_\theta^\ell$ is guaranteed for $\theta \in ]0,2[$, with the rate
@@
 f(\theta,\eta) = \begin{cases}
-\theta-1 & \text{if $\theta$}\\
-\frac12 \theta^2(1-\eta) - (\theta-1) +\frac12 \sqrt{(1-\eta)\theta^2(\theta^2(1-\eta)-4(\theta-1))} & \text{otherwise.}
+\theta-1 & \text{if $\theta>\theta^*$}\\
+\frac12 \theta^2(1-\eta) - (\theta-1) +\frac12 \sqrt{(1-\eta)\theta^2(\theta^2(1-\eta)-4(\theta-1))} & \text{otherwise},
 \end{cases}
 \]
-where $\theta^*=2/(1+\sqrt{\eta})$ is the optimal parameter, for which $f(\theta,\eta)=(1-\sqrt{\eta})/(1+\sqrt{\eta})$. The function $f$ is plotted in Figure \ref{fig:eigtransform}.
+where $\theta^*\coloneqq2/(1+\sqrt{\eta})$ is the optimal parameter, for which $f(\theta^*,\eta)=(1-\sqrt{\eta})/(1+\sqrt{\eta})$. The function $f$ is plotted in Figure \ref{fig:eigtransform}.
 
-To switch from these dual convergence results to primal results, remark that $\gamma^\ell \to \gamma^*$ implies $\KL(\gamma^\ell,\gamma^0)\to \KL(\gamma^*,\gamma^0)$ which in turn implies $E(\alpha^\ell,\beta^\ell) \to \max E$ so invoking the partial strict concavity of $E$, $\pi_S(\alpha^\ell,\beta^\ell)\to (\alpha^*,\beta^*)$.
-The converse implication is direct.
+To switch from these dual convergence results to primal convergence results, remark that $\gamma^\ell \to \gamma^*$ implies $\KL(\gamma^\ell,\gamma^0)\to \KL(\gamma^*,\gamma^0)$ which in turn implies $E(\alpha^\ell,\beta^\ell) \to \max E$ so invoking the partial strict concavity of $E$, $\pi_S(\alpha^\ell,\beta^\ell)\to (\alpha^*,\beta^*)$.
+The converse implication is direct so it holds $[\pi_S(\alpha^\ell,\beta^\ell)\to (\alpha^*,\beta^*)] \Leftrightarrow [\gamma^\ell \to \gamma^*]$.
 We conclude by noting that the fact that $\pi_S(\alpha^\ell, \beta^\ell)$ converges at a linear rate implies the same rate on $\gamma^\ell$, thanks to the relationship between the iterates.
@@
 What we need to show is that eventually one always has $\Theta(\gamma^\ell)=\theta_0$.
-This can be seen from the quadratic Taylor expansion $\varphi_{\theta_0}(1+z)=z^2(\theta_0 -\theta_0^2/2)+ o(z^2)$, which shows that for any choice of $\theta_0<2$ there is a neighborhood of $1$ on which $\varphi_{\theta_0}(\cdot)$ is nonnegative.
+This can be seen from the quadratic Taylor expansion $\varphi_{\theta_0}(1+z)=z^2(\theta_0 -\theta_0^2/2)+ o(z^2)$, which shows that for any choice of $\theta_0\in {]1,2[}$, there is a neighborhood of $1$ on which $\varphi_{\theta_0}(\cdot)$ is nonnegative.
 \end{proof}
@@
 \section{Experimental results}
-We compare Algorithm~\ref{SOR} to SK algorithm in two very different settings.
+We compare Algorithm~\ref{SOR} to SK algorithm on two very different optimal transport settings.
 In setting (a) we consider the domain $[0,1]$ discretized into $100$ samples and the squared Euclidean transport cost on this domain.
-The marginals are densities made of the sum of a base plateau of height $0.1$ and another plateau of heights and boundaries chosen uniformly in $[0,1]$, subsequently normalized.
+The marginals are densities made of the sum of a base plateau of height $0.1$ and another plateau of height and boundaries chosen uniformly in $[0,1]$, subsequently normalized.
 In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.
@@
 For estimating $\eta$, we follow two strategies.
-For strategy ``estimated'' (in blue on figure~\ref{fig:comparison}), $\eta$ is measured by looking at the local convergence rate of SK run on another random problem of the same setting and for the same value of $\epsilon$.
-For strategy ``measured'' (in orange on figure~\ref{fig:comparison}) the parameter is set using the local convergence rate of SK run on the same problem.
-Of course, the latter is unrealistic strategy but our experiment shows that the acceleration power is not so much impacted by estimating once for all the overrelaxation parameter for a class of problem with fixed regularization, as shown on~\ref{fig:comparison}.
+For strategy ``estimated'' (in blue on Figure~\ref{fig:comparison}), $\eta$ is measured by looking at the local convergence rate of SK run on another random problem of the same setting and for the same value of $\epsilon$.
+For strategy ``measured'' (in orange on Figure~\ref{fig:comparison}) the parameter is set using the local convergence rate of SK run on the same problem.
+Of course, the latter is an unrealistic strategy but it is interesting to see in our experiments that the ``estimated'' strategy performs almost as well as the ``measured'' one, as shown on~\ref{fig:comparison}.
 \begin{figure}
@@
 Figure~\ref{fig:comparison} displays the ratio of the number of iterations required to reach a precision of $10^{-6}$ on the dual variable $\alpha$ for SK algorithm and Algorithm \ref{SOR}.
-It is is worth noting that the complexity per iteration is the same, modulo negligible terms, so this ratio is also the runtime ratio (our algorithm can also be parallelized on GPUs just as SK algorithm).
+It is is worth noting that the complexity per iteration of these algorithms is the same modulo negligible terms, so this ratio is also the runtime ratio (our algorithm can also be parallelized on GPUs just as SK algorithm).
 In both experimental settings, for low values of the regularization parameter $\epsilon$, the acceleration ratio is above $20$  with Algorithm \ref{SOR}.

### 2017-10-20 paper.tex
small modifs

@@
 \end{equation}
-where $\gamma^*$ is the solution to the regularized OT problem. We denote this second largest eigenvalue by $1-\eta$, which satisfies $\eta>0$ (see \cite{knight2008sinkhorn}).
+where $\gamma^*$ is the solution to the regularized OT problem (the largest eigenvalue is $1$, associated to the eigenvector $\mathbf{1}$).
+We denote this second largest eigenvalue by $1-\eta$, which satisfies $\eta>0$ (see \cite{knight2008sinkhorn}).
 
@@
 %First notice that if $\gamma^\ell \to \gamma^*$, then $\KL(\gamma^\ell,\gamma^0)\to\KL(\gamma^*,\gamma^0)$, which is equivalent to $E(\alpha^\ell,\beta^\ell)$ converges to the maximum of $E$.
-The dual problem~\eqref{DROT} is invariant by translations of the form $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is otherwise strictly convex on any subspace which does not contain the line $\IR(\mathbf{1},-\mathbf{1})$ (this invariance cancels in the corresponding primal iterates).
-In order to deal with this invariance, consider the subspace $S$ of pairs of dual variables $(\alpha,\beta)$ that satisfy $\alpha_1=0$, let $\pi_S$ be the (non orthogonal) projection on $S$ of kernel $(\mathbf{1},-\mathbf{1})$ and let $(\alpha^*,\beta^*)\in S$ be a dual maximizer.
+The dual problem~\eqref{DROT} is invariant by translations of the form $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is otherwise strictly convex on any subspace which does not contain the line $\IR(\mathbf{1},-\mathbf{1})$.
+In order to deal with this invariance (which cancels in the corresponding primal iterates), consider the subspace $S$ of pairs of dual variables $(\alpha,\beta)$ that satisfy $\alpha_1=0$, let $\pi_S$ be the (non orthogonal) projection on $S$ of kernel $(\mathbf{1},-\mathbf{1})$ and let $(\alpha^*,\beta^*)\in S$ be a dual maximizer.
 
 % 
-Since the SOR iteration followed by $\pi_S$ is a smooth map, its local convergence properties are characterized by the local convergence of its linearization, which corresponds to the SOR method applied to the maximization of the quadratic Taylor expansion of the dual objective $E$ at $(\alpha^*,\beta^*)$.
+Since one SOR iteration is a smooth map, the local convergence properties of the SOR algorithm are characterized by the local convergence of its linearization, which here corresponds to the SOR method applied to the maximization of the quadratic Taylor expansion of the dual objective $E$ at $(\alpha^*,\beta^*)$.
 This defines an affine map $M_\theta : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ whose spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative} (see also \cite[chapter 4]{chizat2017these} for the specific case of convex minimization).
-For the case $\theta=1$, we obtain the matrix $M_1$ defined in \eqref{eq:M1}.
+For the case $\theta=1$, this corresponds to the matrix $M_1$ defined in \eqref{eq:M1}.
 Specifically, in the non strictly concave case~\cite{hadjidimos1985optimization}, we have that the operator $\pi_S\circ M_1^\ell$ converges at the linear rate $1-\eta$ towards the projector on $(\alpha^*,\beta^*)$ and that the convergence of $\pi_S\circ M_\theta^\ell$ is guaranteed for $\theta \in ]0,2[$, with the rate
@@
 \begin{corollary}
-The previous local convergence analysis applies to Algorithm \ref{thm:algo}, where the local convergence rate is governed by the choice of the target extrapolation parameter $\theta_0$.
+The previous local convergence analysis applies to Algorithm \ref{thm:algo} and the local convergence rate is governed by the choice of the target extrapolation parameter $\theta_0$.
 \end{corollary}
@@
 \end{proof}
+
+%Of course, a difficulty is that the number $\eta$ is not known a priori. In the numerical experiments, we have noticed that
 

### 2017-10-20 paper.tex
Relecture Nicolas

@@
 
-Efficient parallel computations can be considered \cite{cuturi13} and one can even reach real-time computation for large scale problem for certain class of cost matrices $c$ \cite{Solomon2015}. 
+Efficient parallel computations can be considered \cite{cuturi13} and one can almost reach real-time computation for large scale problem for certain class of cost matrices $c$  allowing the use of seprable convolutions \cite{Solomon2015}.
 For small values of the parameter $\epsilon$, numerical issues can arise and a stabilization of the algorithm is necessary \cite{2016arXiv160705816C}.
@@
 It uses the function $\Theta$ defined implicitly in \eqref{eq:Theta}, which in practice is approximated with a few iterations of Newton's method on the function $\omega \mapsto \varphi_\omega(\min u)$ which is decreasing is as can be seen on Figure~\ref{phi_omega}.
-With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.is \vspace{-0.1cm}
+With the choice $\theta_0=1$, one recovers exactly the original SK algorithm.is
 \begin{algorithm}
@@
  \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
-\end{algorithmic}\vspace*{-0.1cm}
+\end{algorithmic}
 \end{algorithm}
@@
 where $\gamma^*$ is the solution to the regularized OT problem (the largest eigenvalue is $1$, associated to the eigenvector $\mathbf{1}$).
-We denote this second largest eigenvalue by $1-\eta$, which satisfies $\eta>0$ (see \cite{knight2008sinkhorn}).
+We denote the second largest eigenvalue by $1-\eta$, it satisfies $\eta>0$  \cite{knight2008sinkhorn}.
 
@@
 For the optimal choice of extrapolation parameter $\theta^* = 2/(1+\sqrt{\eta})$, the overrelaxed projection algorithm converges locally linearly at a rate $(1-\sqrt{\eta})/(1+\sqrt{\eta})$.
-More generally, the local convergence of the overrelaxed algorithm is guaranteed as long as $\theta \in {]0,2[}$ and the linear rate is given on Figure~\ref{fig:eigtransform} as a function of $1-\eta$ and $\theta$.
+The local convergence of the overrelaxed algorithm is guaranteed for $\theta \in {]0,2[}$ and the linear rate is given on Figure~\ref{fig:eigtransform} as a function of $1-\eta$ and $\theta$.
 \end{proposition}
@@
 The converse implication is direct so it holds $[\pi_S(\alpha^\ell,\beta^\ell)\to (\alpha^*,\beta^*)] \Leftrightarrow [\gamma^\ell \to \gamma^*]$.
-We conclude by noting that the fact that $\pi_S(\alpha^\ell, \beta^\ell)$ converges at a linear rate implies the same rate on $\gamma^\ell$, thanks to the relationship between the iterates.
+We conclude by noting the fact that $\pi_S(\alpha^\ell, \beta^\ell)$ converges at a linear rate implies the same rate on $\gamma^\ell$, thanks to the relationship between the iterates.
 %Using standard arguments, we have that the local convergence rate of the SOR procedure on the dual objective $E$ is that of the SOR procedure applied to the quadratic form defined by its Hessian and the linear rate of \emph{semi-convergence} to the projection on the set of minimizers.
@@
 \centering
-\includegraphics[scale=0.65]{images/eigen_transform}
+\includegraphics[scale=0.8]{images/eigen_transform}
 \caption{\label{fig:eigtransform} Local linear rate of convergence of the overrelaxed algorithm as a function of $1-\eta$, the local convergence rate of SK algorithm and $\theta$ the overrelaxation parameter.

### 2017-10-27 paper.tex
Relecture Nicolas

@@
 
-Formula \eqref{eq:kl_diff_scal} comes the expression $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$ and the relations \eqref{eq:def_or_proj} and \eqref{scaling}.
+Formula \eqref{eq:kl_diff_scal} comes from the expression $F(\gamma^1)-F(\gamma^2)= \sum_{i,j}\left(\gamma^*_{i,j}\log(\gamma^2_{i,j}/\gamma^1_{i,j})+\gamma^1_{i,j}-\gamma^2_{i,j}\right)$ and the relations \eqref{eq:def_or_proj} and \eqref{scaling}.
 \end{proof}

### 2017-11-19 PosterNIPS2017.tex
Poster files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (11/06/13)
%
% baposter Class Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[portrait,a0paper,fontscale=0.35]{baposter} % Adjust the font scale/size here : fontscale=0.285
\usepackage{textpos}
\usepackage[frenchb]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx} % Required for including images
%\graphicspath{{Figures/}} % Directory in which figures are stored
\usepackage{amsmath} % For typesetting math
\usepackage{amsthm}
\usepackage{amssymb} % Adds new symbols to be used in math mode
\usepackage{booktabs} % Top and bottom rules for tables
\usepackage{enumitem} % Used to reduce itemize/enumerate spacing
\usepackage{palatino} % Use the Palatino font
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{multicol} % Required for multiple columns
\usepackage{tikz} % Required for flow chart
\usetikzlibrary{shapes,arrows} % Tikz libraries required for the flow chart in the template
\usepackage{dsfont}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{color}
\usepackage{nth}
\usepackage{a4wide}
\usepackage{graphicx,color}
\usepackage{array}
\usepackage[nottoc, notlof, notlot]{tocbibind}
\usepackage{subfigure}
\usetikzlibrary{arrows,shapes,positioning,calc}
%\definecolor{CPU}{RGB}{30, 167, 223}
%\definecolor{CPU}{RGB}{8, 138, 104}
\definecolor{gris}{gray}{0.9}
\usepackage{geometry}
\geometry{left=0.5cm,right=0.5cm,bottom=0.5cm,top=0.5cm,centering}
\usepackage{subfig}
\usepackage{float}
\usepackage{multicol}
\usepackage{enumitem,rotating}

\newcommand{\sidecap}[1]{ {\begin{sideways}\parbox{0.2\textwidth}{\centering #1}\end{sideways}} }

\theoremstyle{plain}
\newtheorem*{thm}{Theorem}
%\theoremstyle{definition}
%\newtheorem*{defi}[thm]{Definition}
\theoremstyle{plain}
\newtheorem*{prop}{Properties}
\theoremstyle{plain}
\newtheorem*{hyp}{Assumption}
\theoremstyle{plain}
\newtheorem*{proposition}{Proposition}

\setlength{\columnsep}{1.5em} % Slightly increase the space between columns
\setlength{\columnseprule}{0mm} % No horizontal rule between columns
\def\diag{\mathop{\rm diag}\nolimits}%
\def\argmin{\mathop{\rm arg \; min}\limits}%

\definecolor{lightred}{rgb}{0.5 0 0}

\newcommand{\compresslist}{ % Define a command to reduce spacing within itemize/enumerate environments, this is used right after \begin{itemize} or \begin{enumerate}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
}

\definecolor{CPU}{RGB}{0, 153, 125} % Defines the color used for content box headers

\begin{document}

\begin{poster}
{
background=plain,
headershade=plain,
headerborder=closed, % Adds a border around the header of content boxes
colspacing=0.8em, % Column spacing
bgColorOne=gris, % Background color for the gradient on the left side of the poster
bgColorTwo=gris, % Background color for the gradient on the right side of the poster
borderColor=CPU, % Border color
headerColorOne=CPU, % Background color for the header in the content boxes (left side)
headerColorTwo=CPU, % Background color for the header in the content boxes (right side)
headerFontColor=white, % Text color for the header text in the content boxes
boxColorOne=white, % Background color of the content boxes
textborder=roundedsmall, % Format of the border around content boxes, can be: none, bars, coils, triangles, rectangle, rounded, roundedsmall, roundedright or faded
eyecatcher=true, % Set to false for ignoring the left logo in the title and move the title left
headerheight=0.10\textheight, % Height of the header
headershape=rounded, % Specify the rounded corner in the content box headers, can be: rectangle, small-rounded, roundedright, roundedleft or rounded
headerfont=\Large\bf\textsc, % Large, bold and sans serif font in the headers of content boxes
%textfont={\setlength{\parindent}{1.5em}}, % Uncomment for paragraph indentation
linewidth=1.8pt, % Width of the border lines around content boxes
columns=2,
}
%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
%
{\includegraphics[height=5em]{logo_ens_psl.png}} % First university/lab logo on the left
{\bf Overrelaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport\vspace{0.2em}} 
{{Alexis Thibault, L\'enaïc Chizat, Charles Dossal and  Nicolas Papadakis}} % Author names and institution
{\includegraphics[height=5em]{logo_imb.png}} % Second university/lab logo on the right

%----------------------------------------------------------------------------------------
%	RESEARCH
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	Notations and definitions
%----------------------------------------------------------------------------------------

\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0}{

}

%----------------------------------------------------------------------------------------
%	
%----------------------------------------------------------------------------------------

\headerbox{Overrelaxation}{name=overrelaxation,column=1}{

}

%----------------------------------------------------------------------------------------
% Main result
%----------------------------------------------------------------------------------------

\headerbox{Lyapunov Function}{name=lyapunov,below=Sinkhorn,column=0}{

}

%----------------------------------------------------------------------------------------
%	Bootstrap
%----------------------------------------------------------------------------------------

\headerbox{Experiments}{name=expe,below=overrelaxation,column=1}{

}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\headerbox{References}{name=references,below=lyapunov,span =2}{
\noindent [1] M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
...
}

\end{poster}

\end{document}

### 2017-11-20 PosterNIPS2017.tex
Poster

@@
 \usepackage{enumitem,rotating}
+\usepackage{bbm}
+\usepackage{wrapfig}
 
@@
 \setlength{\columnseprule}{0mm} % No horizontal rule between columns
-\def\diag{\mathop{\rm diag}\nolimits}%
-\def\argmin{\mathop{\rm arg \; min}\limits}%
 
@@
 \definecolor{CPU}{RGB}{0, 153, 125} % Defines the color used for content box headers
+
+% Math operators
+\newcommand{\scal}[2]{\left\langle #1 , #2 \right\rangle}
+\DeclareMathOperator{\IR}{\mathbb{R}}
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator{\One}{\mathbbm{1}}
+\DeclareMathOperator{\Ccal}{\mathcal{C}}
+\DeclareMathOperator{\logsumexp}{logsumexp}
+\DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\KL}{KL}
+\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
+\renewcommand{\epsilon}{\varepsilon}
 
@@
 {\includegraphics[height=5em]{logo_ens_psl.png}} % First university/lab logo on the left
-{\bf Overrelaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport\vspace{0.2em}} 
+{\bf Overrelaxed Sinkhorn--Knopp algorithm\\ for regularized optimal transport\vspace{0.2em}} 
 {{Alexis Thibault, L\'enaïc Chizat, Charles Dossal and  Nicolas Papadakis}} % Author names and institution
@@
 
+% Introduction
+
+\headerbox{Introduction}{name=intro,column=0}{
+	Optimal Transport is an efficient and flexible tool to compare two probability distributions.
+It has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
+The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64} scaling with high dimensional data.\\
+	\begin{minipage}{0.49\textwidth}
+	\begin{center}
+		\includegraphics[width=5cm]{schema_c}
+	\end{center}
+	\caption{A wrapped figure going nicely inside the text.}\label{wrap-fig:1}
+	\end{minipage}
+	\begin{minipage}{0.49\textwidth}
+		Optimal Transport is an efficient and flexible tool to compare two probability distributions.
+It has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
+The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64} scaling with high dimensional data.
+	\end{minipage}
+}
+
 %----------------------------------------------------------------------------------------
@@
 
-\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0}{
-
+\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro}{
+	We aim at numerically solving the entropic regularization of optimal transport~:
+	\begin{equation} \label{eq:problem}
+	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
+	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One)
+	\end{equation}
 }

### 2017-11-20 PosterNIPS2017.tex
Poster

@@
 \documentclass[portrait,a0paper,fontscale=0.35]{baposter} % Adjust the font scale/size here : fontscale=0.285
-\usepackage{textpos}
+%\usepackage{textpos}
 \usepackage[frenchb]{babel}
@@
 \usetikzlibrary{shapes,arrows} % Tikz libraries required for the flow chart in the template
-\usepackage{dsfont}
-\usepackage{listings}
+%\usepackage{dsfont}
+%\usepackage{listings}
 \usepackage{subfigure}
 \usepackage{color}
-\usepackage{nth}
-\usepackage{a4wide}
+%\usepackage{nth}
+%\usepackage{a4wide}
 \usepackage{graphicx,color}
 \usepackage{array}
-\usepackage[nottoc, notlof, notlot]{tocbibind}
+%\usepackage[nottoc, notlof, notlot]{tocbibind}
 \usepackage{subfigure}
@@
 \geometry{left=0.5cm,right=0.5cm,bottom=0.5cm,top=0.5cm,centering}
-\usepackage{subfig}
+%\usepackage{subfig}
 \usepackage{float}
@@
 \usepackage{bbm}
-\usepackage{wrapfig}
+%\usepackage{wrapfig}
 
@@
 	Optimal Transport is an efficient and flexible tool to compare two probability distributions.
-It has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
-The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64} scaling with high dimensional data.\\
-	\begin{minipage}{0.49\textwidth}
+It has been popularized in the computer vision community in the context of discrete histograms [Rubner, 2000].
+The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964c] scaling with high dimensional data.\\
 	\begin{center}
-		\includegraphics[width=5cm]{schema_c}
+		\begin{tabular}{c c c c}
+		\includegraphics[height=3cm]{1d_interp_mu} &
+		\includegraphics[height=3cm]{1d_interp_0_2} &
+		\includegraphics[height=3cm]{1d_interp_2} &
+		\includegraphics[height=3cm]{1d_interp_18} \\
+		&
+		$\epsilon = 5$ &
+		$\epsilon = 0.5$ &
+		$\epsilon = 0.02$
+		\end{tabular}
+	\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.}
 	\end{center}
-	\caption{A wrapped figure going nicely inside the text.}\label{wrap-fig:1}
-	\end{minipage}
-	\begin{minipage}{0.49\textwidth}
-		Optimal Transport is an efficient and flexible tool to compare two probability distributions.
-It has been popularized in the computer vision community in the context of discrete histograms \cite{Rubner2000}.
-The introduction of entropic regularization in \cite{cuturi13} has made possible the use of the fast Sinkhorn--Knopp algorithm \cite{sinkhorn64} scaling with high dimensional data.
-	\end{minipage}
+	For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
 }
@@
 \headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro}{
-	We aim at numerically solving the entropic regularization of optimal transport~:
+	We aim at numerically solving the entropic regularization of optimal transport:
 	\begin{equation} \label{eq:problem}
 	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
-	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One)
+	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
 	\end{equation}
+	In this equation, $\Ccal_1$ and $\Ccal_2$ are the marginal constraint sets corresponding to two discrete measures $\mu^1$ and $\mu^2$,
+	and where $\KL$ is the Kullback--Leibler divergence.
+	\begin{gather*}
+	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
+	\quad\quad\quad
+	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
+	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
+	\end{gather*}
+	As it is strictly convex, it has exactly one solution. It can be written equivalently as the \emph{Bregman projection} of a point onto a convex set:
+	\begin{gather*}
+	P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
+	\gamma^* = P_{\Ccal_1 \cap \Ccal_2}(e^{-c/\epsilon}).
+	\end{gather*}
+	Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is much easier: indeed, it corresponds to scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
+	Therefore the solution can be computed by alternating Bregman projections, as shown in [Bregman, 1964]:
+	\[
+	\lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
+	\]
 }

### 2017-11-21 PosterNIPS2017.tex
Poster

@@
 \usepackage{bbm}
-%\usepackage{wrapfig}
+\usepackage{tcolorbox}
 
@@
 The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964c] scaling with high dimensional data.\\
-	\begin{center}
+	\begin{minipage}{\textwidth}
+		\centering
+		\begin{tcolorbox}[width=0.99\textwidth,/tcb/size=small, /tcb/top=2.0mm]
+			\centering
 			\begin{tabular}{c c c c}
@@
 			\end{tabular}
-	\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.}
-	\end{center}
+			\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
+		\end{tcolorbox}
+	\end{minipage}
 	For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
@@
 	\end{gather*}
-	As it is strictly convex, it has exactly one solution. It can be written equivalently as the \emph{Bregman projection} of a point onto a convex set:
+	\begin{minipage}{0.6\textwidth}
+		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution. It can be written equivalently as the \emph{Bregman projection} of a point onto a convex set:
 		\begin{gather*}
 		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
-	\gamma^* = P_{\Ccal_1 \cap \Ccal_2}(e^{-c/\epsilon}).
+		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(e^{-c/\epsilon}).
 		\end{gather*}
-	Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is much easier: indeed, it corresponds to scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
-	Therefore the solution can be computed by alternating Bregman projections, as shown in [Bregman, 1964]:
-	\[
+		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is much easier.
+Indeed, it corresponds to scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
+		\begin{align*}\label{scaling}
+		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
+		a &=  {\mu^1}\oslash{A_1 \gamma} \\
+		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
+		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
+		\end{align*}
+	\end{minipage}
+	\begin{minipage}{0.39\textwidth}
+		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
+		\centering
+		\input{schema_a}
+		\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
+		\end{tcolorbox}
+	\end{minipage}
+
+	As shown in [Bregman, 1964], the solution can be computed by alternating Bregman projections. The main idea of the Sinkhorn--Knopp algorithm can therefore be written as:
+\begin{equation}
 \lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
-	\]
+\end{equation}
 }
@@
 \headerbox{Overrelaxation}{name=overrelaxation,column=1}{
-
+Although the SK algorithm is much more efficient than linear solvers in a high-dimensional setting, it becomes extremely slow when the regularization $\epsilon$ is small.
+To counteract this limitation, we use the principle of \emph{overrelaxation}.
+\begin{minipage}{0.5\textwidth}
+	\begin{tcolorbox}[size=small, top=2mm]
+		\centering
+		\begin{tabular}{c c}
+		\input{schema_b} &
+		\input{schema_c} \\
+		SK &
+		Overrelaxed SK
+		\end{tabular}
+	\captionof{figure}{For small regularization $\epsilon$, the SK algorithm converges slowly, whereas the overrelaxed SK algorithm converges with a better rate}
+	\end{tcolorbox}
+\end{minipage}
+\begin{minipage}{0.49\textwidth}
+	We define the overrelaxed projections of parameter $\theta \in \IR$:
+	\begin{align*}\label{or_scaling}
+	P_{\Ccal_1}^\theta(\gamma) &= \diag(a)^\theta \gamma\\
+	P_{\Ccal_2}^\theta(\gamma) &= \gamma \diag(b)^\theta \nonumber
+	\end{align*}
+	(element-wise exponentiation)
+\end{minipage}
 }

### 2017-11-22 PosterNIPS2017.tex
Poster (WIP)

@@
 %
-{\includegraphics[height=5em]{logo_ens_psl.png}} % First university/lab logo on the left
+{\includegraphics[height=6em]{logo_ens_psl.png}} % First university/lab logo on the left
 {\bf Overrelaxed Sinkhorn--Knopp algorithm\\ for regularized optimal transport\vspace{0.2em}} 
 {{Alexis Thibault, L\'enaïc Chizat, Charles Dossal and  Nicolas Papadakis}} % Author names and institution
-{\includegraphics[height=5em]{logo_imb.png}} % Second university/lab logo on the right
+{\includegraphics[height=6em]{logo_imb.png}} % Second university/lab logo on the right
 
@@
 	For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
+	This idea has been shown to successfully accelerate the algorithm. However, there is no proof of convergence in the literature, and no automated choice of parameters.
+	In our paper, we suggest a way of ensuring global convergence.
+This is done simply by choosing the overrelaxation parameters at each step of the algorithm so that a given function decreases.
+We will motivate our choice of function with numerical results.
 }
@@
 \headerbox{Overrelaxation}{name=overrelaxation,column=1}{
-Although the SK algorithm is much more efficient than linear solvers in a high-dimensional setting, it becomes extremely slow when the regularization $\epsilon$ is small.
-To counteract this limitation, we use the principle of \emph{overrelaxation}.
-\begin{minipage}{0.5\textwidth}
+\begin{minipage}{0.49\textwidth}
 	\begin{tcolorbox}[size=small, top=2mm]
@@
 \begin{minipage}{0.49\textwidth}
+	\vspace{0.6mm}
+	Although the SK algorithm is much more efficient than linear solvers in a high-dimensional setting, it becomes extremely slow when the regularization $\epsilon$ is small.
+	To counteract this limitation, we use the principle of \emph{overrelaxation}.
+
 	We define the overrelaxed projections of parameter $\theta \in \IR$:
@@
 	P_{\Ccal_1}^\theta(\gamma) &= \diag(a)^\theta \gamma\\
-	P_{\Ccal_2}^\theta(\gamma) &= \gamma \diag(b)^\theta \nonumber
+	P_{\Ccal_2}^\theta(\gamma) &= \gamma \diag(b)^\theta \nonumber .
 	\end{align*}
-	(element-wise exponentiation)
+	In these equations, the exponentiation is applied element-wise.
+
+	In the euclidean setting, which also corresponds to the linearization of the operators around the solution, iterative overrelaxed projections can be shown to converge for any $\theta \in (0,2)$.
 \end{minipage}
+But in general, there is no proof of global convergence for iterative overrelaxed Bregman projections. In order to ensure convergence, we use a Lyapunov function.
 }
@@
 
-\headerbox{Lyapunov Function}{name=lyapunov,below=Sinkhorn,column=0}{
+\headerbox{Lyapunov Function}{name=lyapunov,column=1,below=overrelaxation,bottomaligned=Sinkhorn}{
 
@@
 
-\headerbox{Experiments}{name=expe,below=overrelaxation,column=1}{
+\headerbox{Experiments}{name=expe,below=Sinkhorn,column=0,span=2}{
 
@@
 
-\headerbox{References}{name=references,below=lyapunov,span =2}{
+\headerbox{References}{name=references,below=expe,span =2}{
 \noindent [1] M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\

### 2017-11-23 PosterNIPS2017.tex
Poster

@@
 Indeed, it corresponds to scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
+		Operator $\oslash$ denotes element-wise division:
 		\begin{align*}\label{scaling}
@@
 \end{minipage}
-But in general, there is no proof of global convergence for iterative overrelaxed Bregman projections. In order to ensure convergence, we use a Lyapunov function.
 }
@@
 \headerbox{Lyapunov Function}{name=lyapunov,column=1,below=overrelaxation,bottomaligned=Sinkhorn}{
+	There is no general proof of global convergence for iterative overrelaxed Bregman projections. In order to ensure convergence, we use a Lyapunov function.
+	\begin{tcolorbox}[size=small]
+    	Choose $F(\gamma)$ continuous, coercive function whose unique minimizer is $\gamma^*$.\\
+    	Choose $\theta_1(\gamma)$, $\theta_2(\gamma)$ continuous functions, such that
+    	\begin{equation*}\label{eq:cond_theta_k}
+    	\forall k \in \{1,2\}, \, \forall \gamma \notin \Ccal_k,\quad\quad
+    	F(P_{\Ccal_k}^{\theta_k(\gamma)}(\gamma)) < F(\gamma).
+    	\end{equation*}
+    	Alternately iterate $P_{\Ccal_1}^{\theta_1}$ and $P_{\Ccal_2}^{\theta_2}$, starting from $\gamma^0 = e^{-c/\epsilon}$.\\
+    	This process converges to $\gamma^*$.
+    \end{tcolorbox}
+    Our choice of Lyapunov function is: \[F(\gamma) := \KL(\gamma^*,\gamma)\]
 
+    It is motivated by the fact that this function decreases for regular Bregman projections.
+    Moreover, we can check numerically
+
+    \begin{minipage}{0.35\textwidth}
+    	\begin{tcolorbox}[size=small]
+    	\centering
+    	\includegraphics[width=\textwidth]{images/cvgce_zone_2.png}
+    	\captionof{figure}{
+    		Function $\varphi_\theta(x)$ is positive above the red line, negative below.}
+    	\end{tcolorbox}
+    \end{minipage}
+    \begin{minipage}{0.65\textwidth}
+    	The difference between two evaluations of the function can be computed directly thanks to the formula:
+    	\begin{equation*} \label{eq:kl_diff_scal}
+    	F(\gamma) - F(P^\theta_{\Ccal_k}(\gamma)) = 
+    	\scal{\mu^k}{\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)}.
+    	\end{equation*}
+    	We want this quantity to be positive.
+A sufficient condition is that every coordinate of $\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)$ be positive.
+Functions $\theta_1, \theta_2$ are crafted for this purpose.
+    	\begin{align*}
+    	\Theta^*(u) &:= \max \left\{\theta \mid \varphi_\theta(u) \ge 0 \right\}
+    	\\
+    	\Theta(u) &:= clip_{[1,\theta_0]}(\Theta^*(\min u)-\delta)\\
+    	k \in \{1,2\}, \quad \theta_k (\gamma) &:= \Theta((A_k \gamma) \oslash \mu^k)
+    	\end{align*}
+    \end{minipage}
 }

### 2017-11-24 PosterNIPS2017.tex
Poster

@@
 
-\documentclass[portrait,a0paper,fontscale=0.35]{baposter} % Adjust the font scale/size here : fontscale=0.285
+\documentclass[landscape,a0paper,fontscale=0.35]{baposter} % Adjust the font scale/size here : fontscale=0.285
 %\usepackage{textpos}
@@
 linewidth=1.8pt, % Width of the border lines around content boxes
-columns=2,
+columns=3,
 }
@@
 
-\headerbox{Introduction}{name=intro,column=0}{
+\headerbox{Introduction}{name=intro,column=0,span=2}{
+	\begin{minipage}{0.55\textwidth}
 		Optimal Transport is an efficient and flexible tool to compare two probability distributions.
-It has been popularized in the computer vision community in the context of discrete histograms [Rubner, 2000].
-The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964c] scaling with high dimensional data.\\
-	\begin{minipage}{\textwidth}
+It has been popularized in the computer vision community in the context of discrete histograms.
+The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964] scaling with high dimensional data.
+
+		For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
+		This idea has been shown to successfully accelerate the algorithm. However, there is no proof of convergence in the literature, and no automated choice of parameters.
+		In our paper, we suggest a way of ensuring global convergence.
+This is done simply by choosing the overrelaxation parameters at each step of the algorithm so that a given function decreases.
+We will motivate our choice of function with numerical results.
+
+		The algorithm we obtain converges faster than Sinkhorn--Knopp, while remaining parallelizable, and having almost the same cost per iteration.
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
 		\centering
@@
 	\end{minipage}
-	For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
-	This idea has been shown to successfully accelerate the algorithm. However, there is no proof of convergence in the literature, and no automated choice of parameters.
-	In our paper, we suggest a way of ensuring global convergence.
-This is done simply by choosing the overrelaxation parameters at each step of the algorithm so that a given function decreases.
-We will motivate our choice of function with numerical results.
-}
-
-%----------------------------------------------------------------------------------------
-%	Notations and definitions
-%----------------------------------------------------------------------------------------
-
-\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro}{
-	We aim at numerically solving the entropic regularization of optimal transport:
-	\begin{equation} \label{eq:problem}
-	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
-	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
-	\end{equation}
-	In this equation, $\Ccal_1$ and $\Ccal_2$ are the marginal constraint sets corresponding to two discrete measures $\mu^1$ and $\mu^2$,
-	and where $\KL$ is the Kullback--Leibler divergence.
-	\begin{gather*}
-	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
-	\quad\quad\quad
-	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
-	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
-	\end{gather*}
-	\begin{minipage}{0.6\textwidth}
-		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution. It can be written equivalently as the \emph{Bregman projection} of a point onto a convex set:
-		\begin{gather*}
-		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
-		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(e^{-c/\epsilon}).
-		\end{gather*}
-		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is much easier.
-Indeed, it corresponds to scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
-		Operator $\oslash$ denotes element-wise division:
-		\begin{align*}\label{scaling}
-		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
-		a &=  {\mu^1}\oslash{A_1 \gamma} \\
-		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
-		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
-		\end{align*}
-	\end{minipage}
-	\begin{minipage}{0.39\textwidth}
-		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
-		\centering
-		\input{schema_a}
-		\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
-		\end{tcolorbox}
-	\end{minipage}
-
-	As shown in [Bregman, 1964], the solution can be computed by alternating Bregman projections. The main idea of the Sinkhorn--Knopp algorithm can therefore be written as:
-\begin{equation}
-\lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
-\end{equation}
 }
@@
 
-\headerbox{Overrelaxation}{name=overrelaxation,column=1}{
+\headerbox{Overrelaxation}{name=overrelaxation,column=2}{
 \begin{minipage}{0.49\textwidth}
@@
 %----------------------------------------------------------------------------------------
+%	Bootstrap
+%----------------------------------------------------------------------------------------
+
+\headerbox{Experiments}{name=expe,below=overrelaxation,column=2}{
+	\begin{minipage}{\textwidth}
+		We compare our algorithm to SK algorithm on two different optimal transport settings.
+		\begin{itemize}
+			\item In setting (a) we consider the domain $[0,1]$ discretized into $100$ samples and the squared Euclidean transport cost on this domain.
+			The marginals are densities made of the sum of a base plateau of height $0.1$ and another plateau of height and boundaries chosen uniformly in $[0,1]$, subsequently normalized.
+			\item In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.
+		\end{itemize}
+		We plot the relative speed of our algorithm compared to Sinkhorn, for various values of the regularization parameter $\epsilon$.
+In both experimental settings, our algorithm wins by a factor above $20$ for small values of $\epsilon$.
+	\end{minipage}
+	\begin{minipage}{\textwidth}
+		\begin{tcolorbox}[size=small]
+			\begin{minipage}[b]{.5\linewidth}
+				\centering
+				\includegraphics[scale=0.5]{images/speedratio_image}\\
+				{\centering\small (a) Quadratic cost, random marginals}\label{fig:compare_a}
+			\end{minipage}%
+			\begin{minipage}[b]{.5\linewidth}
+				\centering
+				\includegraphics[scale=0.5]{images/speedratio_ML}\\
+				{\centering\small (b) Random cost, uniform marginals}\label{fig:compare_b}
+			\end{minipage}
+		\end{tcolorbox}
+	\end{minipage}
+}
+
+%----------------------------------------------------------------------------------------
+%	REFERENCES
+%----------------------------------------------------------------------------------------
+
+\headerbox{References}{name=references,below=expe,column=2}{
+	\small
+	\begin{tabular}{r p{0.67\textwidth}}
+		{[Cuturi, 2013]} & M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
+		{[Benamou et al., 2015]} & Benamou, J.-D., Carlier, G., Cuturi, M., Nenna, L., and Peyr\'e, G. (2015).
+		{\it Iterative Bregman projections for regularized transportation problems.} SIAM Journal on Scientific
+		Computing, 37(2):A1111–A1138.\\
+		{[Sinkhorn, 1964]} & Sinkhorn, R. (1964). {\it A relationship between arbitrary positive matrices and doubly stochastic matrices.} The annals of mathematical statistics, 35(2):876–879.
+	\end{tabular}
+}
+
+%----------------------------------------------------------------------------------------
+%	Notations and definitions
+%----------------------------------------------------------------------------------------
+
+\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro,bottomaligned=references}{
+	We aim at numerically solving the entropic regularization of optimal transport:
+	\begin{equation} \label{eq:problem}
+	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
+	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
+	\end{equation}
+	In this equation, $\Ccal_1$ and $\Ccal_2$ are the marginal constraint sets corresponding to two discrete measures $\mu^1$ and $\mu^2$,
+	and where $\KL$ is the Kullback--Leibler divergence.
+	\begin{gather*}
+	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
+	\quad\quad\quad
+	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
+	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
+	\end{gather*}
+	\begin{minipage}{0.6\textwidth}
+		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \emph{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
+		\begin{gather*}
+		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
+		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0).
+		\end{gather*}
+		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is simply performed by scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
+		Operator $\oslash$ denotes element-wise division:
+		\begin{align*}\label{scaling}
+		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
+		a &=  {\mu^1}\oslash{A_1 \gamma} \\
+		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
+		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
+		\end{align*}
+	\end{minipage}
+	\begin{minipage}{0.39\textwidth}
+		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
+			\centering
+			\input{schema_a}
+			\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
+		\end{tcolorbox}
+	\end{minipage}
+
+	As shown in [Bregman, 1964], the solution can be computed by alternating Bregman projections. The main idea of the Sinkhorn--Knopp algorithm can therefore be written as:
+	\begin{equation}
+	\lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
+	\end{equation}
+	This algorithm converges linearly: if we denote by $\gamma^\ell = (P_{\Ccal_2} \circ P_{\Ccal_1})^{(\ell)}(\gamma^0)$ the $\ell$-th iterate, then
+	\[ \norm{\gamma^\ell - \gamma^*} = \mathcal{O}\left((1-\eta)^\ell \right) , \quad \eta > 0 \]
+}
+
+%----------------------------------------------------------------------------------------
 % Main result
@@
 
-\headerbox{Lyapunov Function}{name=lyapunov,column=1,below=overrelaxation,bottomaligned=Sinkhorn}{
+\headerbox{Lyapunov Function}{name=lyapunov,column=1,below=intro,bottomaligned=Sinkhorn}{
 	There is no general proof of global convergence for iterative overrelaxed Bregman projections. In order to ensure convergence, we use a Lyapunov function.
@@
     Our choice of Lyapunov function is: \[F(\gamma) := \KL(\gamma^*,\gamma)\]
-
-    It is motivated by the fact that this function decreases for regular Bregman projections.
-    Moreover, we can check numerically
+    It should be noted that this function decreases for regular Bregman projections.
 
@@
     	\end{equation*}
-    	We want this quantity to be positive.
+    	We want this difference to be positive.
 A sufficient condition is that every coordinate of $\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)$ be positive.
 Functions $\theta_1, \theta_2$ are crafted for this purpose.
-    	\begin{align*}
-    	\Theta^*(u) &:= \max \left\{\theta \mid \varphi_\theta(u) \ge 0 \right\}
+    	\begin{gather*}
+    	\Theta^*(u) := \max \left\{\theta \mid \varphi_\theta(u) \ge 0 \right\}
     	\\
-    	\Theta(u) &:= clip_{[1,\theta_0]}(\Theta^*(\min u)-\delta)\\
-    	k \in \{1,2\}, \quad \theta_k (\gamma) &:= \Theta((A_k \gamma) \oslash \mu^k)
-    	\end{align*}
+    	\Theta(u) := clip_{[1,\theta_0]}(\Theta^*(\min u)-\delta)\\
+    	k \in \{1,2\}, \quad \theta_k (\gamma) := \Theta((A_k \gamma) \oslash \mu^k)
+    	\end{gather*}
     \end{minipage}
-}
-
-%----------------------------------------------------------------------------------------
-%	Bootstrap
-%----------------------------------------------------------------------------------------
-
-\headerbox{Experiments}{name=expe,below=Sinkhorn,column=0,span=2}{
-
-}
-
-%----------------------------------------------------------------------------------------
-%	REFERENCES
-%----------------------------------------------------------------------------------------
-
-\headerbox{References}{name=references,below=expe,span =2}{
-\noindent [1] M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
-...
+    This adaptive choice allows to choose the overrelaxation parameter $\theta_0$ that is used asymptotically.
+A linear study around the solution shows that overrelaxation modifies the eigenvalues in a predictable way: in particular, if the linear convergence rate of the Sinkhorn--Knopp algorithm is $1-\eta$, then the optimal asymptotic overrelaxation parameter is $\theta_0 = \frac{2}{1+\sqrt{\eta}}$.
 }

### 2017-11-24 PosterNIPS2017.tex
Poster

@@
 \usepackage{geometry}
-\geometry{left=0.5cm,right=0.5cm,bottom=0.5cm,top=0.5cm,centering}
+\geometry{left=0.9cm,right=0.9cm,bottom=0.9cm,top=0.9cm,centering}
 %\usepackage{subfig}
@@
 				\centering
-				\includegraphics[scale=0.5]{images/speedratio_image}\\
+				\includegraphics[scale=0.4]{images/speedratio_image}\\
 				{\centering\small (a) Quadratic cost, random marginals}\label{fig:compare_a}
@@
 				\centering
-				\includegraphics[scale=0.5]{images/speedratio_ML}\\
+				\includegraphics[scale=0.4]{images/speedratio_ML}\\
 				{\centering\small (b) Random cost, uniform marginals}\label{fig:compare_b}

### 2017-11-25 PosterNIPS2017.tex
Poster

@@
 
-		The algorithm we obtain converges faster than Sinkhorn--Knopp, while remaining parallelizable, and having almost the same cost per iteration.
+		The algorithm we obtain converges between 2 and 20 times faster than Sinkhorn--Knopp, while remaining parallelizable, and having almost the same cost per iteration.
 	\end{minipage}
@@
 
-\headerbox{Overrelaxation}{name=overrelaxation,column=2}{
+\headerbox{Overrelaxation}{name=overrelaxation,column=1,below=intro}{
 \begin{minipage}{0.49\textwidth}
@@
 	In these equations, the exponentiation is applied element-wise.
-
-	In the euclidean setting, which also corresponds to the linearization of the operators around the solution, iterative overrelaxed projections can be shown to converge for any $\theta \in (0,2)$.
 \end{minipage}
-}
-
-%----------------------------------------------------------------------------------------
-%	Bootstrap
-%----------------------------------------------------------------------------------------
-
-\headerbox{Experiments}{name=expe,below=overrelaxation,column=2}{
-	\begin{minipage}{\textwidth}
-		We compare our algorithm to SK algorithm on two different optimal transport settings.
-		\begin{itemize}
-			\item In setting (a) we consider the domain $[0,1]$ discretized into $100$ samples and the squared Euclidean transport cost on this domain.
-			The marginals are densities made of the sum of a base plateau of height $0.1$ and another plateau of height and boundaries chosen uniformly in $[0,1]$, subsequently normalized.
-			\item In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.
-		\end{itemize}
-		We plot the relative speed of our algorithm compared to Sinkhorn, for various values of the regularization parameter $\epsilon$.
-In both experimental settings, our algorithm wins by a factor above $20$ for small values of $\epsilon$.
-	\end{minipage}
-	\begin{minipage}{\textwidth}
-		\begin{tcolorbox}[size=small]
-			\begin{minipage}[b]{.5\linewidth}
-				\centering
-				\includegraphics[scale=0.4]{images/speedratio_image}\\
-				{\centering\small (a) Quadratic cost, random marginals}\label{fig:compare_a}
-			\end{minipage}%
-			\begin{minipage}[b]{.5\linewidth}
-				\centering
-				\includegraphics[scale=0.4]{images/speedratio_ML}\\
-				{\centering\small (b) Random cost, uniform marginals}\label{fig:compare_b}
-			\end{minipage}
-		\end{tcolorbox}
-	\end{minipage}
-}
-
-%----------------------------------------------------------------------------------------
-%	REFERENCES
-%----------------------------------------------------------------------------------------
-
-\headerbox{References}{name=references,below=expe,column=2}{
-	\small
-	\begin{tabular}{r p{0.67\textwidth}}
-		{[Cuturi, 2013]} & M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
-		{[Benamou et al., 2015]} & Benamou, J.-D., Carlier, G., Cuturi, M., Nenna, L., and Peyr\'e, G. (2015).
-		{\it Iterative Bregman projections for regularized transportation problems.} SIAM Journal on Scientific
-		Computing, 37(2):A1111–A1138.\\
-		{[Sinkhorn, 1964]} & Sinkhorn, R. (1964). {\it A relationship between arbitrary positive matrices and doubly stochastic matrices.} The annals of mathematical statistics, 35(2):876–879.
-	\end{tabular}
+As in the Sinkhorn--Knopp algorithm, we iterate these operators alternately.
+If we consider the linearization of the operators around the solution, iterative overrelaxed projections can be shown to converge for any $\theta \in (0,2)$.
+In the global, non-euclidean setting, we use an additional criterion based on a Lyapunov function.
 }
@@
 
-\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro,bottomaligned=references}{
+\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro}{
 	We aim at numerically solving the entropic regularization of optimal transport:
@@
 	\end{gather*}
-	\begin{minipage}{0.6\textwidth}
+	\begin{minipage}{0.55\textwidth}
 		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \emph{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
@@
 	\end{minipage}
-	\begin{minipage}{0.39\textwidth}
+	\begin{minipage}{0.44\textwidth}
 		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
@@
 
-\headerbox{Lyapunov Function}{name=lyapunov,column=1,below=intro,bottomaligned=Sinkhorn}{
-	There is no general proof of global convergence for iterative overrelaxed Bregman projections. In order to ensure convergence, we use a Lyapunov function.
+\headerbox{Lyapunov Function}{name=lyapunov,column=2}{
+	Although the overrelaxation of iterative Bregman projections as presented here is already in use in the context of numerical calculations of optimal transport, there is no proof that this algorithm converges to the solution.
+In order to ensure this result, we use a Lyapunov function, and slightly modify the algorithm.
 	\begin{tcolorbox}[size=small]
@@
 
-    \begin{minipage}{0.35\textwidth}
-    	\begin{tcolorbox}[size=small]
-    	\centering
-    	\includegraphics[width=\textwidth]{images/cvgce_zone_2.png}
-    	\captionof{figure}{
-    		Function $\varphi_\theta(x)$ is positive above the red line, negative below.}
-    	\end{tcolorbox}
-    \end{minipage}
+    The difference between two evaluations of the function can be computed directly thanks to the formula:
+
     \begin{minipage}{0.65\textwidth}
-    	The difference between two evaluations of the function can be computed directly thanks to the formula:
     	\begin{equation*} \label{eq:kl_diff_scal}
@@
     \end{minipage}
+	\begin{minipage}{0.34\textwidth}
+		\begin{tcolorbox}[size=small]
+			\centering
+			\includegraphics[width=\textwidth]{images/cvgce_zone_2.png}
+			\captionof{figure}{
+				Function $\varphi_\theta(x)$ is positive above the red line, negative below.}
+		\end{tcolorbox}
+	\end{minipage}
 This adaptive choice allows to choose the overrelaxation parameter $\theta_0$ that is used asymptotically.
 A linear study around the solution shows that overrelaxation modifies the eigenvalues in a predictable way: in particular, if the linear convergence rate of the Sinkhorn--Knopp algorithm is $1-\eta$, then the optimal asymptotic overrelaxation parameter is $\theta_0 = \frac{2}{1+\sqrt{\eta}}$.
+}
+
+%----------------------------------------------------------------------------------------
+%	Bootstrap
+%----------------------------------------------------------------------------------------
+
+\headerbox{Experiments}{name=expe,below=lyapunov,column=2}{
+	\begin{minipage}{\textwidth}
+		We compare our algorithm to SK algorithm on two different settings.
+		\begin{itemize}
+			\item In setting (a) we consider the domain $[0,1]$ discretized into $100$ samples and the squared Euclidean cost.
+			The marginals are randomly generated, piecewise constant on $[0,1]$.
+			\item In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.
+		\end{itemize}
+		We plot the relative speed of our algorithm compared to Sinkhorn, for various values of the regularization parameter $\epsilon$.
+In both experimental settings, our algorithm wins by a factor above $20$ for small values of $\epsilon$.
+	\end{minipage}
+	\begin{minipage}{\textwidth}
+		\begin{tcolorbox}[size=small]
+			\begin{minipage}[b]{.5\linewidth}
+				\centering
+				\includegraphics[scale=0.35]{images/speedratio_image}\\
+				{\centering\small (a) Quadratic cost, random marginals}\label{fig:compare_a}
+			\end{minipage}%
+			\begin{minipage}[b]{.5\linewidth}
+				\centering
+				\includegraphics[scale=0.35]{images/speedratio_ML}\\
+				{\centering\small (b) Random cost, uniform marginals}\label{fig:compare_b}
+			\end{minipage}
+		\end{tcolorbox}
+	\end{minipage}
+}
+
+%----------------------------------------------------------------------------------------
+%	REFERENCES
+%----------------------------------------------------------------------------------------
+
+\headerbox{References}{name=references,below=overrelaxation,column=1}{
+	\small
+	\begin{tabular}{r p{0.67\textwidth}}
+		{[Cuturi, 2013]} & M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
+		{[Benamou et al., 2015]} & Benamou, J.-D., Carlier, G., Cuturi, M., Nenna, L., and Peyr\'e, G. (2015).
+		{\it Iterative Bregman projections for regularized transportation problems.} SIAM Journal on Scientific
+		Computing, 37(2):A1111–A1138.\\
+		{[Sinkhorn, 1964]} & Sinkhorn, R. (1964). {\it A relationship between arbitrary positive matrices and doubly stochastic matrices.} The annals of mathematical statistics, 35(2):876–879.
+	\end{tabular}
 }

### 2017-11-25 PosterNIPS2017.tex
Poster (fini)

@@
 		The algorithm we obtain converges between 2 and 20 times faster than Sinkhorn--Knopp, while remaining parallelizable, and having almost the same cost per iteration.
+Moreover, the log-domain techniques used to stabilize the Sinkhorn--Knopp algorithm with regard to numerical approximations {[Schmitzer, 2016]} can be directly extended to this algorithm.
 	\end{minipage}
@@
 In the global, non-euclidean setting, we use an additional criterion based on a Lyapunov function.
-}
-
-%----------------------------------------------------------------------------------------
-%	Notations and definitions
-%----------------------------------------------------------------------------------------
-
-\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro}{
-	We aim at numerically solving the entropic regularization of optimal transport:
-	\begin{equation} \label{eq:problem}
-	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
-	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
-	\end{equation}
-	In this equation, $\Ccal_1$ and $\Ccal_2$ are the marginal constraint sets corresponding to two discrete measures $\mu^1$ and $\mu^2$,
-	and where $\KL$ is the Kullback--Leibler divergence.
-	\begin{gather*}
-	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
-	\quad\quad\quad
-	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
-	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
-	\end{gather*}
-	\begin{minipage}{0.55\textwidth}
-		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \emph{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
-		\begin{gather*}
-		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
-		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0).
-		\end{gather*}
-		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is simply performed by scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
-		Operator $\oslash$ denotes element-wise division:
-		\begin{align*}\label{scaling}
-		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
-		a &=  {\mu^1}\oslash{A_1 \gamma} \\
-		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
-		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
-		\end{align*}
-	\end{minipage}
-	\begin{minipage}{0.44\textwidth}
-		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
-			\centering
-			\input{schema_a}
-			\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
-		\end{tcolorbox}
-	\end{minipage}
-
-	As shown in [Bregman, 1964], the solution can be computed by alternating Bregman projections. The main idea of the Sinkhorn--Knopp algorithm can therefore be written as:
-	\begin{equation}
-	\lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
-	\end{equation}
-	This algorithm converges linearly: if we denote by $\gamma^\ell = (P_{\Ccal_2} \circ P_{\Ccal_1})^{(\ell)}(\gamma^0)$ the $\ell$-th iterate, then
-	\[ \norm{\gamma^\ell - \gamma^*} = \mathcal{O}\left((1-\eta)^\ell \right) , \quad \eta > 0 \]
 }
@@
 
-\headerbox{References}{name=references,below=overrelaxation,column=1}{
+\headerbox{References}{name=references,below=overrelaxation,column=1,bottomaligned=expe}{
 	\small
@@
 		Computing, 37(2):A1111–A1138.\\
-		{[Sinkhorn, 1964]} & Sinkhorn, R. (1964). {\it A relationship between arbitrary positive matrices and doubly stochastic matrices.} The annals of mathematical statistics, 35(2):876–879.
+		{[Sinkhorn, 1964]} & Sinkhorn, R. (1964). {\it A relationship between arbitrary positive matrices and doubly stochastic matrices.} The annals of mathematical statistics, 35(2):876–879.\\
+		{[Schmitzer, 2016]} & Schmitzer, B. 2016. Stabilized sparse scaling algorithms for entropy regularized transport problems. {\it arXiv preprint arXiv:1610.06519.}
 	\end{tabular}
+}
+
+%----------------------------------------------------------------------------------------
+%	Notations and definitions
+%----------------------------------------------------------------------------------------
+
+\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro,bottomaligned=expe}{
+	We aim at numerically solving the entropic regularization of optimal transport:
+	\begin{equation} \label{eq:problem}
+	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
+	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
+	\end{equation}
+	In this equation, $\Ccal_1$ and $\Ccal_2$ are the marginal constraint sets corresponding to two discrete measures $\mu^1$ and $\mu^2$,
+	and where $\KL$ is the Kullback--Leibler divergence.
+	\begin{gather*}
+	\Ccal_1 = \left\{ \gamma \mid A_1 \gamma = \mu^1 \right\},
+	\quad\quad\quad
+	\Ccal_2 = \left\{ \gamma \mid A_2 \gamma = \mu^2 \right\}.\\
+	\KL(\gamma,\xi) = \sum_{i,j} \gamma_{i,j} \left( \log \left( \frac{\gamma_{i,j}}{\xi_{i,j}} \right) -1  \right) + \sum_{i,j} \xi_{i,j}.
+	\end{gather*}
+	\begin{minipage}{0.55\textwidth}
+		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \emph{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
+		\begin{gather*}
+		P_{\Ccal}(\xi) := \argmin_{\gamma \in \Ccal} \KL(\gamma,\xi),\\
+		\text{(\ref{eq:problem})} \quad \Longleftrightarrow \quad \gamma^* = P_{\Ccal_1 \cap \Ccal_2}(\gamma^0).
+		\end{gather*}
+		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is simply performed by scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
+		Operator $\oslash$ denotes element-wise division:
+		\begin{align*}\label{scaling}
+		P_{\Ccal_1}(\gamma) &= \diag(a) \gamma &\text{with}\quad
+		a &=  {\mu^1}\oslash{A_1 \gamma} \\
+		P_{\Ccal_2}(\gamma) &= \gamma \diag(b) &\text{with}\quad
+		b &= {\mu^2}\oslash{A_2 \gamma}\nonumber
+		\end{align*}
+	\end{minipage}
+	\begin{minipage}{0.44\textwidth}
+		\begin{tcolorbox}[/tcb/size=small, /tcb/top=2.0mm]
+			\centering
+			\input{schema_a}
+			\captionof{figure}{The Sinkhorn--Knopp algorithm iterates Bregman projections on the two constraint sets.}
+		\end{tcolorbox}
+	\end{minipage}
+
+	As shown in [Bregman, 1964], the solution can be computed by alternating Bregman projections. The main idea of the Sinkhorn--Knopp algorithm can therefore be written as:
+	\begin{equation}
+	\lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
+	\end{equation}
+	This algorithm converges linearly: if we denote by $\gamma^\ell = (P_{\Ccal_2} \circ P_{\Ccal_1})^{(\ell)}(\gamma^0)$ the $\ell$-th iterate, then
+	\[ \norm{\gamma^\ell - \gamma^*} = \mathcal{O}\left((1-\eta)^\ell \right) , \quad \eta > 0 \]
 }

### 2017-11-26 PosterNIPS2017.tex
Poster, agrandissement courbes

--- 2017-11-25 PosterNIPS2017.tex
@@
 	\begin{minipage}{0.55\textwidth}
-		Optimal Transport is an efficient and flexible tool to compare two probability distributions.
-It has been popularized in the computer vision community in the context of discrete histograms.
-The introduction of entropic regularization in [Cuturi, 2013] has made possible the use of the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964] scaling with high dimensional data.
+		\begin{itemize}
+			\item[] Optimal Transport is an efficient and flexible tool to compare two probability distributions.
+Thanks to entropic regularization [Cuturi, 2013], the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964] can be used to numerically calculate the solution.
 
-		For small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
-		This idea has been shown to successfully accelerate the algorithm. However, there is no proof of convergence in the literature, and no automated choice of parameters.
-		In our paper, we suggest a way of ensuring global convergence.
-This is done simply by choosing the overrelaxation parameters at each step of the algorithm so that a given function decreases.
-We will motivate our choice of function with numerical results.
+			\item[] Yet, for small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
+			This idea has been numerically shown to accelerate the algorithm, but with no proof of convergence.
+			We ensure global convergence by choosing the overrelaxation parameters at each step of the algorithm so that a given Lyapunov function decreases.
 
-		The algorithm we obtain converges between 2 and 20 times faster than Sinkhorn--Knopp, while remaining parallelizable, and having almost the same cost per iteration.
+			\item[] The algorithm we obtain converges between 2 and 20 times faster than Sinkhorn--Knopp, while having almost the same cost per iteration, and retaining other nice properties.
+		\end{itemize}
 	\end{minipage}
@@
 %----------------------------------------------------------------------------------------
+% Main result
+%----------------------------------------------------------------------------------------
+
+\headerbox{Lyapunov Function}{name=lyapunov,column=2}{
+	To ensure convergence of iterative overrelaxed Bregman projections, we assert the decrease of a Lyapunov function.
+	\begin{tcolorbox}[size=small]
+    	Choose $F(\gamma)$ continuous, coercive function whose unique minimizer is $\gamma^*$.\\
+    	Choose $\theta_1(\gamma)$, $\theta_2(\gamma)$ continuous functions, such that
+    	\begin{equation*}\label{eq:cond_theta_k}
+    	\forall k \in \{1,2\}, \, \forall \gamma \notin \Ccal_k,\quad\quad
+    	F(P_{\Ccal_k}^{\theta_k(\gamma)}(\gamma)) < F(\gamma).
+    	\end{equation*}
+    	Alternately iterate $P_{\Ccal_1}^{\theta_1}$ and $P_{\Ccal_2}^{\theta_2}$, starting from $\gamma^0 = e^{-c/\epsilon}$.
+    	This process converges to $\gamma^*$.
+    \end{tcolorbox}
+    Our choice of Lyapunov function is: $F(\gamma) := \KL(\gamma^*,\gamma)$
+    Note that this function decreases for regular Bregman projections.
+    The difference between two evaluations can be computed directly:
+
+    \begin{minipage}{0.65\textwidth}
+    	\begin{equation*} \label{eq:kl_diff_scal}
+    	F(\gamma) - F(P^\theta_{\Ccal_k}(\gamma)) = 
+    	\scal{\mu^k}{\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)}.
+    	\end{equation*}
+    	We want this difference to be positive.
+A sufficient condition is that every coordinate of $\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)$ be positive.
+Functions $\theta_1, \theta_2$ are crafted for this purpose.
+    	\begin{gather*}
+    	\Theta^*(u) := \max \left\{\theta \mid \varphi_\theta(u) \ge 0 \right\}
+    	\\
+    	\Theta(u) := clip_{[1,\theta_0]}(\Theta^*(\min u)-\delta)\\
+    	k \in \{1,2\}, \quad \theta_k (\gamma) := \Theta((A_k \gamma) \oslash \mu^k)
+    	\end{gather*}
+    \end{minipage}
+	\begin{minipage}{0.34\textwidth}
+		\begin{tcolorbox}[size=small]
+			\centering
+			\includegraphics[width=\textwidth]{images/cvgce_zone_2.png}
+			\captionof{figure}{
+				Function $\varphi_\theta(x)$ is positive above the red line, negative below.}
+		\end{tcolorbox}
+	\end{minipage}
+This adaptive choice allows to choose the overrelaxation parameter $\theta_0$ that is used asymptotically.
+A linear study around the solution shows that overrelaxation modifies the eigenvalues in a predictable way: in particular, if the linear convergence rate of the Sinkhorn--Knopp algorithm is $1-\eta$, then the optimal asymptotic overrelaxation parameter is $\theta_0 = \frac{2}{1+\sqrt{\eta}}$.
+}
+
+%----------------------------------------------------------------------------------------
+%	Bootstrap
+%----------------------------------------------------------------------------------------
+
+\headerbox{Experiments}{name=expe,below=overrelaxation,column=1,span=2}{
+	\begin{minipage}{0.32\textwidth}
+		We compare our algorithm to SK algorithm on two different settings of size $100 \times 100$.
+		\begin{itemize}
+			\item[(a)] Cost $c$ is the squared euclidean norm on $[0,1]$. Marginals are randomly generated, piecewise constant.
+			\item[(b)] Cost $c$ is a $100\times 100$ random matrix with entries uniform in $[0,1]$. Marginals are uniform.
+		\end{itemize}
+		In both experimental settings, our algorithm wins by a factor above $20$ for small $\epsilon$.
+	\end{minipage}
+	\begin{minipage}{0.68\textwidth}
+		\begin{tcolorbox}[size=small]
+			\begin{minipage}{0.5\textwidth}
+				\centering
+				\includegraphics[width=\textwidth]{images/speedratio_image}
+			\end{minipage}%
+			\begin{minipage}{0.5\textwidth}
+				\centering
+				\includegraphics[width=\textwidth]{images/speedratio_ML}
+			\end{minipage}
+		\captionof{figure}{Relative speed of our algorithm compared to Sinkhorn--Knopp, for various values of $\epsilon$}
+		\end{tcolorbox}
+	\end{minipage}
+}
+
+%----------------------------------------------------------------------------------------
+%	REFERENCES
+%----------------------------------------------------------------------------------------
+
+\headerbox{References}{name=references,below=lyapunov,column=2,bottomaligned=overrelaxation}{
+	\scriptsize
+	\begin{multicols}{2}
+		\textbf{[Cuturi, 2013]} M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.
+
+		\textbf{[Benamou et al., 2015]} Benamou, J.-D., Carlier, G., Cuturi, M., Nenna, L., and Peyr\'e, G. (2015).
+		{\it Iterative Bregman projections for regularized transportation problems.} SIAM Journal on Scientific
+		Computing, 37(2):A1111–A1138.
+
+		\textbf{[Sinkhorn, 1964]} Sinkhorn, R. (1964). {\it A relationship between arbitrary positive matrices and doubly stochastic matrices.} The annals of mathematical statistics, 35(2):876–879.
+
+		\textbf{[Schmitzer, 2016]} Schmitzer, B. 2016. Stabilized sparse scaling algorithms for entropy regularized transport problems. {\it arXiv preprint arXiv:1610.06519.}
+	\end{multicols}
+}
+
+%----------------------------------------------------------------------------------------
 %	Notations and definitions
@@
 
-\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro}{
+\headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro,bottomaligned=expe}{
 	We aim at numerically solving the entropic regularization of optimal transport:
@@
 
-%----------------------------------------------------------------------------------------
-% Main result
-%----------------------------------------------------------------------------------------
-
-\headerbox{Lyapunov Function}{name=lyapunov,column=2}{
-	Although the overrelaxation of iterative Bregman projections as presented here is already in use in the context of numerical calculations of optimal transport, there is no proof that this algorithm converges to the solution.
-In order to ensure this result, we use a Lyapunov function, and slightly modify the algorithm.
-	\begin{tcolorbox}[size=small]
-    	Choose $F(\gamma)$ continuous, coercive function whose unique minimizer is $\gamma^*$.\\
-    	Choose $\theta_1(\gamma)$, $\theta_2(\gamma)$ continuous functions, such that
-    	\begin{equation*}\label{eq:cond_theta_k}
-    	\forall k \in \{1,2\}, \, \forall \gamma \notin \Ccal_k,\quad\quad
-    	F(P_{\Ccal_k}^{\theta_k(\gamma)}(\gamma)) < F(\gamma).
-    	\end{equation*}
-    	Alternately iterate $P_{\Ccal_1}^{\theta_1}$ and $P_{\Ccal_2}^{\theta_2}$, starting from $\gamma^0 = e^{-c/\epsilon}$.\\
-    	This process converges to $\gamma^*$.
-    \end{tcolorbox}
-    Our choice of Lyapunov function is: \[F(\gamma) := \KL(\gamma^*,\gamma)\]
-    It should be noted that this function decreases for regular Bregman projections.
-
-    The difference between two evaluations of the function can be computed directly thanks to the formula:
-
-    \begin{minipage}{0.65\textwidth}
-    	\begin{equation*} \label{eq:kl_diff_scal}
-    	F(\gamma) - F(P^\theta_{\Ccal_k}(\gamma)) = 
-    	\scal{\mu^k}{\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)}.
-    	\end{equation*}
-    	We want this difference to be positive.
-A sufficient condition is that every coordinate of $\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)$ be positive.
-Functions $\theta_1, \theta_2$ are crafted for this purpose.
-    	\begin{gather*}
-    	\Theta^*(u) := \max \left\{\theta \mid \varphi_\theta(u) \ge 0 \right\}
-    	\\
-    	\Theta(u) := clip_{[1,\theta_0]}(\Theta^*(\min u)-\delta)\\
-    	k \in \{1,2\}, \quad \theta_k (\gamma) := \Theta((A_k \gamma) \oslash \mu^k)
-    	\end{gather*}
-    \end{minipage}
-	\begin{minipage}{0.34\textwidth}
-		\begin{tcolorbox}[size=small]
-			\centering
-			\includegraphics[width=\textwidth]{images/cvgce_zone_2.png}
-			\captionof{figure}{
-				Function $\varphi_\theta(x)$ is positive above the red line, negative below.}
-		\end{tcolorbox}
-	\end{minipage}
-This adaptive choice allows to choose the overrelaxation parameter $\theta_0$ that is used asymptotically.
-A linear study around the solution shows that overrelaxation modifies the eigenvalues in a predictable way: in particular, if the linear convergence rate of the Sinkhorn--Knopp algorithm is $1-\eta$, then the optimal asymptotic overrelaxation parameter is $\theta_0 = \frac{2}{1+\sqrt{\eta}}$.
-}
-
-%----------------------------------------------------------------------------------------
-%	Bootstrap
-%----------------------------------------------------------------------------------------
-
-\headerbox{Experiments}{name=expe,below=lyapunov,column=2}{
-	\begin{minipage}{\textwidth}
-		We compare our algorithm to SK algorithm on two different settings.
-		\begin{itemize}
-			\item In setting (a) we consider the domain $[0,1]$ discretized into $100$ samples and the squared Euclidean cost.
-			The marginals are randomly generated, piecewise constant on $[0,1]$.
-			\item In setting (b) the cost is a $100\times 100$ random matrix with entries uniform in $[0,1]$ and the marginals are uniform.
-		\end{itemize}
-		We plot the relative speed of our algorithm compared to Sinkhorn, for various values of the regularization parameter $\epsilon$.
-In both experimental settings, our algorithm wins by a factor above $20$ for small values of $\epsilon$.
-	\end{minipage}
-	\begin{minipage}{\textwidth}
-		\begin{tcolorbox}[size=small]
-			\begin{minipage}[b]{.5\linewidth}
-				\centering
-				\includegraphics[scale=0.35]{images/speedratio_image}\\
-				{\centering\small (a) Quadratic cost, random marginals}\label{fig:compare_a}
-			\end{minipage}%
-			\begin{minipage}[b]{.5\linewidth}
-				\centering
-				\includegraphics[scale=0.35]{images/speedratio_ML}\\
-				{\centering\small (b) Random cost, uniform marginals}\label{fig:compare_b}
-			\end{minipage}
-		\end{tcolorbox}
-	\end{minipage}
-}
-
-%----------------------------------------------------------------------------------------
-%	REFERENCES
-%----------------------------------------------------------------------------------------
-
-\headerbox{References}{name=references,below=overrelaxation,column=1}{
-	\small
-	\begin{tabular}{r p{0.67\textwidth}}
-		{[Cuturi, 2013]} & M. Cuturi. Sinkhorn distances: Lightspeed computation of optimal transport. {\it Advances in Neural Information Processing Systems 26}, 2013.\\
-		{[Benamou et al., 2015]} & Benamou, J.-D., Carlier, G., Cuturi, M., Nenna, L., and Peyr\'e, G. (2015).
-		{\it Iterative Bregman projections for regularized transportation problems.} SIAM Journal on Scientific
-		Computing, 37(2):A1111–A1138.\\
-		{[Sinkhorn, 1964]} & Sinkhorn, R. (1964). {\it A relationship between arbitrary positive matrices and doubly stochastic matrices.} The annals of mathematical statistics, 35(2):876–879.
-	\end{tabular}
-}
-
 \end{poster}

### 2017-11-26 PosterNIPS2017.tex
Allegement intro poster

@@
 
-			\item[] Yet, for small regularization $\epsilon$, the SK algorithm becomes extremely slow. To counter this, we propose a new algorithm that accelerates SK by \emph{overrelaxation}.
-			This idea has been numerically shown to accelerate the algorithm, but with no proof of convergence.
-			We ensure global convergence by choosing the overrelaxation parameters at each step of the algorithm so that a given Lyapunov function decreases.
+			\item[] To counter the slowing of SK for small $\epsilon$, we propose a new algorithm based on \emph{overrelaxation}.
+			Global convergence is ensured by choosing the overrelaxation parameters at each step of the algorithm so that a given Lyapunov function decreases.
 

### 2017-11-27 PosterNIPS2017.tex
Poster : aération, ajout figure

@@
 		\begin{itemize}
-			\item[] Optimal Transport is an efficient and flexible tool to compare two probability distributions.
-Thanks to entropic regularization [Cuturi, 2013], the fast Sinkhorn--Knopp algorithm (SK) [Sinkhorn, 1964] can be used to numerically calculate the solution.
+			\item Optimal Transport: efficient and flexible tool to compare two probability distributions.
 
-			\item[] To counter the slowing of SK for small $\epsilon$, we propose a new algorithm based on \emph{overrelaxation}.
-			Global convergence is ensured by choosing the overrelaxation parameters at each step of the algorithm so that a given Lyapunov function decreases.
+			\item The fast \textbf{Sinkhorn--Knopp} algorithm (SK) can be used to numerically calculate an approximate solution [Cuturi, 2013].
 
-			\item[] The algorithm we obtain converges between 2 and 20 times faster than Sinkhorn--Knopp, while having almost the same cost per iteration, and retaining other nice properties.
+			\item Our contribution: new algorithm based on \textbf{overrelaxation} of SK.
+
+			\item Always converges to the solution
+
+			\item Adaptive choice of parameters
+
+			\item Converges \textbf{between 2 and 20 times faster} than SK, with almost the same cost per iteration, and other nice properties.
 		\end{itemize}
@@
 \headerbox{Overrelaxation}{name=overrelaxation,column=1,below=intro}{
-\begin{minipage}{0.49\textwidth}
+\begin{minipage}{0.6\textwidth}
 	\begin{tcolorbox}[size=small, top=2mm]
@@
 \end{minipage}
-\begin{minipage}{0.49\textwidth}
-	\vspace{0.6mm}
-	Although the SK algorithm is much more efficient than linear solvers in a high-dimensional setting, it becomes extremely slow when the regularization $\epsilon$ is small.
-	To counteract this limitation, we use the principle of \emph{overrelaxation}.
-
-	We define the overrelaxed projections of parameter $\theta \in \IR$:
+\begin{minipage}{0.4\textwidth}
+	\textbf{Overrelaxed projections}: elevate the scaling factors to some power $\theta \in \IR$.
 	\begin{align*}\label{or_scaling}
@@
 	\end{align*}
-	In these equations, the exponentiation is applied element-wise.
+	(element-wise exponentiation)
+
+	Iterative overrelaxed projections \textbf{converge locally} for any $\theta \in (0,2)$.
+	In the global, non-euclidean setting, we use an additional criterion based on a Lyapunov function.
 \end{minipage}
-As in the Sinkhorn--Knopp algorithm, we iterate these operators alternately.
-If we consider the linearization of the operators around the solution, iterative overrelaxed projections can be shown to converge for any $\theta \in (0,2)$.
-In the global, non-euclidean setting, we use an additional criterion based on a Lyapunov function.
+
+Overrelaxation modifies the eigenvalues of the linearized operator in a predictable way: if SK converges at rate $1-\eta$, then the optimal asymptotic overrelaxation parameter is $\theta_0 = \frac{2}{1+\sqrt{\eta}}$, for which the convergence rate is $\frac{1-\sqrt{\eta}}{1+\sqrt{\eta}}$.
 }
@@
 \headerbox{Lyapunov Function}{name=lyapunov,column=2}{
-	To ensure convergence of iterative overrelaxed Bregman projections, we assert the decrease of a Lyapunov function.
+	To ensure convergence of iterative overrelaxed Bregman projections, we assert the decrease of a \textbf{Lyapunov function}.
 	\begin{tcolorbox}[size=small]
@@
     \end{tcolorbox}
-    Our choice of Lyapunov function is: $F(\gamma) := \KL(\gamma^*,\gamma)$
+    Our choice of Lyapunov function is: $F(\gamma) := \KL(\gamma^*,\gamma)$.
     Note that this function decreases for regular Bregman projections.
-    The difference between two evaluations can be computed directly:
 
     \begin{minipage}{0.65\textwidth}
-    	\begin{equation*} \label{eq:kl_diff_scal}
+    	\vspace{0.6em}
+    	Differences of $F$ can be calculated:
+    	\begin{gather*} \label{eq:kl_diff_scal}
     	F(\gamma) - F(P^\theta_{\Ccal_k}(\gamma)) = 
-    	\scal{\mu^k}{\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)}.
-    	\end{equation*}
-    	We want this difference to be positive.
-A sufficient condition is that every coordinate of $\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)$ be positive.
-Functions $\theta_1, \theta_2$ are crafted for this purpose.
+    	\scal{\mu^k}{\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)}.\\
+    	\varphi_\theta(x) = x(1-x^{-\theta}) - \theta \log x
+    	\end{gather*}
+    	We want this difference to be positive. A sufficient condition is that every coordinate of $\varphi_\theta \left((A_k \gamma) \oslash \mu^k \right)$ be positive.
     	\begin{gather*}
@@
 This adaptive choice allows to choose the overrelaxation parameter $\theta_0$ that is used asymptotically. 
-A linear study around the solution shows that overrelaxation modifies the eigenvalues in a predictable way: in particular, if the linear convergence rate of the Sinkhorn--Knopp algorithm is $1-\eta$, then the optimal asymptotic overrelaxation parameter is $\theta_0 = \frac{2}{1+\sqrt{\eta}}$.
 }
@@
 \headerbox{Experiments}{name=expe,below=overrelaxation,column=1,span=2}{
-	\begin{minipage}{0.32\textwidth}
+	\begin{minipage}{\textwidth}
 		We compare our algorithm to SK algorithm on two different settings of size $100 \times 100$.
-		\begin{itemize}
-			\item[(a)] Cost $c$ is the squared euclidean norm on $[0,1]$. Marginals are randomly generated, piecewise constant.
-			\item[(b)] Cost $c$ is a $100\times 100$ random matrix with entries uniform in $[0,1]$. Marginals are uniform.
-		\end{itemize}
 		In both experimental settings, our algorithm wins by a factor above $20$ for small $\epsilon$.
 	\end{minipage}
-	\begin{minipage}{0.68\textwidth}
+	\begin{minipage}{0.6\textwidth}
 		\begin{tcolorbox}[size=small]
@@
 				\centering
-				\includegraphics[width=\textwidth]{images/speedratio_image}
+				\includegraphics[width=0.9\textwidth]{images/speedratio_image}\\
+				(a) Squared euclidean norm on $[0,1]$. Marginals randomly generated, piecewise constant.
 			\end{minipage}%
@@
 				\centering
-				\includegraphics[width=\textwidth]{images/speedratio_ML}
+				\includegraphics[width=0.9\textwidth]{images/speedratio_ML}\\
+				(b) Random cost, uniform marginals.
 			\end{minipage}
 		\captionof{figure}{Relative speed of our algorithm compared to Sinkhorn--Knopp, for various values of $\epsilon$}
+		\end{tcolorbox}
+	\end{minipage}
+	\begin{minipage}{0.4\textwidth}
+		\begin{tcolorbox}[size=small]
+			\centering
+			\includegraphics[width=0.89\textwidth]{typical_convergence} \\
+			\captionof{figure}{Typical behavior of the algorithm: value of the Lyapunov function and of the overrelaxation parameters during the iterations.}
 		\end{tcolorbox}
@@
 \headerbox{Sinkhorn-Knopp Algorithm}{name=Sinkhorn,column=0,below=intro,bottomaligned=expe}{
-	We aim at numerically solving the entropic regularization of optimal transport:
-	\begin{equation} \label{eq:problem}
+	Entropic regularization of optimal transport:
+	\begin{equation*} \label{eq:problem}
 	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
 	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
-	\end{equation}
-	In this equation, $\Ccal_1$ and $\Ccal_2$ are the marginal constraint sets corresponding to two discrete measures $\mu^1$ and $\mu^2$,
-	and where $\KL$ is the Kullback--Leibler divergence.
+	\end{equation*}
+	Marginal contraint sets $\Ccal_1$ and $\Ccal_2$ correspond to two discrete measures $\mu^1$ and $\mu^2$, and $\KL$ is the Kullback--Leibler divergence.
 	\begin{gather*}
@@
 	\begin{minipage}{0.55\textwidth}
-		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \emph{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
+		As problem (\ref{eq:problem}) is strictly convex, it has exactly one solution, that can be written equivalently as the \textbf{Bregman projection} of $\gamma^0 = e^{-c/\epsilon}$ onto a convex set:
 		\begin{gather*}
@@
 		\end{gather*}
-		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is simply performed by scaling the rows or columns of the matrix so that the corresponding marginal sum is correct.
+		Whereas projecting onto $\Ccal_1 \cap \Ccal_2$ is challenging, projecting onto $\Ccal_1$ or $\Ccal_2$ individually is simply performed by \textbf{scaling the rows or columns} of the matrix.
 		Operator $\oslash$ denotes element-wise division:
@@
 
-	As shown in [Bregman, 1964], the solution can be computed by alternating Bregman projections. The main idea of the Sinkhorn--Knopp algorithm can therefore be written as:
-	\begin{equation}
+	\vspace{0.6em}
+	\textbf{Sinkhorn--Knopp algorithm:}
+	\begin{equation*}
 	\lim P_{\Ccal_2}\circ P_{\Ccal_1} \circ \ldots \circ P_{\Ccal_2} \circ P_{\Ccal_1} (\gamma^0) = \gamma^*
-	\end{equation}
-	This algorithm converges linearly: if we denote by $\gamma^\ell = (P_{\Ccal_2} \circ P_{\Ccal_1})^{(\ell)}(\gamma^0)$ the $\ell$-th iterate, then
+	\end{equation*}
+	This algorithm \textbf{converges linearly}: if we denote by $\gamma^\ell = (P_{\Ccal_2} \circ P_{\Ccal_1})^{(\ell)}(\gamma^0)$ the $\ell$-th iterate, then
 	\[ \norm{\gamma^\ell - \gamma^*} = \mathcal{O}\left((1-\eta)^\ell \right) , \quad \eta > 0 \]

### 2017-11-28 PosterNIPS2017.tex
Last image of poster
Poster: Final details

@@
 			\end{tabular}
-			\captionof{figure}{Regularized optimal transport maps for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
+			\captionof{figure}{Regularized optimal transport maps become more and more precise for decreasing regularization parameter $\epsilon$.} \label{fig:1d_ot}
 		\end{tcolorbox}
@@
 			\centering
-			\includegraphics[width=0.89\textwidth]{typical_convergence} \\
-			\captionof{figure}{Typical behavior of the algorithm: value of the Lyapunov function and of the overrelaxation parameters during the iterations.}
+			\includegraphics[width=0.83\textwidth]{typical_convergence_2} \\
+			\captionof{figure}{Typical run of the algorithm: value of the Lyapunov function and of the overrelaxation parameters during the iterations.}
 		\end{tcolorbox}
@@
 	Entropic regularization of optimal transport:
-	\begin{equation*} \label{eq:problem}
+	\begin{equation} \label{eq:problem}
 	\gamma^* = \argmin_{\gamma \in \Ccal_1 \cap \Ccal_2}
 	\scal{c}{\gamma} + \epsilon \KL(\gamma,\One).
-	\end{equation*}
+	\end{equation}
 	Marginal contraint sets $\Ccal_1$ and $\Ccal_2$ correspond to two discrete measures $\mu^1$ and $\mu^2$, and $\KL$ is the Kullback--Leibler divergence.
